### 2009-05-07 edpm-paper.tex
copy of files as they stand from previous ACL'08 style submission to MetricsMATR

% -*-LaTeX-*- 
% derived from file acl08.tex
%

\documentclass[11pt]{article}
\usepackage{acl08}
\usepackage{times}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{dcolumn}
\usepackage{url}
\usepackage{comment}
\usepackage{graphicx}
\usepackage{qtree}
\setlength\titlebox{6.5cm}    % Expanding the titlebox

%\newcommand{\blindthis}[1]{#1}
\newcommand{\blindthis}[1]{}

\newcolumntype{d}{D{.}{.}{3}}
\DeclareMathOperator*{\corr}{corr}

\title{Automatic Syntactic MT Evaluation with Expected Dependency
  Pair Match\protect\blindthis{\Thanks{Thanks to Kevin Knight for providing data for
    earlier revisions of this research.}}}

\author{\blindthis{
    Jeremy G. Kahn \and Mari Ostendorf \\
    Signal, Speech, and Language Interpretation Laboratory \\
    University of Washington \\
    Seattle, WA 98195, USA\\
    {\tt \{jgk,mo\}@ssli.ee.washington.edu}
    \And
    Brian Roark\\
    OGI
  }
}
\date{}

\newcommand{\headlabel}[2]{\ensuremath{\underset{\textrm{#2}}{\textrm{#1}}}}
\newcommand{\arclabel}[1]{\ensuremath{\stackrel{#1}{\to}}}
\newcommand{\DPM}[1]{\ensuremath{\mathrm{DPM}_{#1}}}
\newcommand{\DPMempty}{\ensuremath{\DPM{}}}
% \newcommand{\EDPM}[2]{
%   \DPM{ E^{#2} \left[ #1 \right] }
% }
\newcommand{\EDPM}[2]{\ensuremath{\DPM{#1}{#2}}}
\newcommand{\mDPM}[2]{\ensuremath{\DPM{#1}{0}}}
\newcommand{\bDPM}[1]{\ensuremath{\mathrm{b}\DPM{#1}}}
\newcommand{\myEDPM}[0]{\ensuremath{\mathrm{EDPM}}}
\newcommand{\BoNG}[1]{\ensuremath{\textrm{bag-of-ngrams}(#1)}}
% \newcommand{\BoNG}[1]{
%   \ensuremath{F_\cup \left(n_1^{#1}  \textrm{-gram} \right)
%   }}
\newcommand{\bBoNG}[1]{\ensuremath{\textrm{av-bags-of-ngrams}(#1)}}
% \newcommand{\bBoNG}[1]{
%   \ensuremath{F_\forall \left(
%       n_1^#1\textrm{-gram}
%     \right)}}

\newcommand{\ngram}[0]{\ensuremath{n\textrm{-gram}}}
\newcommand{\precision}[0]{\ensuremath{\textrm{Precision}}}
\newcommand{\recall}[0]{\ensuremath{\textrm{Recall}}}

\begin{document}
\maketitle
\begin{abstract}
  % TO DO: cut further per M.O. comments?
  Previous work on dependency-based machine translation evaluation
  % \cite{owczarzak07labelleddepseval} 
  has shown that an LFG-based F-measure of labeled partial
  dependencies over an $n$-best list %(\textbf{d\_$n$\_var})
  has improved correlation with human judgments of fluency and
  adequacy as compared to BLEU %\cite{papineni02bleu}
  and TER. %\cite{snover06ter}
  Inspired by \textsc{SParseval}, % \cite{roark06:sparseval}
  we demonstrate that a statistical syntactic parser based on PCFGs
  % \cite{charniak-johnson:2005:ACL} 
  may be used in place of the LFG dependencies, and we explore
  variations on other aspects of the %\textbf{d\_$n$\_var}
  dependency-matching method. We
  find that separating the match quality into different bins, as in
  BLEU, does \emph{not} improve correlations with human judgments, but
  that including $n>1$ parses helps, especially if one incorporates
  probabilistic weights from the parser.  Additionally, we find that
  matching simple words (along with partial-dependencies) further improves
  correlations with human judgments. % beyond \textbf{d\_$n$\_var}.

  From these results, we design a new scoring metric ``Expected
  Dependency Pair Match'' (\myEDPM), and 
%   explore \myEDPM's quality
%   with respect to the GALE % \cite{olive06gale}
%   2.5 evaluations corpus.  We 
  demonstrate that $\Delta$\myEDPM{} is superior to $\Delta$BLEU and
  $\Delta$TER as a per-document and per-sentence predictor of
  $\Delta$HTER.
% , and is superior to $\Delta$TER in
%   the written contexts (newswire and web text).
\end{abstract}

\section{Introduction}

Machine translation (MT) evaluation is a challenge for research
because the space of good translations is large, and two equally good
translations may appear to be quite different at first glance. 
%
The challenges of choosing among translations are compounded when this
evaluation is done automatically.
%
Human evaluation, however, is both time-consuming and difficult, so
research % since at least \cite{papineni02bleu}
has turned increasingly towards automatic measures of translation
quality, usually by comparing the system translation to one or more
reference (human) translations.
%
Automatic measures of this kind (e.g.\ BLEU \cite{papineni02bleu}) not
only provide a well-defined evaluation standard but are also required
for training on error criteria, e.g.\ with minimum error rate training
\cite{och03mert}.

The most popular evaluation measures (BLEU and related measure
NIST \cite{doddington02nist}) are based on $n$-gram precision.  More
recent research has found that these measures may not accurately track
translation quality both empirically \cite{charniak03syntaxlmmt} and
theoretically \cite{callisonburch06bleuproblems}.

% \subsection{Syntactic machine translation evaluation}
One direction to look for improving metrics is to try to model
acceptable variation, whether word-choice (e.g.\ METEOR
\cite{banerjee05meteor}, which does progressively more forgiving
word matching) or by modeling syntactic information
\cite{liu05syntaxformteval,owczarzak07labelleddepseval}.  
\newcite{owczarzak07labelleddepseval} explore the correlation of
their measure \textbf{d} and \textbf{d\_var} with human judgment, and
report substantial improvements relative to the popular measures BLEU
and TER \cite{snover06ter}.  
%
Keeping to the syntactic approach, the work here follows and extends
the labelled-dependency match version of \textsc{SParseval}
\cite{roark06:sparseval} and the \textbf{d}/\textbf{d\_var}
\cite{owczarzak07labelleddepseval} measures, all of which evaluate
hypothesis-reference similarity with an $F$ measure over fragments of
a labelled dependency structure.  
%
The measures requiring dependency structure have generated that structure by a
PCFG with deterministic head-finding
\cite{liu05syntaxformteval,roark06:sparseval} or by extracting the
semantic dependencies from an LFG parser
(\newcite{cahill04lfg} in \newcite{owczarzak07labelleddepseval}).
In this work, we extend the dependency-scoring strategies of
\newcite{owczarzak07labelleddepseval} but do so with a
widely used and publically available PCFG parser and
deterministic head-finding rules instead of an LFG system.

% Prior work with
% \textsc{SParseval} (which has generally been used for
% speech-recognition evaluation rather than MT) has generally used a PCFG and the \textbf{d}
% measure is the source of the dependency graph; the latter uses an LFG
% parser \cite{cahill04lfg} while the former uses a PCFG parse and a
% deterministic head-finding strategy.

%% TO DO: MO: SParseval has multiple versions, only one of which uses the
%% head-finding structure. I don't think that SParseval *requires* the
%% use of a PCFG which is the model.  It simply requires a certain
%% form of the resulting parse.

%% Comparing SParseval and d might make
%% sense if d has been used for parsing as well as MT, but then you
%% should be citing the parsing paper. If not, you can distinguish the
%% LFG and PCFG aspects of the two approaches work, but should
%% separate that from SParseval and d/dvar.

%% In the list of diffs, you
%% might also talk about looking at correlation with two types of
%% human assessments: FL & Ad and HTER.

%\subsection{Evaluation of MT metrics}
These measures have been evaluated in a number of ways.  Some
\cite{banerjee05meteor,liu05syntaxformteval,owczarzak07labelleddepseval}
have evaluated their success by comparing the measure to human
judgments of fluency and adequacy.  In other work,
e.g. \newcite{snover06ter}, 
%and the GALE translation and summarization project
measures are evaluated by comparison to
HTER, a distance to a human-revised reference that uses wording closer
to the MT system choices (keeping the original meaning) that
%\emph{post hoc} human-targeted reference, which
is intended to measure the post-editing work required after
translation.  In this paper, we pursue both kinds of evaluation.

The remainder of this work is as follows. In section~\ref{sec:dpm}, we
define a family of measures \DPM{} that include \textsc{SParseval} and
\textbf{d}/\textbf{d\_var} measures and describe our strategy for
extracting labelled dependencies. In section~\ref{sec:facorr}, we
explore a selection of variants of that family with human judgments
over the Multiple Translation Chinese corpus
\cite{LDC03MTC2,LDC06MTC4}. We explore questions of dependency-graph
decomposition, binning choices, using multiple parses, and
parse-confidence and select a member of that family as a new best
measure \myEDPM{} that we believe to be a superior representative of
the \DPM{} family.  In section~\ref{sec:hter}, we explore \myEDPM{}'s
ability to predict changes in a human-generated score (HTER) over a
corpus of hypothesis translations in the GALE 2.5 \cite{darpa08gale}
% \cite{olive06gale}
% TO DO: MO: find better citation for GALE (cf. NIST web site or
% Google)
evaluation. In section~\ref{sec:conclusion}, we discuss future work
and conclude.

% Somewhere in here, mention the idea: skip-ngram but defer to parser to
% choose which ones to skip

\section{Definition of \DPMempty{} and its variants}
\label{sec:dpm}
In this work, we define a family of measures Dependency Pair Match
(\DPMempty{}) that is composed of extensions of the methods described in
\newcite{owczarzak07labelleddepseval}.  \DPMempty{} is defined as the $F$
measure (harmonic mean of precision and recall) over bags-of-subtrees
of the hypothesis translation dependency tree as compared to
bags-of-subtrees of the reference translation dependency tree.
%
\begin{figure*}
  \centering
  \begin{tabular}{rcc}
    & \textbf{Reference} & \textbf{Hypothesis} \\
    \textbf{dependency tree}
    & \includegraphics[scale=0.6]{dpm-example-ref.png} & 
    \includegraphics[scale=0.6]{dpm-example-hyp.png}\\
    \textbf{$dlh$ list} &
    \begin{tabular}{@{$\langle$}c@{,~}c@{,~}c@{$\rangle$}}
      the & \arclabel{\texttt{det}}  &  cat \\
      red & \arclabel{\texttt{mod}}  &  cat \\
      cat & \arclabel{\texttt{subj}} &  ate \\
      ate & \arclabel{\texttt{root}} &  $<$root$>$ \\
    \end{tabular} & 
    \begin{tabular}{@{$\langle$}c@{,~}c@{,~}c@{$\rangle$}}
      the &      \arclabel{\texttt{det}}  &  cat \\
      cat &      \arclabel{\texttt{subj}} &  stumbled \\
      stumbled & \arclabel{\texttt{root}} &  $<$root$>$ \\
    \end{tabular}\\
  \end{tabular}\\
  Precision$_{dlh}$ is $\frac{1}{3}$ and Recall$_{dlh}$ is
  $\frac{1}{4}$, yielding  $\DPM{dlh} = F = \frac{2}{7} = 0.286$
  \caption{Example hypothesis and reference dependency trees and the $dlh$ decomposition of each.}
  \label{fig:dpmexample}
\end{figure*}
Figure~\ref{fig:dpmexample} demonstrates a toy example of the
bags-of-dependencies extracted from a hypothesis and reference
tree.  In the examples presented here, we extract only $dlh$
dependencies, which are tuples of the form $\langle \textrm{Dependent},
\textrm{arc-Label}, \textrm{Head} \rangle$.
We describe different variants in this family below, and compare their
effectiveness in section~\ref{sec:facorr}.

\subsection{\DPMempty{} variations in subtree extraction}
Different members of the \DPMempty{} family may extract different
subtrees.  In this section, we denote the set of tree-components extracted via a
trailing subscript: \DPM{dlh} extracts all $\langle
\textrm{Dependent}, \textrm{arc-Label}, \textrm{Head} \rangle$ subtree
tuples for the $F$ measure, which is equivalent (modulo the
dependency-extraction methods) to labeled \textsc{SParseval} and the
\newcite{owczarzak07labelleddepseval} \textbf{d} measure.
\DPM{dl,lh}, by contrast, extracts all the subtrees $\langle
\textrm{Dependent}, \textrm{arc-Label} \rangle$ and $\langle
\textrm{arc-Label}, \textrm{Head} \rangle$, which is equivalent to the
\newcite{owczarzak07labelleddepseval} \textbf{d\_var} method (again,
modulo the dependency-extraction method).  In the example trees in
figure~\ref{fig:dpmexample}, the hypothesis tree produces the following
six items for scoring with \DPM{dl,lh}:
\begin{center}
  \begin{tabular}{cc}
    $dl$ & $lh$ \\
    \hline
    \begin{tabular}{@{$\langle$}c@{,~}c@{$\rangle$}}
      the & \arclabel{\texttt{det}}  \\
      cat & \arclabel{\texttt{subj}} \\
      stumbled & \arclabel{\texttt{root}} \\
    \end{tabular} &
    \begin{tabular}{@{$\langle$}c@{,~}c@{$\rangle$}}
      \arclabel{\texttt{det} } &  cat \\
      \arclabel{\texttt{subj}} &  stumbled \\
      \arclabel{\texttt{root}} &  $<$root$>$ \\
    \end{tabular}\\
  \end{tabular}
\end{center}
so that Precision$_{dl,lh}$ is $\frac{3}{6}$ and Recall$_{dl,lh}$ is
$\frac{3}{8}$, giving a \DPM{dl,lh} of $\frac{3}{7} = 0.429$ for the
example.

\subsection{\DPMempty{} variations in binning subscores}
When the \DPMempty{} measure is composed of more than one class of
tuple (e.g.\ \DPM{dl,lh}, but not \DPM{dlh}), we may consider adopting
a strategy like BLEU \cite{papineni02bleu} or the syntactic measures
from \newcite{liu05syntaxformteval} and compute an average of
subscores over each tuple class. We define a binned DPM \bDPM{} as the
harmonic mean of separate precision and recall
scores\footnote{\bDPM{} differs from both \newcite{papineni02bleu}
  and \newcite{liu05syntaxformteval} in that the earlier metrics use
  averages over precision subscores, necessitating an ``anti-gaming''
  brevity penalty (in BLEU), while the \bDPM{} measure is an average
  across precisions and recalls.} for each class of tuple involved,
rather than simple $F$ (the harmonic mean of a single precision and
recall score over all the tuples together).

\subsection{\DPMempty{} variations using $n$-best lists and expected counts}
%% Multiple parses
Since the dependency structure is hidden, we may also be interested in
exploring alternative dependency structures (in both hypothesis and
reference) predicted by the parser, to cope with both error in the
parser and ambiguity in the translation hypothesis and
reference. \DPMempty{} is well-defined over the $n$-best list of
dependency-structures: when $n>1$, \DPMempty{} uses the expectation of
bags-of-subtrees (rather than the bags-of-subtrees derived from the
1-best parse).

An expectation requires a probability distribution over the $n$-best
list, and we consider three options: uniform, the parser
probabilities, and a flattened version of the parser probabilities
such that $\tilde{p}(x) = \frac{p(x)^\gamma}{\sum_ip(i)^\gamma}$
(where $\gamma$ is a free parameter) to account for the fact that the
parser tends to be over-confident.  In all cases, the probabilities
are normalized to sum to one over the the $n$-best list, where the
maximum $n$ in this work is 50.  The uniform distribution ($\gamma =
0$) is intended to be equivalent to the
\newcite{owczarzak07labelleddepseval} \textbf{d\_50} and
\textbf{d\_50\_var} measures.\footnote{Since
  \newcite{owczarzak07labelleddepseval} report no use of parse
  weights, \textbf{d\_50} and \textbf{d\_50\_var} may be using a sum
  of counts over the 50-best list rather than expected-counts over a
  uniform distribution. These two approaches are equivalent --- so
  long as the $n$-best list is always the same length for hypothesis
  and reference.  In our implementation
  (section~\ref{sec:implementation}), the $n$-best list does not
  always reach 50 candidate parses on short sentences, so the
  expectation matches our intent better than a sum of counts over the
  $n$-best.}
%
We note in
table~\ref{tab:measurecorrespondences} those measures in the
\DPMempty{} family that correspond to
\newcite{owczarzak07labelleddepseval} measures.
\begin{table}
  \centering
  \begin{tabular}{|r|r|r||l|}
    \hline
    \multicolumn{3}{|c||}{\DPM{}} & \newcite{owczarzak07labelleddepseval}\\
    Sub-graph & $n$ & $\gamma$ & \\ % & \textbf{d\_*} measure \\
    \hline
    $dlh$ & 1  & --- & \textbf{d} \\
    $dlh$ & 50 & 0   & \textbf{d\_50} \\
    \hline
    $dl,lh$ & 1 & --- & \textbf{d\_var} \\
    $dl,lh$ & 50 & 0 & \textbf{d\_50\_var}\\
    \hline
  \end{tabular}
  \caption{\textbf{Correspondences between the \DPMempty{} family of
      measures and \newcite{owczarzak07labelleddepseval} \textbf{d\_*} measures.}
    Differences between dependency extraction methods are ignored for
    these equivalencies.}
  \label{tab:measurecorrespondences}
\end{table}
% TO DO: MO note: if tight on space, delete this table.

\subsection{Implementation of the \DPMempty{} family}
\label{sec:implementation}
In principle, the \DPM{} family of measures may be implemented with
any parser that generates a dependency graph (a single labelled arc
for each word, pointing to its head-word). Prior work
\cite{owczarzak07labelleddepseval} on related
measures has used an LFG parser \cite{cahill04lfg} or
an unlabelled dependency tree \cite{liu05syntaxformteval}.

%
In this work, we use a state-of-the-art PCFG (the first stage of
\newcite{charniak-johnson:2005:ACL}) and context-free head-finding
rules \cite{magerman95headfinding} to generate a 50-best list of
dependency trees for each hypothesis and reference translation.  We
use the parser's default Wall Street Journal training parameters.
Head-finding uses the Charniak parser's rules, with three
modifications: prepositional and complementizer phrases choose nominal
and verbal heads respectively (rather than functional heads) and
auxiliary verbs are modifiers of main verbs (rather than the
converse).

Having constructed the dependency tree, we label the arcs as $d
\stackrel{A/B}{\to} h$, where the arc label $A/B$ between dependent
$d$ and its head $h$ is composed of $A$ (the lowest constituent-label
headed by $h$ and dominating $d$) and $B$ (the highest constituent
label headed by $d$). 
\begin{figure}
  % \centering
  \Tree
  [.\headlabel{S}{stumbled}
    [.\headlabel{NP}{cat}
      [.\headlabel{DT}{the} the ]
      [.\headlabel{NN}{cat} cat ]
    ]
    [.\headlabel{VP}{stumbled} 
      [.\headlabel{VBD}{stumbled} stumbled ]
    ]
  ]
  \\
  \begin{center}
    \includegraphics[scale=0.6]{dpm-example-depextract.png}
  \end{center}
  \caption{An example constituent tree (heads of each constituent are
    listed small below the label) and the labelled dependency tree
    derived from it using the strategy described in
    section~\ref{sec:implementation}.}
  \label{fig:depextract}
\end{figure}
For example, in figure~\ref{fig:depextract}, the S node is the
lowest node headed by \emph{stumbled} that dominates \emph{cat}, and
the NP node is the highest constituent label headed by \emph{cat}, so
the arc between \emph{cat} and \emph{stumbled} is labelled \arclabel{S/NP}.
%
This strategy is very similar to one adopted in the reference
implementation of labelled-dependency \textsc{SParseval}, and may be
considered as an approximation of the rich semantics generated by
\cite{cahill04lfg} or another heavily knowledge-engineered
parser, but with much less knowledge-engineering required.

The $A/B$ labels are not as descriptive as the LFG semantics, but they
have a similar resolution, e.g.\ the $\stackrel{S/NP}{\to}$ arc label
usually represents a subject dependent of a sentential verb.

\section{Correlation with human judgments of fluency \& adequacy}
\label{sec:facorr}

To select a good member of the \DPMempty{} family, we explore the
correlation of these measures against a corpus of human judgments of
fluency and adequacy.

\subsection{Corpus}

For these experiments, we use LDC Multiple Translation Chinese corpus
parts 2~\cite{LDC03MTC2} and 4~\cite{LDC06MTC4}.  These corpora
include multiple human judgments of fluency and adequacy for each
sentence, with each judgment using a different human judge and a
different reference translation.  For a rough\footnote{Our segment
  count differs from \newcite{owczarzak07labelleddepseval}, who report
  16,800 segments over the same corpus. We find baseline correlations
  (BLEU$_4$ and TER) lower than those reported there as well, so the
  results presented here are not directly comparable with that paper,
  though we demonstrate similar gains over those baselines in
  essentially the same corpus (section~\ref{sec:facorr:newdeps}).}
comparison with \newcite{owczarzak07labelleddepseval}, we treat each
judgment as a separate segment.
%
This treatment of this corpus yields 16,815 tuples of
$\langle$hypothesis, reference, fluency, adequacy$\rangle$.  In these
experiments, we extend this tuple with automatic scores derived from
$\langle$hypothesis, reference$\rangle$ and examine the
correlations\footnote{The independence of each of these segments is
  questionable, since the same hypothesis translations are used in
  multiple items, but for the sake of methodological comparison with
  prior work, this strategy is preserved.}  between those automatic
scores and the arithmetic mean of the fluency and adequacy measures.
% TO DO: anything more to say here.

\subsection{Utility of alternative dependency extraction}
\label{sec:facorr:newdeps}
Our research suggests that the dependency extraction strategy from
section~\ref{sec:implementation} works in place of the richer
semantics from an LFG parser.  
\begin{table}
  \centering
  \begin{tabular}{r|d}
    \textbf{Measure} & \multicolumn{1}{c}{$r$} \\
    % vs.\ $\frac{\mathrm{fl} + \mathrm{ad}}{2}$} \\
    \hline
    \DPM{dl,lh} ($\sim$\textbf{d\_var}) & 0.226\\
    \DPM{dlh} ($\sim$\textbf{d}) & 0.185 \\
    TER & -0.173 \\
    BLEU$_4$ & 0.139 \\
  \end{tabular}
  \caption{
    Correlation of various measures with the average of fluency and
    adequacy over the sentences in the MTC corpus. These \DPMempty{} results use
    only the one-best parse ($n=1$).
    \label{tab:facorr:statsparser}}
\end{table}
The results in table~\ref{tab:facorr:statsparser}, which use only the
top parse for each sentence, are very similar to those reported in
\newcite{owczarzak07labelleddepseval}, in that the syntactic measure
\DPM{dl,lh} ($\sim$\textbf{d\_var}) is much-better correlated with
human judgment than the non-syntactic measures BLEU$_4$ and TER.
%% TO DO: look at the following thought
% The difference between the syntactic and non-syntactic measures are
% here not quite as large as the differences reported in earlier work,

\subsection{Alternative dependency sub-graphs}

The \DPMempty{} family allows us to easily explore alternative
sub-graphs of the dependency graph, and we find that we achieve small
improvements in correlation with human judgments by including the
unigram ($1g$) and bigram ($2g$) in the dependency-tree decomposition,
ignoring the dependency arc.  \newcite{owczarzak07labelleddepseval}
found that their original proposal, \textbf{d}, scoring full
dependent-arc-head triples, was not as well-correlated with human
judgment as \textbf{d\_var}, which examined only dependent-arc and
arc-head tuples. Table~\ref{tab:facorr:statsparser} confirms this as
well for a statistical constituent parser with simple
dependency-extraction.  In table~\ref{tab:facorr:subgraphs}, we extend
this search to consider whether it is useful to include other
subgraphs of the dependency tree into the bag of tree-fragments to be
scored.
\begin{table}
  \centering
  \begin{tabular}{r|d}
    \textbf{Measure} & \multicolumn{1}{c}{$r$}\\
    % vs.\ $\frac{\mathrm{fl} + \mathrm{ad}}{2}$} \\
    \hline
    \DPM{1g,2g,dl,lh} & 0.237\\
    \DPM{1g,dl,lh}   & 0.234\\
    \\
    \DPM{1g,2g}($\equiv\BoNG{2}$)     & 0.227 \\
    \DPM{dl,lh}     & 0.226\\
    \\
    % \DPM{d} ($\equiv$ bag-of-words F) & 0.214 \\
    \DPM{1g,dl,dlh} & 0.227\\
    \DPM{dlh} & 0.185 \\
    TER       & -0.173 \\
    BLEU$_4$  & 0.139 \\
  \end{tabular}
  \caption{
    As in table~\ref{tab:facorr:statsparser}, but with 
    alternative dependency-graph constituents to compute the $F$
    measure.  Again, $n=1$ for all \DPMempty{} correlations.
    \label{tab:facorr:subgraphs}}
\end{table}

Table~\ref{tab:facorr:subgraphs} shows that we can combine the
benefits of string-local $n$-grams (\DPM{1g,2g}) with the benefits of
dependency information (\DPM{dl,lh}) for a further improved
correlation with human judgment, with the best correlation in
\DPM{1g,2g,dl,lh}. Including progressively larger chunks of the
dependency graph (as in \DPM{1g,dl,dlh}, which is inspired by the
BLEU$_k$ idea of progressively larger $n$-grams) does not seem to be
an improvement over \DPM{dl,lh}.

\subsection{Binning subscores with \bDPM{}}
BLEU \cite{papineni02bleu} and the related NIST
\cite{doddington02nist} measure, as well as the earlier proposed
syntactic measures \cite{liu05syntaxformteval}, choose to sort their
components into bins on length, and take average scores over those
bins.  \newcite{owczarzak07labelleddepseval} measures, by contrast, and
the \DPM{} measures by extension, score all components together in one
bin, measuring precision and recall of all bins combined.
\begin{table}
  \centering
  \begin{tabular}{r|d}
    \textbf{\DPM{} measures} & \multicolumn{1}{c}{$r$} \\
    % vs.\  $\frac{\mathrm{fl} + \mathrm{ad}}{2}$} \\
    \hline
    \DPM{1g,2g,dl,lh}  & 0.237\\
    \bDPM{1g,2g,dl,lh} & 0.217 \\
%     \\
    \DPM{1g,dl,lh}  & 0.227\\
    \bDPM{1g,dl,lh} & 0.212 \\
    \DPM{dl,lh} & 0.226\\
    \bDPM{dl,lh} & 0.208 \\
%     \hline
%     \DPM{d} & 0.214 \\
    \hline
    \hline
    \textbf{$n$-gram measures} \\
    \hline
    \BoNG{2} & 0.227 \\
    \bBoNG{2} & 0.215\\
    BLEU$_2$ & 0.211 \\
    \hline
    \BoNG{3} & 0.227 \\
    BLEU$_3$ & 0.179 \\
    \bBoNG{3} & 0.177\\
    \hline
    \BoNG{4} & 0.225 \\
    BLEU$_4$ & 0.139 \\
    \bBoNG{4} & 0.135 \\
    \hline
%     TER & -0.173 \\
  \end{tabular}
  \caption{
    As in
    table~\ref{tab:facorr:subgraphs}, but also exploring the
    possibility of binning syntactic and $n$-gram components of
    different sizes into  subscores and combining, on the model of
    BLEU.  $n=1$ for all \DPM{} measures.
    \label{tab:facorr:multibin}
  }
\end{table}  
In table~\ref{tab:facorr:multibin} we explore which strategy is
preferred, and show that binning with \bDPM{} does not improve
correlations for the syntactic measures.  In fact, \bDPM{} measures
are consistently worse-correlated than their corresponding \DPMempty{}
measure.

For comparison, we try the same approach with $n$-grams, comparing
\bBoNG{k} (the
harmonic mean of $k$ precisions and $k$ recalls) to \BoNG{k} (a
single $F$ over one bin of $n$-grams of lengths 1-$k$), and
we find that the single bin again performs better than the separated
bins.  For comparison, the \BoNG{1} (a bag-of-words $F$) has an $r$ of
0.214 over the same corpus; thus, for $n$-grams, the main improvement
is in moving from 1-grams to 2-grams, but only for the non-binned
variants.  We also include BLEU in table~\ref{tab:facorr:multibin},
which allows a comparison of BLEU's strategy (average precision and
brevity penalty) to an $F$-style harmonic mean of precision and recall
(binned separately and together).  We find that BLEU$_k$ does about as
well as the corresponding \bBoNG{k}, while the \BoNG{k} does
consistently better than BLEU$_k$, especially for larger $k$.
% TO DO: Worth noting that \bDPM{dl,lh} is actually worse than \DPM{d}?

% note: \DPM{d,dh} actually performs \emph{worse} (0.224) than
%   \BoNG{2} (0.227).  Looking at alternative parses doesn't help really
%   at all, either.  Both of these suggest that arc labels --- even more
%   than arc-positions --- make a difference here.   Is this
%   worth discussing? Exploring POS-tagging matches?

\subsection{Using parse $n$-best lists}
We explore the use of multiple parses in
table~\ref{tab:facorr:multiparse}, which presents \DPM{} variants with
$\gamma=0$ of the most successful \DPM{} sub-graph lists shown in previous
tables.
\begin{table}
  \centering
  \begin{tabular}{r|r|d}
    \textbf{Measure} & \textbf{parameters} & \multicolumn{1}{c}{$r$}\\
    % vs.\ $\frac{\mathrm{fl} + \mathrm{ad}}{2}$} \\
    \hline
    \DPM{1g,2g,dl,lh} & $\gamma=0, n={50}$ & 0.239 \\
    \DPM{1g,2g,dl,lh} &  $n=1$ & 0.237\\
    \DPM{1g,dl,lh} & $\gamma=0, n={50}$ & 0.237 \\
    \DPM{1g,dl,lh} &  $n=1$ & 0.234\\
%    \\
    \DPM{dl,lh} ($\sim\textbf{d\_50\_var}$) & $\gamma=0, n={50}$  & 0.234 \\
    \DPM{dl,lh} ($\sim\textbf{d\_var}$)& $n=1$ & 0.226\\
  \end{tabular}
  \caption{
    As in table~\ref{tab:facorr:subgraphs}, but considering variants
    of the best \DPM{} measures uniform probability distribution over
    multiple parses ($\gamma=0, n=50$).
    \label{tab:facorr:multiparse}}
\end{table}
We use $\gamma=0$ (uniform probability over the $n$-best list) to
compare as closely as possible to
\newcite{owczarzak07labelleddepseval}, which uses a parser with ranks
but no weights.

We find that using multiple parses with a uniform distribution
improves correlations further, although the improvement from varying
the dependency-tree sub-graphs is not as large for $n=50$ variants of
\DPM{} as for $n=1$ variants.
% \footnote{Although the \BoNG{3} correlations are
%   impressive, they cannot take advantage of the extra knowledge source
%   provided by the parser, since they only use surface features.}

%% Not sure how true this next thought is:
% Also note that the improvements are not as large
%   with this implementation as the improvements reported in
%   \cite{owczarzak07labelleddepseval}.  \footnote{possible that the
%     alternative LFG dependencies are more semantially-diverse/less
%     subject to (e.g.) morphological scatter, so $n$-best is more
%     diverse in LFG?  a note for future work?}

\subsection{Including parse confidence}
Since the parser in our implementation provides a confidence in each
parse, we explore the use of that confidence with the $\gamma$ free
parameter.
\begin{table}
  \centering
  \begin{tabular}{r|l||d}
    \multicolumn{3}{c}{\DPM{1g,2g,dl,lh}}  \\
    \multicolumn{1}{c|}{$n$} & \multicolumn{1}{c||}{$\gamma$} & \multicolumn{1}{c}{$r$} \\
    \hline
    % vs.\ $\frac{\mathrm{fl} + \mathrm{ad}}{2}$} \\
    \hline
    % TO DO: fill these in.
    50 & {0.25} & 0.240 \\
    50 & {0.5\phantom{0}}  & 0.240 \\
    50 & {0.75} & 0.240 \\
    50 & {1}    & 0.239 \\
    50 & 0      & 0.239 \\
     1 & ---    & 0.237 \\
    \hline
  \end{tabular}~
  \begin{tabular}{r|l||d}
    \multicolumn{3}{c}{\DPM{dl,lh}}   \\
    \multicolumn{1}{c|}{$n$} & \multicolumn{1}{c||}{$\gamma$} & \multicolumn{1}{c}{$r$} \\
    \hline
    % \end{tabular}\\
    % \begin{tabular}{r|d}
    %   \textbf{Measure} & \multicolumn{1}{c}{$r$ vs.\
    %     $\frac{\mathrm{fl} + \mathrm{ad}}{2}$} \\
    \hline
    50 & {0.25} & 0.234 \\
    50 & {0.5} & 0.234 \\
    50 & 0 & 0.234 \\
    50 & 0.75 & 0.233 \\
    50 & {1} & 0.232 \\
    1 & --- & 0.226 \\
    \hline
%     \hline
%     \textbf{Comparison measures}\\
%     \hline
%     (\BoNG{3}) & (0.227) \\
%     (\BoNG{2}) & (0.227) \\
%     (TER) & (-0.173) \\
%     (BLEU$_4$) & (0.139) \\
  \end{tabular}
  \caption{
    As in table~\ref{tab:facorr:multiparse}, but
    considering various values of $\gamma$ and $n$ for two different
    \DPMempty{} sub-graph lists ($dl,lh$ and $1g,2g,dl,lh$).
%   All variants of EDPM use
%     the 50-best parses.
    \label{tab:facorr:parseweight}}
\end{table}
Table~\ref{tab:facorr:parseweight} explores various ``flattenings''
(values of $\gamma$) of the parse confidence in the \DPM{}
measure. $\gamma=1$ is not always the best, suggesting that the parse
probabilities $p(\mathrm{tree}|\mathrm{words})$ are overconfident. We
find that $\gamma=0.25$ is generally the best flattening of the parse
confidence for the variants of \DPM{} that we have tested.  The
differences are small, but the trends are consistent across the
variants.

\subsection{Summary}

In this section, we have presented experiments exploring a number of
parameters to the \DPMempty{} measure. The experiments suggest a
best-case variant \myEDPM{}, where we set:
\begin{displaymath}
  \myEDPM{} = \DPM{1g,2g,dl,lh}, n=50, \gamma=0.25
\end{displaymath}
in which we choose a $1g,2g,dl,lh$ sub-graph decomposition based on
the improvements from better sub-graphs
(table~\ref{tab:facorr:subgraphs}), multiple parses ($n=50$) based on
table~\ref{tab:facorr:multiparse}, and $\gamma=0.25$, hinted at by
table~\ref{tab:facorr:parseweight}.  We use these \myEDPM{}
parameter-settings in the experiments exploring correlations with HTER
(below).

% TO DO: MO says: I really don't like the final result tied to a
% specific gamma.  the best gamma is a parameter that would be tuned
% for particular tasks and parsers in the same what that one might
% tune n for n-best, and you don't have 50 in the name.

% MO: You could avoid the hat by not using EDPM until this section.
% JGK: TO DO: don't use EDPM until here.

\section{Correlations with HTER}
\label{sec:hter}

Having chosen \myEDPM{} as a good candidate member of the \DPMempty{}
family, we explore its utility on another task: predicting the
human-targeted translation edit rate (HTER) on the (unsequestered) GALE 2.5
% \cite{olive06gale}
evaluation results.

\subsection{Corpus}

The GALE 2.5 translation corpus is made up of system translations into
English from three sites.
% MO added:
The three sites all use system combination to integrate results from
multiple systems, some of which are phrase-based and some which may
use syntax on either the source or target side. No system provided
system-generated parses.
%
The corpus being translated comes from
Arabic and Chinese in four genres: \texttt{bc} (broadcast
conversation), \texttt{bn} (broadcast news), \texttt{nw} (newswire),
and \texttt{wb} (web text), with corpus sizes shown in
table~\ref{tab:galestats}.
\begin{table}
  \centering
  \begin{tabular}{r|rr|rr|rr}
     & \multicolumn{2}{c|}{Arabic} & \multicolumn{2}{c|}{Chinese}
     & \multicolumn{2}{c}{Total}\\
     & doc & sent & doc & sent & doc   & sent\\
     \hline
     \texttt{bc}    & 59  & 750 & 56 & 1061 & 115 & 1811\\
     \texttt{bn}    & 63  & 666 & 63 & 620  & 126 & 1286\\
     \texttt{nw}    & 68  & 494 & 70 & 440  & 138 & 934 \\
     \texttt{wb}    & 69  & 683 & 68 & 588  & 137 & 1271\\
     \hline
     Total & 259 & 2593& 257& 2709 & 516 & 5302\\
  \end{tabular}
  \caption{Corpus statistics for the GALE 2.5 translation
    corpus.}
  \label{tab:galestats}
\end{table}
The corpus includes one English reference translation $r_i$
\cite{gale08phase2_5references} for each sentence $i$ and a system
translation $t_{i,z}$ for each of the three systems
$z$. Additionally, each of the system translations of each segment $i$
has a corresponding human-targeted reference aligned at the sentence
level, so we have available the HTER score of each segment
$s_{\textsc{hter}}(t_{i,z})$ at both the sentence and document level.

\subsection{Prediction of $\Delta$HTER per-document}

We evaluate our new measure \myEDPM{} by testing its prediction of
improvements in HTER over different translations of a document.
We name one of the three translation systems arbitrarily as baseline
$b$ and define
\begin{eqnarray}
  \label{eq:delta}
  \Delta s(i,b,z) & = & s(t_{i,z}) - s(t_{i,b})
\end{eqnarray}
as a measure of change % in measure $s$
that system $z$ provides over segment $i$ with respect to system $b$
(by providing translation $t_{i,z}$ instead of $t_{i,b}$).
% $\Delta_b(s,t_{i,z})$ may be either positive or negative.

We would like an automatic measure $s$ that provides a good
correlation $r$ between $\Delta s$ and $\Delta \textsc{hter}$.  For
each of the 516 documents\footnote{Using per-document comparison
  avoids the problems of per-sentence comparisons, e.g.\ that BLEU
  falls to zero too easily on a sentence.} in the corpus described in
table~\ref{tab:galestats}, for each of the three pairs of systems, we
generate the $\Delta s$ scores, ordering the systems such that
$\Delta\textsc{hter} \geq 0$ for each document pair.

% For candidate score
% measures $s$ (including target measure \textsc{hter}), we generate
% three $\Delta s$ data (all possible system pairs, ordered such that
% $\Delta$\textsc{hter} is always positive) for
% % Using this $\Delta_b$ measure for each score, we have two data for
% each document of the 516 documents in the corpus
% described in table~\ref{tab:galestats}.

Table~\ref{tab:hter:perdoc1} shows the per-document correlations with
$\Delta\textsc{hter}$ over this set of data, broken out into per-genre
and per-language correlations.

\begin{table}
  \centering
  \begin{tabular}{r|r|r||r}
     \multicolumn{1}{c|}{Measure $s$} 
     & \multicolumn{1}{c|}{all-Arabic} &
     \multicolumn{1}{c||}{all-Chinese} &\multicolumn{1}{c}{all}\\
    \hline{}
    TER      &  0.51 &  0.19 &  0.39 \\
    BLEU$_4$ & -0.40 & -0.19 & -0.32 \\
    \myEDPM  & 
       \textbf{-0.61}&
               \textbf{-0.25}&
                       \textbf{-0.47}\\
  \end{tabular}
  \caption{Per-doc $r$ of $\Delta s$ with $\Delta\textsc{hter}$ over
    various measures $s$, examined for each genre in the corpus, for each
    language in the corpus, and as a whole. }
  \label{tab:hter:perdoc1}
\end{table}
We compare \myEDPM{}'s correlation with HTER against both that of TER
and that of BLEU, two very popular automatic measures. It is worth
mentioning that TER has an 
%a not-so-hidden 
advantage in that HTER uses a TER
measure to calculate the post-editing work between the hypothesized
translation and the human-targeted reference which could, in
principle, bias HTER towards a TER measure. \myEDPM{} shares no such
advantage.  \myEDPM{} nevertheless has the best
correlation of the three measures in both Arabic and
Chinese, as well as over the entire corpus.

\begin{table*}
  \centering
  \begin{tabular}{r|rrrr|}
    & \multicolumn{4}{c|}{Arabic} \\
    \multicolumn{1}{c|}{Measure $s$} & \multicolumn{1}{c}{\texttt{bc}} &
    \multicolumn{1}{c}{\texttt{bn}} & \multicolumn{1}{c}{\texttt{nw}} & \multicolumn{1}{c|}{\texttt{wb}}\\
    \hline{}
    %          bc     &  bn    & nw     & wb \\
    TER      &  0.59  & 
                 \textbf{0.24} &  0.22  &  0.26 \\
    BLEU$_4$ & -0.50  & -0.10  &
                         \textbf{-0.30} & -0.31 \\
    \myEDPM  &
       \textbf{-0.80 }& -0.10 &
                         \textbf{-0.31} &
                                  \textbf{-0.33} \\
  \end{tabular}
  \begin{tabular}{|rrrr|}
    %& 
    \multicolumn{4}{|c|}{Chinese} \\
        % Measure & 
        \multicolumn{1}{|c}{\texttt{bc}} &
    \multicolumn{1}{c}{\texttt{bn}} & \multicolumn{1}{c}{\texttt{nw}} & \multicolumn{1}{c|}{\texttt{wb}}\\
    \hline{}
    %  
    % bc       &  bn  & nw    & wb \\
%     TER      &
     0.06     &  0.13 &  0.35 &  0.14 \\
    %BLEU$_4$ &
    -0.01     & -0.22 &
                \textbf{-0.36}& -0.07 \\
%   \myEDPM   &
\textbf{-0.14}&
        \textbf{-0.30}&
                \textbf{-0.37}&
                        \textbf{-0.16}\\
  \end{tabular}
  \caption{Per-document $r$ of $\Delta s$ with $\Delta\textsc{hter}$ for various
    measures $s$. Correlations by individual 
    language$\times$genre subcorpora.
  }
  \label{tab:hter:perdoc2}
\end{table*}
In table~\ref{tab:hter:perdoc2}, we break out the results into
individual language$\times$genre subcorpora for further comparison. We
find that \myEDPM{} is the best measure in nearly all subcorpora
(Arabic \texttt{bn} standing as an exception).

\subsection{Prediction of $\Delta\textsc{hter}$ per-sentence}
Like the per-document correlations, at the sentence level we are
interested in identifying the predictive power of changes in score $s$
with respect to changes in the gold-standard score (here,
\textsc{HTER}).  In table~\ref{tab:hter:persent2}, we present
per-sentence correlations of $\Delta\textsc{hter}$ improvements
weighted by sentence length\footnote{When calcluating per-sentence
  correlations, we want to scale differences in $\Delta s(i,b,z)$ by
  the length of reference $r_i$, since an improvement of $x$\% on
  score $s$ in a long sentence is ordinarily understood to be worth
  more than the same $x$\% on a short sentence.}, also broken out by
individual language$\times$genre subcorpora.

\begin{table*}
  \centering
  \begin{tabular}{r|rrrr|}
    & \multicolumn{4}{c|}{Arabic} \\
    Measure $s$& \texttt{bc} & \texttt{bn} & \texttt{nw} & \texttt{wb} \\
    \hline{}
    TER        &  0.54 & 
                  \textbf{0.18}&  0.11 &  0.19 \\
    BLEU$_4$   & -0.29 & -0.15 & -0.06 & -0.17 \\
    \myEDPM &
         \textbf{-0.59}& -0.12 &
                         \textbf{-0.15}&
                                 \textbf{-0.21} \\
  \end{tabular}
  \begin{tabular}{|rrrr|}
    \multicolumn{4}{|c|}{Chinese} \\
    \texttt{bc} & \texttt{bn} & \texttt{nw} & \texttt{wb} \\
    % TER
    \hline{}
    0.15   &  0.14  &
              \textbf{0.26}& 
                      \textbf{0.13} \\
    % BLEU$_4$
    -0.06  & -0.10  &-0.22 &
                      \textbf{-0.13}\\
    % \myEDPM
    \textbf{-0.18}&
      \textbf{-0.17}&
             \textbf{-0.27}&
                      \textbf{-0.13} \\
  \end{tabular}
  \caption{Per-sentence $r$ of length-weighted $\Delta s$ with
    length-weighted $\Delta\textsc{hter}$.  Correlations by individual 
    language$\times$genre subcorpora.}
  \label{tab:hter:persent2}
\end{table*}
Table~\ref{tab:hter:persent2}'s per-sentence results are largely
similar to the per-document analysis (table~\ref{tab:hter:perdoc2}),
although both the absolute and relative differences are smaller.

\section{Conclusion and future work}
\label{sec:conclusion}
% Leave blank space for acknowledgements, since you can't put them in
% now for anonymizing reasons, but we need to later.

% TO DO: If needed for saving space, you can take the location out of
% the conference paper refs.  Or use first initials, or both.

In this paper, we described \DPM{}, a family of metrics for evaluating
machine translation quality using a labelled dependency tree.  Using
the Multiple Translation Chinese corpus, we selected a member of that
family \myEDPM{} and found a good value for free parameter $\gamma$,
taking advantage of parse structure and parser probabilities.

We then tested the \myEDPM{} measure against the GALE 2.5 translation
corpus and evaluated its ability to predict $\Delta\textsc{hter}$
(the change in \textsc{hter}) on a per-document and per-sentence
level.  We have shown \myEDPM{} to be superior to both BLEU$_4$
and TER in most cases and for most of the subcorpora available to us.

\myEDPM{} has the same advantages as the
\newcite{owczarzak07labelleddepseval} \textbf{d\_*} measures, but is
implemented in a more portable way.  Both methods require
substantially more time to run than BLEU or TER, because parses of
both the hypothesis and the reference are required, but the method
described in this paper may be implemented with any PCFG-based parser
--- a treebanked corpus for training that parser should be sufficient.

% In the second half of the work presented here, we have seen that both
% $\Delta\myEDPM{}$ and $\Delta$TER are relatively well-correlated with
% $\Delta\textsc{hter}$ and that $\Delta$BLEU is almost always
% worse-correlated than the other two. 
The correlations with BLEU and TER are both worth a closer examination.
%
Many of the systems used
in the comparison were optimized on BLEU themselves, which raises the
possibility that current MT systems are over-fitted to BLEU, reducing
BLEU's utility as a predictor of quality translation. Conversely,
TER's relationship to HTER makes one suspicious about its good
correlation; it might be worth testing against another human-directed
measure as well.

In future work, we would like to explore a number of related
questions.  How useful is it to increase $n$ beyond 50 for this
approach?  We could, for example, use a different PCFG parser as
another way of exploring the trade-offs between parse-quality, MT
quality prediction, and speed.  On the other hand, a poor-quality
translation that is very difficult to parse may interfere with the
quality of the measure; assessing this measure's sensitivity to
sentence quality would also be worthwhile. In a different direction,
we would like to ask whether other segmentations of the dependency
tree are more appropriate than those explored here, following up on an
approach suggested in \newcite{liu05syntaxformteval}, which uses
linked words in chains much larger than 2.

From the results correlating \myEDPM{} with both \textsc{hter} and
human judgments of fluency and accuracy, we see that \myEDPM{} seems
to be a superior tool for identifying improvements at the document
level, and a competitive one at the sentence level, which is often
where parameter tuning takes place. One possible use for this measure,
since it is more computationally-costly than BLEU or TER, might be as a late
pass evaluation metric in training to select among translation outputs
already deemed to be very good.

\section*{Acknowledgments}
\blindthis{
Grant numbers etc.\ go here.
}

% MO: TO DO Could save space by using first initials.
\bibliographystyle{acl}
\bibliography{edpm-paper}

\end{document}

### 2009-05-08 edpm-paper.tex
monkeying with original edpm-paper to use mean-subtracted

@@
   Dependency Pair Match'' (\myEDPM), and 
-%   explore \myEDPM's quality
-%   with respect to the GALE % \cite{olive06gale}
-%   2.5 evaluations corpus.  We 
-  demonstrate that $\Delta$\myEDPM{} is superior to $\Delta$BLEU and
-  $\Delta$TER as a per-document and per-sentence predictor of
-  $\Delta$HTER.
-% , and is superior to $\Delta$TER in
-%   the written contexts (newswire and web text).
+  demonstrate that mean-subtracted \myEDPM{} is superior to
+  mean-subtracted BLEU and
+  mean-subtracted TER as a per-document and per-sentence predictor of
+  mean-subtracted HTER.
 \end{abstract}
@@
 
-\subsection{Prediction of $\Delta$HTER per-document}
+\subsection{Prediction of mean-subtracted HTER per-document}
 
@@
 improvements in HTER over different translations of a document.
-We name one of the three translation systems arbitrarily as baseline
-$b$ and define
-\begin{eqnarray}
-  \label{eq:delta}
-  \Delta s(i,b,z) & = & s(t_{i,z}) - s(t_{i,b})
-\end{eqnarray}
-as a measure of change % in measure $s$
-that system $z$ provides over segment $i$ with respect to system $b$
-(by providing translation $t_{i,z}$ instead of $t_{i,b}$).
-% $\Delta_b(s,t_{i,z})$ may be either positive or negative.
+Rather than consider absolute scores $s$ per-document, we normalize for
+the underlying difficulty of the document by examining $\mu s(t)$, the difference
+of translation $t_i$ from the mean score over the $I$ translations of
+the same document, where 
+\begin{equation}
+  \label{eq:meansub}
+  \mu s(t_i) = s(t_i) - \sum_{j=1}^I\frac{s(t_j)}{I}
+\end{equation}
 
 We would like an automatic measure $s$ that provides a good
-correlation $r$ between $\Delta s$ and $\Delta \textsc{hter}$.  For
+correlation $r$ between $\mu s$ and $\mu \textsc{hter}$.  For
 each of the 516 documents\footnote{Using per-document comparison
@@
 table~\ref{tab:galestats}, for each of the three pairs of systems, we
-generate the $\Delta s$ scores, ordering the systems such that
-$\Delta\textsc{hter} \geq 0$ for each document pair.
+generate the $\mu s$ scores, calculating a mean for ea.
 

### 2009-05-08 kluwer/klusampl.tex
kluwer subdir added and mtjournal.tex initialized

\documentclass{kluwer}    % Specifies the document style.

\newdisplay{guess}{Conjecture}

\begin{document}                                                                                   
\begin{article}
\begin{opening}         
\title{A Sample Document\thanks{Footnote 
            to the title with the `thanks' command.}} 
\author{Leslie \surname{Lamport}}  
\runningauthor{Leslie Lamport}
\runningtitle{A Sample Document}
\institute{Author's Affiliation}
\date{April 15, 1993}

\begin{abstract}
This is a sample input file.  Comparing it with the output it
generates can show you how to produce a simple document of
your own.
\end{abstract}
\keywords{sample, \LaTeX}

\end{opening}

\section{Ordinary Text}  
                    % Produces section heading.  Lower-level
                    % sections are begun with similar 
                    % \subsection and \subsubsection commands.

The ends  of words and sentences are marked 
  by   spaces. It  doesn't matter how many 
spaces    you type; one is as good as 100.  The
end of   a line counts as a space.

One   or more   blank lines denote the  end 
of  a paragraph.

Since any number of consecutive spaces are treated like a single
one, the formatting of the input file makes no difference to
      \TeX,         % The \TeX command generates the TeX logo.
but it makes a difference to you.  
When you use
      \LaTeX,       % The \LaTeX command generates the LaTeX logo.
making your input file as easy to read as possible
will be a great help as you write your document and when you
change it.  This sample file shows how you can add comments to
your own input file.

Because printing is different from typewriting, there are a 
number of things that you have to do differently when preparing 
an input file than if you were just typing the document directly.  
Quotation marks like 
       ``this'' 
have to be handled specially, as do quotes within quotes: 
       ``\,`this'                  % \, separates the double and single quote.
    is what I just 
    wrote, not  `that'\,''.

Dashes come in three sizes: an 
       intra-word 
dash, a medium dash for number ranges like 
       1--2, 
and a punctuation 
       dash---like 
this.

A sentence-ending space should be larger than the space between words
within a sentence.  You sometimes have to type special commands in
conjunction with punctuation characters to get this right, as in the
following sentence.
       Gnats, gnus, etc.\    % `\ ' makes an inter-word space.
       all begin with G\@.   % \@ marks end-of-sentence punctuation.
You should check the spaces after periods when reading your output to
make sure you haven't forgotten any special cases.
Generating an ellipsis 
       \ldots\    % `\ ' needed because TeX ignores spaces after 
          % command names like \ldots made from \ + letters.
          %
          % Note how a `%' character causes TeX to ignore the 
          % end of the input line, so these blank lines do not
          % start a new paragraph.
with the right spacing around the periods 
requires a special  command.

\TeX\ interprets some common characters as commands, so you must type
special commands to generate them.  These characters include the
following: 
       \$ \& \% \# \{ and~\}.

In printing, text is emphasized by using an
       {\em italic\/}  % The \/ command produces the tiny extra space that
               % should be added between a slanted and a following
               % unslanted letter.
type style.

\begin{em}
   A long segment of text can also be emphasized in this way.  Text within
   such a segment given additional emphasis 
      with\/ {\em Roman} 
   type.  Italic type loses its ability to emphasize and become simply
   distracting when used excessively.  
\end{em}

It is sometimes necessary to prevent \TeX\ from breaking a line where
it might otherwise do so.  This may be at a space, as between the
``Mr.'' and ``Jones'' in
       ``Mr.~Jones'',        % ~ produces an unbreakable interword space.
or within a word---especially when the word is a symbol like
       \mbox{\em itemnum\/} 
that makes little sense when hyphenated across 
       lines.

\TeX\ is good at typesetting mathematical formulas like
       \( x-3y = 7 \) 
or
       \( a_{1} > x^{2n} / y^{2n} > x' \).
Remember that a letter like
       $x$        % $ ... $  and  \( ... \)  are equivalent
is a formula when it denotes a mathematical symbol, and should
be treated as one.

\section{Notes}
Footnotes\footnote{This is an example of a footnote.}
pose no problem\footnote{And another one}.
Endnotes\endnote{This is an example of an endnote.} pose no 
problem either\endnote{And another one.}.

When using the {\sc kluwer} style file, you can produce endnotes 
analogous to footnotes. Instead of the \LaTeX\ command \verb+\footnote{}+,
you can use \verb+\endnote{}+. The command \verb+\theendnotes+ can be used
to place the endnotes in the text. They will be put in a separate section in
the \verb+\footnotesize+ font.

\section{Displayed Text}

Text is displayed by indenting it from the left margin.
Quotations are commonly displayed.  There are short quotations
\begin{quote}
   This is a short a quotation.  It consists of a 
   single paragraph of text.  There is no paragraph
   indentation.
\end{quote}
and longer ones.
\begin{quotation}
   This is a longer quotation.  It consists of two paragraphs
   of text.  The beginning of each paragraph is indicated
   by an extra indentation.

   This is the second paragraph of the quotation.  It is just
   as dull as the first paragraph.
\end{quotation}
Another frequently-displayed structure is a list.
The following is an example of an {\em itemized} list, four levels deep.
\begin{itemize}
\item  This is the first item of an itemized list.  Each item 
      in the list is marked with a ``tick''.  The document
      style determines what kind of tick mark is used.
\item  This is the second item of the list.  It contains another
      list nested inside it.  The three inner lists are an {\em itemized}
      list.
    \begin{itemize}
       \item This is the first item of an enumerated list that
            is nested within the itemized list.
          \item This is the second item of the inner list.  \LaTeX\
            allows you to nest lists deeper than you really should.
      \end{itemize}
      This is the rest of the second item of the outer list.  It
      is no more interesting than any other part of the item.
   \item  This is the third item of the list.
\end{itemize}

The following is an example of an {\em enumerated} list, four levels deep.
\begin{enumerate}
\item  This is the first item of an enumerated list.  Each item 
      in the list is marked with a ``tick''.  The document
      style determines what kind of tick mark is used.
\item  This is the second item of the list.  It contains another
      list nested inside it.  The three inner lists are an {\em enumerated}
      list.
    \begin{enumerate}
       \item This is the first item of an enumerated list that
            is nested within the enumerated list.
          \item This is the second item of the inner list.  \LaTeX\
            allows you to nest lists deeper than you really should.
      \end{enumerate}
      This is the rest of the second item of the outer list.  It
      is no more interesting than any other part of the item.
   \item  This is the third item of the list.
\end{enumerate}

The following is an example of a {\em description} list.
\begin{description}
\item[Cow] Highly intelligent animal that can produce milk out of grass.
\item[Horse] Stupid animal that has legs.
\item[Human being] Stupid animal that can talk and think.
\end{description}

You can even display poetry.
\begin{verse}
   There is an environment for verse \\    % The \\ command separates lines
   Whose features some poets will curse.   % within a stanza.

               % One or more blank lines separate stanzas.

   For instead of making\\
   Them do {\em all\/} line breaking, \\
   It allows them to put too many words on a line when they'd 
   rather be forced to be terse.
\end{verse}

Mathematical formulas may also be displayed.  A displayed formula is
one-line long; multiline formulas require special formatting
instructions.
   \[  x' + y^{2} = z_{i}^{2}\]
Don't start a paragraph with a displayed equation, nor make
one a paragraph by itself.

Example of a theorem:

\begin{guess}
All conjectures are interesting, but some conjectures are more
interesting than others.
\end{guess}

\section{Tables and Figures}
Cross reference to labelled table: As you can see in Table~\ref{sphericcase} on 
page~\pageref{sphericcase} and also in Table~\ref{parset} on page~\pageref{parset}.

\begin{table} % 
\begin{tabular}{lrl}                                        
\hline
$Q_{s,\max}$   & [g/g DM h]  & 0.18\\
$K_{s}$       & [g/L]        & 1.0\\
$Y_{x/s}$     & [g DM/g]     & 0.5\\
$Y_{p/s}$     & [g/g]        & 0.854\\
$Q_{p,\max}$   & [g/g DM h]  & 0.0045\\
$\mu_{\rm crit}$  & [h$^{-1}$]  & 0.01\\
$k_{h}$       & [h$^{-1}$]  & 0.002\\
$m_{s}$       & [g/g DM h]  & 0.025\\                              
\hline
\end{tabular}
\caption[]{Parameter set used in the model of \shortcite{Bunt}}\label{parset}
\end{table}

\begin{table*}
\caption[]{The spherical case ($I_1=0$, $I_2=0$).}
\label{sphericcase}
\begin{tabular}{crrrrc}
\hline
Equil. \\
Points & \multicolumn{1}{c}{$x$} & \multicolumn{1}{c}{$y$} & \multicolumn{1}{c}{$z$} & \multicolumn{1}{c}{$C$} &
S \\
\hline
$~~L_1$ & $-$2.485252241 & 0.000000000 & 0.017100631 & 8.230711648 & U \\
$~~L_2$ &    0.000000000 & 0.000000000 & 3.068883732 & 0.000000000 & S \\
$~~L_3$ &    0.009869059 & 0.000000000 & 4.756386544 & $-$0.000057922 & U \\
$~~L_4$ &    0.210589855 & 0.000000000 & $-$0.007021459 & 9.440510897 & U \\
$~~L_5$ &    0.455926604 & 0.000000000 & $-$0.212446624 & 7.586126667 & U \\
$~~L_6$ &    0.667031314 & 0.000000000 & 0.529879957 & 3.497660052 & U \\
$~~L_7$ &    2.164386674 & 0.000000000 & $-$0.169308438 & 6.866562449 & U \\
$~~L_8$ &    0.560414471 & 0.421735658 & $-$0.093667445 & 9.241525367 & U \\
$~~L_9$ &    0.560414471 & $-$0.421735658 & $-$0.093667445 & 9.241525367 & U
\\
$~~L_{10}$ & 1.472523232 & 1.393484549 & $-$0.083801333 & 6.733436505 & U \\
$~~L_{11}$ & 1.472523232 & $-$1.393484549 & $-$0.083801333 & 6.733436505 & U
\\ \hline
\end{tabular}
\end{table*}

A major
point of difference lies in the value of the specific production rate $\pi$ for
large values of the specific growth rate $\mu$.
Already in the early publications \cite{Falzon87}
it appeared that high glucose
concentrations in the production phase are well correlated with a
low penicillin yield (the 
`glucose effect'). It has been confirmed recently 
\cite{Bunt,Cahour-thesis,BrownAndBurton,Carr-Goldstein}
that
high glucose concentrations inhibit the synthesis of the enzymes of the
penicillin pathway, but not the actual penicillin biosynthesis.
In other words, glucose represses (and not inhibits) the penicillin
biosynthesis.

These findings do not contradict the results of 
\citeauthor{Chin88-book} (on which \shortcite{Bunt} based their 
production kinetics) and of \citeyear{ChinThesis} which were obtained for
continuous culture fermentations.                
Because for high values of the specific
growth rate $\mu$ it is most likely (as shall be discussed below) that 
maintenance metabolism occurs, it can be shown that
in steady state continuous culture conditions, and with $\mu$ described by a Monod kinetics
\begin{equation}
    C_{s}  =  K_{M} \frac{\mu/\mu_{x}}{1-\mu/\mu_{x}} \label{cs}
\end{equation}
Pirt \& Rhigelato determined $\pi$ for $\mu$ between 
$0.023$ and $0.086$ h$^{-1}$.
They also reported a value $\mu_{x} \approx 0.095$
h$^{-1}$, so that for their experiments $\mu/\mu_{x}$ is in the range 
of $0.24$ to $0.9$. 
Substituting $K _M$ in Eq. (\ref{cs}) by 
the value $K_{M}=1$ g/L as used by \cite{Bunt}, one finds
with the above equation $0.3 < C_{s} < 9$ g/L. This agrees well with 
the work of  \shortcite{Bunt}, who reported that penicillin biosynthesis 
repression only occurs at glucose concentrations from $C_{s}=10$ g/L on.
The conclusion is that the glucose concentrations in the experiments of 
Pirt \& Rhigelato probably were too low for glucose repression to be 
detected. The experimental data published by Ryu \& Hospodka 
are not detailed sufficiently to permit a similar analysis.

\begin{table} 
\caption[]{Parameter sets used by Bajpai \& Reu\ss\ } 
\begin{tabular}{lrll}                                       
\hline
\multicolumn{2}{l}{\it parameter} & {\it Set 1} & {\it Set 2}\\ 
\hline  
$\mu_{x}$           & [h$^{-1}$]  & 0.092       & 0.11          \\
$K_{x}$             & [g/g DM]     & 0.15        & 0.006         \\
$\mu_{p}$           & [g/g DM h]  & 0.005       & 0.004         \\
$K_{p}$             & [g/L]        & 0.0002      & 0.0001        \\
$K_{i}$             & [g/L]        & 0.1         & 0.1           \\
$Y_{x/s}$           & [g DM/g]     & 0.45        & 0.47          \\
$Y_{p/s}$           & [g/g]        & 0.9         & 1.2           \\
$k_{h}$             & [h$^{-1}$]  & 0.04        & 0.01          \\
$m_{s}$             & [g/g DM h]  & 0.014       & 0.029         \\ 
\hline
\end{tabular}
\end{table}

Bajpai \& Reu\ss\ decided to disregard the
differences between time constants for the two regulation mechanisms
(glucose repression or inhibition) because of the
relatively very long fermentation times, and therefore proposed a Haldane
expression for $\pi$.

It is interesting that simulations with the \cite{Bunt} model for the
initial conditions given by these authors indicate that, when the
remaining substrate is fed at a constant rate, a considerable and 
unrealistic amount of penicillin is 
produced when the glucose concentration is still very high \cite{CarberryCL88}
Simulations with the Bajpai \& Reu\ss\ model correctly predict almost        
no penicillin production in similar conditions.

The maintenance coefficient used by \cite{Bunt} 
($m_{s}=0.025$ g/g DM h) corresponds well to
the value $m_{s}=0.029$ g/g DM h (Set 2 of \cite{BuchananRBES}), to the 
value $m_{s}=0.024$ g/g DM h reported by \citeauthor{MMI2-d3}, and to the
value used by \shortcite{CawseyIJMMS} ($m_{s}=0.022$ g/g DM h) (1983).
However, these values differ from the value in Set 1 of 
\cite{BuchananRBES} ($m_{s}=0.014$ g/g DM h). 
It is not clear where this difference originated from.
Simulations indicated that the dynamic behaviour of the model is rather
sensitive with respect to the value of $m_{s}$.

In the model van \shortcite{Bunt}, at severe substrate limitation conditions, and thus most probably corresponding
to endogenous metabolic behaviour, the biomass consumption due to maintenance
and production requirements may exceed the conversion of substrate into
biomass and $\mu$ eventually may become negative. This situation may occur at
the end of the growth phase during a fed-batch fermentation.
For these conditions $\pi$ is not defined. A
straightforward extension of the $\pi(\mu)$ kinetics (10)
could be $\pi(\mu \leq 0)=0$, but there
are some biochemical indications that 
the penicillin biosynthesis actually does not stop in that case.

\begin{figure} % figuur 1
\vspace{6pc}
\caption[]{Pathway of the penicillin G biosynthesis.}
\label{penG}
\end{figure}

Sample of cross-reference to figure.
Figure~\ref{penG} shows that is not easy to get something on paper.

\section{Headings}

\subsection{Subsection}
based their model on balancing methods and
biochemical know\-ledge. The original model (1980) contained an equation for the
oxygen dynamics which has been omitted in a second paper 
(1981). This simplified model shall be discussed here.

\subsubsection{Subsubsection}
\cite{Carr-Goldstein,Cohen-Jones88-book} 
based their model on balancing methods and
biochemical know\-ledge. The original model (1980) contained an equation for the
oxygen dynamics which has been omitted in a second paper 
(1981). This simplified model shall be discussed here.

\paragraph{Paragraph.}
\cite{Carr-Goldstein,Cohen-Jones88-book} 
based their model on balancing methods and
biochemical know\-ledge. The original model (1980) contained an equation for the
oxygen dynamics which has been omitted in a second paper 
(1981). This simplified model shall be discussed here.

\subparagraph{Subparagraph.}
\cite{Carr-Goldstein,Cohen-Jones88-book} 
based their model on balancing methods and
biochemical know\-ledge. The original model (1980) contained an equation for the
oxygen dynamics which has been omitted in a second paper 
(1981). This simplified model shall be discussed here.

\section{Equations and the Like}

Two equations:
\begin{equation}
    C_{s}  =  K_{M} \frac{\mu/\mu_{x}}{1-\mu/\mu_{x}} \label{ccs}
\end{equation}

and

\begin{equation}
    G = \frac{P_{\rm opt} - P_{\rm ref}}{P_{\rm ref}} \mbox{\ }100 \mbox{\ }(\%)
\end{equation}

Two equation arrays:

\begin{eqnarray}
  \frac{dS}{dt} & = & - \sigma X + s_{F} F\\
  \frac{dX}{dt} & = &   \mu    X\\
  \frac{dP}{dt} & = &   \pi    X - k_{h} P\\
  \frac{dV}{dt} & = &   F
\end{eqnarray}

and,

\begin{eqnarray}
 \mu_{\rm substr} & = & \mu_{x} \frac{C_{s}}{K_{x}C_{x}+C_{s}}  \\
 \mu              & = & \mu_{\rm substr} - Y_{x/s}(1-H(C_{s}))(m_{s}+\pi /Y_{p/s}) \\
 \sigma           & = & \mu_{\rm substr}/Y_{x/s}+ H(C_{s}) (m_{s}+ \pi /Y_{p/s})
\end{eqnarray}

\section{References in the THEBIBLIOGRAPHY Environment}
The first reference in the list below has the key \{BrownAndBurton\}.
Together with the bibitem option $\backslash$citeauthoryear\{Brown and
Burton\}\{1978\} the follwing cross-references can be used:\\
$\backslash$cite\{BrownAndBurton\} produces: \cite{BrownAndBurton}\\
$\backslash$shortcite\{BrownAndBurton\}  produces: \shortcite{BrownAndBurton}\\  
$\backslash$citeauthor\{BrownAndBurton\}  produces: \citeauthor{BrownAndBurton}\\  
$\backslash$citeyear\{BrownAndBurton\}  produces: \citeyear{BrownAndBurton}\\

\appendix

And this is my Appendix.

\acknowledgements
And this is an acknowledgements section with a heading that was produced by the
$\backslash$acknowledgements command. Thank you all for helping me writing this
\LaTeX\ sample file.

% The endnotes section will be placed here.

\theendnotes

\begin{thebibliography}{}

\bibitem[\protect\citeauthoryear{Brown and Burton}{1978}]{BrownAndBurton}
J.~S. Brown and R.~R. Burton.
\newblock {Diagnostic Models for Procedural Bugs in Basic Mathematical Skills}.
\newblock {\em Cognitive Science}, 2(2):155--192, 1978.

\bibitem[\protect\citeauthoryear{Buchanan}{1984}]{BuchananRBES}
Bruce~G. Buchanan and Edward~H. Shortliffe.
\newblock {\em Rule-Based Expert Systems: The MYCIN Experiments of the Stanford
  Heuristic Programming Project}.
\newblock Addison-Wesley Publishing Company, 1984.

\bibitem[\protect\citeauthoryear{Bunt}{1990}]{Bunt}
H.~C. Bunt.
\newblock {Modular Incremental Modelling of Belief and Intention}.
\newblock In {\em Proceedings of the Second International Workshop on User
  Modeling}, 1990.

\bibitem[\protect\citeauthoryear{Cahour}{1990}]{Cahour-90-Montreal}
B\'eatrice Cahour.
\newblock {Competence Modelling in Consultation Dialogs}.
\newblock In L.~Berlinguet and D.~Berthelette, editors, {\em Proceedings of the
  International Congress, Work With Dispay Units' 89}, Montreal, Canada,
  September 1990. North Holland, Amsterdam.

\bibitem[\protect\citeauthoryear{Cahour}{1988}]{Cahour-thesis}
B\'eatrice Cahour.
\newblock {\em {La Mod\'elisation de l'Interlocuteur: Elaboration du Mod\`ele
  et Effets au Cours de Dialogues de Consultation}}.
\newblock PhD thesis, Universit\'e Paris 8, France, 1991.
\newblock Cognitive psychology PhD.

\bibitem[\protect\citeauthoryear{Carberry}{1988}]{CarberryCL88}
Sandra~M. Carberry.
\newblock {Modeling the User's Plans and Goals}.
\newblock {\em Computational Linguistics}, 14(3):23--37, September 1988.

\bibitem[\protect\citeauthoryear{Carbonell}{1970}]{CarbonellScholar}
J.~R. Carbonell.
\newblock {AI in CAI: An Artificial Intelligence Approach to Computer-Aided
  Instruction}.
\newblock {\em IEEE Transactions on Man-Machine Systems}, 11:190--202, 1970.

\bibitem[\protect\citeauthoryear{Carr and Goldstein}{1977}]{Carr-Goldstein}
B.~Carr and I.~Goldstein.
\newblock {Overlays: A Theory of Modelling for Computed Aided Instruction}.
\newblock AI Memo 406, 1977.

\bibitem[\protect\citeauthoryear{Cawsey}{(in press)}]{CawseyIJMMS}
Alison Cawsey.
\newblock {Planning Interactive Explanations}.
\newblock {\em International Journal of Man-Machine Studies}, in press.

\bibitem[\protect\citeauthoryear{Chandrasekaran and Swartout}{1991}]{chandra-swartout}
B.~Chandrasekaran and William Swartout.
\newblock {Explanations in Knowledge Systems: The Role of Explicit
  Representation of Design Knowledge}.
\newblock {\em IEEE Expert}, 6(3):47--50, June 1991.

\bibitem[\protect\citeauthoryear{Chappel and Cahour}{1991}]{MMI2-d3}
Helen Chappel and B\'eatrice Cahour.
\newblock {User Modeling for Multi-Modal Co-Operative Dialogue with KBS}.
\newblock Deliverable D3, Esprit Project P2474, 1991.

\bibitem[\protect\citeauthoryear{Chin}{1987}]{ChinThesis}
David~N. Chin.
\newblock {\em Intelligent Agents as a Basis for Natural Language Interfaces}.
\newblock PhD thesis, University of California at Berkeley, 1987.

\bibitem[\protect\citeauthoryear{Chin}{1989}]{Chin88-book}
David~N. Chin.
\newblock {KNOME}: {M}odeling {W}hat the {U}ser {K}nows in {UC}.
\newblock In Alfred Kobsa and Wolfgang Wahlster, editors, {\em User Models in
  Dialog Systems}. Springer-Verlag, Symbolic Computation Series, Berlin
  Heidelberg New York Tokyo, 1989.

\bibitem[\protect\citeauthoryear{Cohen and Jones}{1989}]{Cohen-Jones88-book}
Robin Cohen and Marlene Jones.
\newblock {Incorporating User Models into Expert Systems for Educational
  Diagnosis}.
\newblock In Alfred Kobsa and Wolfgang Wahlster, editors, {\em User Models in
  Dialog Systems}, pages 35 -- 51. Springer-Verlag, Symbolic Computation
  Series, Berlin Heidelberg New York Tokyo, 1989.

\bibitem[\protect\citeauthoryear{Falzon}{1987}]{Falzon87}
Pierre Falzon.
\newblock {Les Dialogues de Diagnostic: L'\'evaluation des Connaissances de
  l'Interlocuteur}.
\newblock Technical Report 747, INRIA, Rocquencourt, France, 1987.

\bibitem[\protect\citeauthoryear{Finin et al.}{1986}]{FininJoshiWebber}
Timothy~W. Finin, Aravind~K. Joshi, and Bonnie~Lynn Webber.
\newblock {Natural Language Interactions with Artificial Experts}.
\newblock {\em Proceedings of the IEEE}, 74(7), July 1986.

\end{thebibliography}
\end{article}
\end{document}

### 2009-05-08 mtjournal.tex
fleshed out title and opening

--- 2009-05-08 kluwer/klusampl.tex
@@
 \begin{opening}         
-\title{A Sample Document\thanks{Footnote 
-            to the title with the `thanks' command.}} 
-\author{Leslie \surname{Lamport}}  
-\runningauthor{Leslie Lamport}
-\runningtitle{A Sample Document}
-\institute{Author's Affiliation}
-\date{April 15, 1993}
+\title{Expected Dependency Pair Match:\\
+Predicting improvements to {HTER} with expected syntactic structure} 
+\author{Jeremy G. \surname{Kahn}\email{jgk@u.washington.edu}}  
+\institute{University of Washington}
+\author{Matthew \surname{Snover}\email{snover@cs.umd.edu}}
+\institute{University of Maryland}
+\author{Mari \surname{Ostendorf}\email{ostendor@u.washington.edu}}  
+\institute{University of Washington}
+\runningauthor{Kahn, Snover \& Ostendorf}
+\runningtitle{Expected Dependency Pair Match}
+\date{May 10, 2009}
 
 \begin{abstract}
-This is a sample input file.  Comparing it with the output it
-generates can show you how to produce a simple document of
-your own.
+Abstract to go here
 \end{abstract}
-\keywords{sample, \LaTeX}
+\keywords{machine translation evaluation, syntax, dependency trees}
 

### 2009-05-08 mtjournal.tex
put outline of intro, paradigm sections into kluwer tex

@@
 
-\section{Ordinary Text}  
-                    % Produces section heading.  Lower-level
-                    % sections are begun with similar 
-                    % \subsection and \subsubsection commands.
+\section{Introduction}
 
-The ends  of words and sentences are marked 
-  by   spaces. It  doesn't matter how many 
-spaces    you type; one is as good as 100.  The
-end of   a line counts as a space.
+\begin{itemize}
+\item evaluation [esp. automatic evaluation]
+\item popular current
+  measures BLEU, TER
 
-One   or more   blank lines denote the  end 
-of  a paragraph.
+\item challenges - motivate search for better measures
 
-Since any number of consecutive spaces are treated like a single
-one, the formatting of the input file makes no difference to
-      \TeX,         % The \TeX command generates the TeX logo.
-but it makes a difference to you.  
-When you use
-      \LaTeX,       % The \LaTeX command generates the LaTeX logo.
-making your input file as easy to read as possible
-will be a great help as you write your document and when you
-change it.  This sample file shows how you can add comments to
-your own input file.
+\item other extensions:
+  \begin{itemize}
+  \item METEOR 
+    \begin{itemize}
+    \item stemming and synonym tables
+    \item tuned but not generally     tuneable
+    \end{itemize}
+  \item TERp
+    \begin{itemize}
+    \item development-set tuning by user
+    \item stemming and synonym tables
+    \item paraphrase tables
+    \end{itemize}
+  \end{itemize}
+\item prior work with (partial) syntactic locality:
+  \begin{itemize}
+  \item Liu \& Gildea 
+  \item Roark et al. Sparseval
+  \item Owczarzak et al. 2007
+  \end{itemize}
 
-Because printing is different from typewriting, there are a 
-number of things that you have to do differently when preparing 
-an input file than if you were just typing the document directly.  
-Quotation marks like 
-       ``this'' 
-have to be handled specially, as do quotes within quotes: 
-       ``\,`this'                  % \, separates the double and single quote.
-    is what I just 
-    wrote, not  `that'\,''.
+\item Evaluation of evaluation measures
+  \begin{itemize}
+  \item correlation vs. human fluency/adequacy 
+  \item correlation vs. human-in-the-loop measures (HTER)
+  \end{itemize}
+\item one challenge of metric-evaluation: underlying doc difficulty 
+  \begin{itemize}
+  \item describe mean-subtraction
+  \end{itemize}
+\end{itemize}
 
-Dashes come in three sizes: an 
-       intra-word 
-dash, a medium dash for number ranges like 
-       1--2, 
-and a punctuation 
-       dash---like 
-this.
+\section{Experimental paradigm}
 
-A sentence-ending space should be larger than the space between words
-within a sentence.  You sometimes have to type special commands in
-conjunction with punctuation characters to get this right, as in the
-following sentence.
-       Gnats, gnus, etc.\    % `\ ' makes an inter-word space.
-       all begin with G\@.   % \@ marks end-of-sentence punctuation.
-You should check the spaces after periods when reading your output to
-make sure you haven't forgotten any special cases.
-Generating an ellipsis 
-       \ldots\    % `\ ' needed because TeX ignores spaces after 
-          % command names like \ldots made from \ + letters.
-          %
-          % Note how a `%' character causes TeX to ignore the 
-          % end of the input line, so these blank lines do not
-          % start a new paragraph.
-with the right spacing around the periods 
-requires a special  command.
+Section explains space of measures to explore
 
-\TeX\ interprets some common characters as commands, so you must type
-special commands to generate them.  These characters include the
-following: 
-       \$ \& \% \# \{ and~\}.
+\begin{itemize}
+\item F-measure over various bags of components of parse tree
 
-In printing, text is emphasized by using an
-       {\em italic\/}  % The \/ command produces the tiny extra space that
-               % should be added between a slanted and a following
-               % unslanted letter.
-type style.
+  [ Figure describing Sparseval-ish variant ]
 
-\begin{em}
-   A long segment of text can also be emphasized in this way.  Text within
-   such a segment given additional emphasis 
-      with\/ {\em Roman} 
-   type.  Italic type loses its ability to emphasize and become simply
-   distracting when used excessively.  
-\end{em}
+\item Listing possible component valences
+  \begin{itemize}
+  \item dependent + labeled link + head 
+  \item outbound only (dependent +
+    label)
+  \item inbound only (label + head)
+  \item word-word (skipping label)
+  \end{itemize}
+  \begin{itemize}
+  \item 1-grams
+  \item 2-grams
+  \item \ldots{}
+  \end{itemize}
 
-It is sometimes necessary to prevent \TeX\ from breaking a line where
-it might otherwise do so.  This may be at a space, as between the
-``Mr.'' and ``Jones'' in
-       ``Mr.~Jones'',        % ~ produces an unbreakable interword space.
-or within a word---especially when the word is a symbol like
-       \mbox{\em itemnum\/} 
-that makes little sense when hyphenated across 
-       lines.
+  Note: inbound links \& full pairs "overweight" head words; this is
+  a good thing.
 
-\TeX\ is good at typesetting mathematical formulas like
-       \( x-3y = 7 \) 
-or
-       \( a_{1} > x^{2n} / y^{2n} > x' \).
-Remember that a letter like
-       $x$        % $ ... $  and  \( ... \)  are equivalent
-is a formula when it denotes a mathematical symbol, and should
-be treated as one.
+  [ figure demonstrating F-measure over d+l+h decomposition of toy
+  trees ]
+
+\item Tree features may be extracted from a 1-best tree, or from an
+  n-best tree list
+
+  \begin{itemize}
+  \item  counts are an expectation over the forest
+  \item n parameter: how many trees in the forest
+  \item gamma parameter for flattening overconfident parse hyps
+  \end{itemize}
+  note intentional similarities to Owczarzak paper
+
+\item Components in the basic DPM measures combined in various ways:
+
+  \begin{itemize}
+  \item combination of precision, recall scores for different
+    components (on analogy with BLEU combination), naive assumption:
+    all weighted equally
+
+  \item simple F over items of all valences (naive assumption: all
+    weighted equally)
+  \end{itemize}
+\item Tuneable weights:
+  \begin{itemize}
+  \item recall that TERp has tuneable system over small number of parameters
+    \begin{itemize}
+    \item insertion weight
+    \item deletion weight
+    \item substitution weight
+    \item synonym weight (motivated with WordNet)
+    \end{itemize}
+  \item learn weights for separate precision, recall for various
+    components
+  \end{itemize}
+
+\item Implementation
+  \begin{itemize}
+  \item Charniak parser
+  \item n=1..50
+  \item head-finding, with some modifications
+  \item label-construction
+  \end{itemize}
+\end{itemize}
 

### 2009-05-08 mtjournal.tex
removed dummy notes and displayed-text sections

@@
 \end{itemize}
-
-\section{Notes}
-Footnotes\footnote{This is an example of a footnote.}
-pose no problem\footnote{And another one}.
-Endnotes\endnote{This is an example of an endnote.} pose no 
-problem either\endnote{And another one.}.
-
-When using the {\sc kluwer} style file, you can produce endnotes 
-analogous to footnotes. Instead of the \LaTeX\ command \verb+\footnote{}+,
-you can use \verb+\endnote{}+. The command \verb+\theendnotes+ can be used
-to place the endnotes in the text. They will be put in a separate section in
-the \verb+\footnotesize+ font.
-
-\section{Displayed Text}
-
-Text is displayed by indenting it from the left margin.
-Quotations are commonly displayed.  There are short quotations
-\begin{quote}
-   This is a short a quotation.  It consists of a 
-   single paragraph of text.  There is no paragraph
-   indentation.
-\end{quote}
-and longer ones.
-\begin{quotation}
-   This is a longer quotation.  It consists of two paragraphs
-   of text.  The beginning of each paragraph is indicated
-   by an extra indentation.
-
-   This is the second paragraph of the quotation.  It is just
-   as dull as the first paragraph.
-\end{quotation}
-Another frequently-displayed structure is a list.
-The following is an example of an {\em itemized} list, four levels deep.
-\begin{itemize}
-\item  This is the first item of an itemized list.  Each item 
-      in the list is marked with a ``tick''.  The document
-      style determines what kind of tick mark is used.
-\item  This is the second item of the list.  It contains another
-      list nested inside it.  The three inner lists are an {\em itemized}
-      list.
-    \begin{itemize}
-       \item This is the first item of an enumerated list that
-            is nested within the itemized list.
-          \item This is the second item of the inner list.  \LaTeX\
-            allows you to nest lists deeper than you really should.
-      \end{itemize}
-      This is the rest of the second item of the outer list.  It
-      is no more interesting than any other part of the item.
-   \item  This is the third item of the list.
-\end{itemize}
-
-The following is an example of an {\em enumerated} list, four levels deep.
-\begin{enumerate}
-\item  This is the first item of an enumerated list.  Each item 
-      in the list is marked with a ``tick''.  The document
-      style determines what kind of tick mark is used.
-\item  This is the second item of the list.  It contains another
-      list nested inside it.  The three inner lists are an {\em enumerated}
-      list.
-    \begin{enumerate}
-       \item This is the first item of an enumerated list that
-            is nested within the enumerated list.
-          \item This is the second item of the inner list.  \LaTeX\
-            allows you to nest lists deeper than you really should.
-      \end{enumerate}
-      This is the rest of the second item of the outer list.  It
-      is no more interesting than any other part of the item.
-   \item  This is the third item of the list.
-\end{enumerate}
-
-The following is an example of a {\em description} list.
-\begin{description}
-\item[Cow] Highly intelligent animal that can produce milk out of grass.
-\item[Horse] Stupid animal that has legs.
-\item[Human being] Stupid animal that can talk and think.
-\end{description}
-
-You can even display poetry.
-\begin{verse}
-   There is an environment for verse \\    % The \\ command separates lines
-   Whose features some poets will curse.   % within a stanza.
-
-               % One or more blank lines separate stanzas.
-
-   For instead of making\\
-   Them do {\em all\/} line breaking, \\
-   It allows them to put too many words on a line when they'd 
-   rather be forced to be terse.
-\end{verse}
-
-Mathematical formulas may also be displayed.  A displayed formula is
-one-line long; multiline formulas require special formatting
-instructions.
-   \[  x' + y^{2} = z_{i}^{2}\]
-Don't start a paragraph with a displayed equation, nor make
-one a paragraph by itself.
-
-Example of a theorem:
-
-\begin{guess}
-All conjectures are interesting, but some conjectures are more
-interesting than others.
-\end{guess}
-
-\section{Tables and Figures}
-Cross reference to labelled table: As you can see in Table~\ref{sphericcase} on 
-page~\pageref{sphericcase} and also in Table~\ref{parset} on page~\pageref{parset}.
-
-\begin{table} % 
-\begin{tabular}{lrl}                                        
-\hline
-$Q_{s,\max}$   & [g/g DM h]  & 0.18\\
-$K_{s}$       & [g/L]        & 1.0\\
-$Y_{x/s}$     & [g DM/g]     & 0.5\\
-$Y_{p/s}$     & [g/g]        & 0.854\\
-$Q_{p,\max}$   & [g/g DM h]  & 0.0045\\
-$\mu_{\rm crit}$  & [h$^{-1}$]  & 0.01\\
-$k_{h}$       & [h$^{-1}$]  & 0.002\\
-$m_{s}$       & [g/g DM h]  & 0.025\\                              
-\hline
-\end{tabular}
-\caption[]{Parameter set used in the model of \shortcite{Bunt}}\label{parset}
-\end{table}
-
-\begin{table*}
-\caption[]{The spherical case ($I_1=0$, $I_2=0$).}
-\label{sphericcase}
-\begin{tabular}{crrrrc}
-\hline
-Equil. \\
-Points & \multicolumn{1}{c}{$x$} & \multicolumn{1}{c}{$y$} & \multicolumn{1}{c}{$z$} & \multicolumn{1}{c}{$C$} &
-S \\
-\hline
-$~~L_1$ & $-$2.485252241 & 0.000000000 & 0.017100631 & 8.230711648 & U \\
-$~~L_2$ &    0.000000000 & 0.000000000 & 3.068883732 & 0.000000000 & S \\
-$~~L_3$ &    0.009869059 & 0.000000000 & 4.756386544 & $-$0.000057922 & U \\
-$~~L_4$ &    0.210589855 & 0.000000000 & $-$0.007021459 & 9.440510897 & U \\
-$~~L_5$ &    0.455926604 & 0.000000000 & $-$0.212446624 & 7.586126667 & U \\
-$~~L_6$ &    0.667031314 & 0.000000000 & 0.529879957 & 3.497660052 & U \\
-$~~L_7$ &    2.164386674 & 0.000000000 & $-$0.169308438 & 6.866562449 & U \\
-$~~L_8$ &    0.560414471 & 0.421735658 & $-$0.093667445 & 9.241525367 & U \\
-$~~L_9$ &    0.560414471 & $-$0.421735658 & $-$0.093667445 & 9.241525367 & U
-\\
-$~~L_{10}$ & 1.472523232 & 1.393484549 & $-$0.083801333 & 6.733436505 & U \\
-$~~L_{11}$ & 1.472523232 & $-$1.393484549 & $-$0.083801333 & 6.733436505 & U
-\\ \hline
-\end{tabular}
-\end{table*}
-
-A major
-point of difference lies in the value of the specific production rate $\pi$ for
-large values of the specific growth rate $\mu$.
-Already in the early publications \cite{Falzon87}
-it appeared that high glucose
-concentrations in the production phase are well correlated with a
-low penicillin yield (the 
-`glucose effect'). It has been confirmed recently 
-\cite{Bunt,Cahour-thesis,BrownAndBurton,Carr-Goldstein}
-that
-high glucose concentrations inhibit the synthesis of the enzymes of the
-penicillin pathway, but not the actual penicillin biosynthesis.
-In other words, glucose represses (and not inhibits) the penicillin
-biosynthesis.
-
-These findings do not contradict the results of 
-\citeauthor{Chin88-book} (on which \shortcite{Bunt} based their 
-production kinetics) and of \citeyear{ChinThesis} which were obtained for
-continuous culture fermentations.                
-Because for high values of the specific
-growth rate $\mu$ it is most likely (as shall be discussed below) that 
-maintenance metabolism occurs, it can be shown that
-in steady state continuous culture conditions, and with $\mu$ described by a Monod kinetics
-\begin{equation}
-    C_{s}  =  K_{M} \frac{\mu/\mu_{x}}{1-\mu/\mu_{x}} \label{cs}
-\end{equation}
-Pirt \& Rhigelato determined $\pi$ for $\mu$ between 
-$0.023$ and $0.086$ h$^{-1}$.
-They also reported a value $\mu_{x} \approx 0.095$
-h$^{-1}$, so that for their experiments $\mu/\mu_{x}$ is in the range 
-of $0.24$ to $0.9$. 
-Substituting $K _M$ in Eq. (\ref{cs}) by 
-the value $K_{M}=1$ g/L as used by \cite{Bunt}, one finds
-with the above equation $0.3 < C_{s} < 9$ g/L. This agrees well with 
-the work of  \shortcite{Bunt}, who reported that penicillin biosynthesis 
-repression only occurs at glucose concentrations from $C_{s}=10$ g/L on.
-The conclusion is that the glucose concentrations in the experiments of 
-Pirt \& Rhigelato probably were too low for glucose repression to be 
-detected. The experimental data published by Ryu \& Hospodka 
-are not detailed sufficiently to permit a similar analysis.
-
-\begin{table} 
-\caption[]{Parameter sets used by Bajpai \& Reu\ss\ } 
-\begin{tabular}{lrll}                                       
-\hline
-\multicolumn{2}{l}{\it parameter} & {\it Set 1} & {\it Set 2}\\ 
-\hline  
-$\mu_{x}$           & [h$^{-1}$]  & 0.092       & 0.11          \\
-$K_{x}$             & [g/g DM]     & 0.15        & 0.006         \\
-$\mu_{p}$           & [g/g DM h]  & 0.005       & 0.004         \\
-$K_{p}$             & [g/L]        & 0.0002      & 0.0001        \\
-$K_{i}$             & [g/L]        & 0.1         & 0.1           \\
-$Y_{x/s}$           & [g DM/g]     & 0.45        & 0.47          \\
-$Y_{p/s}$           & [g/g]        & 0.9         & 1.2           \\
-$k_{h}$             & [h$^{-1}$]  & 0.04        & 0.01          \\
-$m_{s}$             & [g/g DM h]  & 0.014       & 0.029         \\ 
-\hline
-\end{tabular}
-\end{table}
-
-Bajpai \& Reu\ss\ decided to disregard the
-differences between time constants for the two regulation mechanisms
-(glucose repression or inhibition) because of the
-relatively very long fermentation times, and therefore proposed a Haldane
-expression for $\pi$.
-
-It is interesting that simulations with the \cite{Bunt} model for the
-initial conditions given by these authors indicate that, when the
-remaining substrate is fed at a constant rate, a considerable and 
-unrealistic amount of penicillin is 
-produced when the glucose concentration is still very high \cite{CarberryCL88}
-Simulations with the Bajpai \& Reu\ss\ model correctly predict almost        
-no penicillin production in similar conditions.
-
-The maintenance coefficient used by \cite{Bunt} 
-($m_{s}=0.025$ g/g DM h) corresponds well to
-the value $m_{s}=0.029$ g/g DM h (Set 2 of \cite{BuchananRBES}), to the 
-value $m_{s}=0.024$ g/g DM h reported by \citeauthor{MMI2-d3}, and to the
-value used by \shortcite{CawseyIJMMS} ($m_{s}=0.022$ g/g DM h) (1983).
-However, these values differ from the value in Set 1 of 
-\cite{BuchananRBES} ($m_{s}=0.014$ g/g DM h). 
-It is not clear where this difference originated from.
-Simulations indicated that the dynamic behaviour of the model is rather
-sensitive with respect to the value of $m_{s}$.
-
-In the model van \shortcite{Bunt}, at severe substrate limitation conditions, and thus most probably corresponding
-to endogenous metabolic behaviour, the biomass consumption due to maintenance
-and production requirements may exceed the conversion of substrate into
-biomass and $\mu$ eventually may become negative. This situation may occur at
-the end of the growth phase during a fed-batch fermentation.
-For these conditions $\pi$ is not defined. A
-straightforward extension of the $\pi(\mu)$ kinetics (10)
-could be $\pi(\mu \leq 0)=0$, but there
-are some biochemical indications that 
-the penicillin biosynthesis actually does not stop in that case.
-
-\begin{figure} % figuur 1
-\vspace{6pc}
-\caption[]{Pathway of the penicillin G biosynthesis.}
-\label{penG}
-\end{figure}
-
-Sample of cross-reference to figure.
-Figure~\ref{penG} shows that is not easy to get something on paper.
 

### 2009-05-08 mtjournal.tex
included correlations section from outline

@@
   \end{itemize}
+\end{itemize}
+
+\section{Correlation with human judgments of fluency \& adequacy}
+
+Corpus: LDC Multiple Translation Chinese corpus 3\&4 treat each
+segment and each judgment as independent data point
+
+\begin{itemize}
+\item EXPT
+
+  using n=1 parse for DPM variants
+
+\begin{verbatim}
+   metric      r
+  --------   -----
+  F[dl;lh]   0.226
+  BLEU4      0.218  # +1 smoothing
+  F[dlh]     0.185
+  TER       -0.173
+\end{verbatim}
+
+  point: using partial-labeling (dl,lh) much better than full
+  link. Also, using Charniak parser as labeler seems to work, LFG not
+  necessary
+
+\item EXPT
+
+  combining precision, recall naively vs. combining items
+
+\begin{verbatim}
+  F(1g,2g,dl,lh)        0.237
+  prmeans(1g,2g,dl,lh)  0.217
+
+  F(dl,lh)              0.226
+  prmeans(dl,lh)        0.208
+\end{verbatim}
+
+  point: combining individual items naively (before computing
+  precision and recall) seems better than naively combining precisions
+  and recalls (too many chances for zeroes?)
+
+\item EXPT
+  expanding n, gamma=0
+
+\begin{verbatim}
+  F[1g,2g,dl,lh],n=50   0.239
+    ...          n=1    0.237
+  F[dl,lh],n=50         0.234
+    ...    n=1          0.226
+\end{verbatim}
+  point: larger n $\to$ better r
+  (holds for other comparisons too, but these are good examples)
+
+\item  EXPT
+
+  setting gamma
+
+  Tuning experiments finds that gamma of 0.25 is slightly better
+  (especially when using F[dl,lh]) than gammas of other values in
+  range 0-1 (these values probably not significant)
+
 \end{itemize}

### 2009-05-08 mtjournal.tex
removed dummy sections on Headings and on Equations

@@
 
-\section{Headings}
-
-\subsection{Subsection}
-based their model on balancing methods and
-biochemical know\-ledge. The original model (1980) contained an equation for the
-oxygen dynamics which has been omitted in a second paper 
-(1981). This simplified model shall be discussed here.
-
-\subsubsection{Subsubsection}
-\cite{Carr-Goldstein,Cohen-Jones88-book} 
-based their model on balancing methods and
-biochemical know\-ledge. The original model (1980) contained an equation for the
-oxygen dynamics which has been omitted in a second paper 
-(1981). This simplified model shall be discussed here.
-
-\paragraph{Paragraph.}
-\cite{Carr-Goldstein,Cohen-Jones88-book} 
-based their model on balancing methods and
-biochemical know\-ledge. The original model (1980) contained an equation for the
-oxygen dynamics which has been omitted in a second paper 
-(1981). This simplified model shall be discussed here.
-
-\subparagraph{Subparagraph.}
-\cite{Carr-Goldstein,Cohen-Jones88-book} 
-based their model on balancing methods and
-biochemical know\-ledge. The original model (1980) contained an equation for the
-oxygen dynamics which has been omitted in a second paper 
-(1981). This simplified model shall be discussed here.
-
-\section{Equations and the Like}
-
-Two equations:
-\begin{equation}
-    C_{s}  =  K_{M} \frac{\mu/\mu_{x}}{1-\mu/\mu_{x}} \label{ccs}
-\end{equation}
-
-and
-
-\begin{equation}
-    G = \frac{P_{\rm opt} - P_{\rm ref}}{P_{\rm ref}} \mbox{\ }100 \mbox{\ }(\%)
-\end{equation}
-
-Two equation arrays:
-
-\begin{eqnarray}
-  \frac{dS}{dt} & = & - \sigma X + s_{F} F\\
-  \frac{dX}{dt} & = &   \mu    X\\
-  \frac{dP}{dt} & = &   \pi    X - k_{h} P\\
-  \frac{dV}{dt} & = &   F
-\end{eqnarray}
-
-and,
-
-\begin{eqnarray}
- \mu_{\rm substr} & = & \mu_{x} \frac{C_{s}}{K_{x}C_{x}+C_{s}}  \\
- \mu              & = & \mu_{\rm substr} - Y_{x/s}(1-H(C_{s}))(m_{s}+\pi /Y_{p/s}) \\
- \sigma           & = & \mu_{\rm substr}/Y_{x/s}+ H(C_{s}) (m_{s}+ \pi /Y_{p/s})
-\end{eqnarray}
-
 \section{References in the THEBIBLIOGRAPHY Environment}

### 2009-05-09 mtjournal.tex
included section on initial correlations with HTER

@@
 
+\section{Correlations with HTER}
+
+  Corpus
+    Gale translation corpus 2.5; 3 sites, 2 source languages, 4 genres each
+
+  prediction:
+    Mean-subtracted HTER per-document
+
+  base measure EDPM: F[1g,2g,dl,lh],n=50,gamma=0.25
+
+  [EXPT]
+\begin{verbatim}
+            Ar     Zh    All
+   TER     0.51   0.32  0.44
+   BLEU_4 -0.42  -0.33 -0.37
+   EDPM   -0.60* -0.39 -0.50*
+\end{verbatim}
+  point: base EDPM does better on all docs, and on Arabic. difference
+  is not quite significant at the p=0.05 on the Chinese documents
+
+  Note 1: also tried pairwise deltas per-document, with similar
+  results, as presented in the MetricsMATR competition submission.
+
+  Note 2: when using per-segment mean-subtracted, r values are smaller
+  and *Arabic* was the language where EDPM did not perform best.
+
+\begin{verbatim}
+               Ar     Zh     All
+      TER     0.38*  0.04   0.19
+      BLEU_4 -0.10  -0.16  -0.14
+      EDPM   -0.31  -0.19* -0.24*
+\end{verbatim}
+
+  (Go into details of per-genre differences here? or leave out, for
+  space reasons?)
+
 \section{References in the THEBIBLIOGRAPHY Environment}

### 2009-05-09 mtjournal.tex
added section on weight-tuning from outline

@@
   (Go into details of per-genre differences here? or leave out, for
-  space reasons?)
+  space reasons? -- leave out, I'd say)
+
+\section{Weight-tuning to combine syntax and other knowledge sources}
+
+TERp optimizer [Matt, a brief description here]
+
+TERp comes with a variety of features (described earlier)
+
+include optionally new features
+
+Corpus: GALE 2.5 data, again documents randomly split into two groups,
+evenly split across language and genre
+
+tuning target: weighted correlation with mean-removed segment HTER
+
+[we use weighted correlation to avoid emphasis on short segments
+(which hurts document-level and system-level utility) ]
+
+Select features from set:
+
+\begin{description}
+\item[E]: features from EDPM, specifically, P,R for inbound, outbound,
+  full-links, and unlabeled links  (8 features)
+\item[T]: features from basic TERp
+  (inserts, deletes, substitutions, shifts)  (7 features)
+\item[P]: features from TERp paraphraser
+  (4 features)
+\item[N]: precision and recall for 1-grams, 2-grams
+\item[B]: brevity/prolixity (2 features: one for longer-than-ref, one for
+  shorter-than-ref)
+\end{description}
+
+tune weights on one set, test on the other (and vice versa).
+reporting average r between the two.
+
+\begin{verbatim}
+   == Tuned For Weighted Mean Removed Pearson
+     - Examining WGT_MEAN_RM_SEG_PEAR ==
+            Average
+      Cond   Test  
+     ---------------
+      ETPB   0.4096
+      ETP    0.4038
+      TPB    0.4010
+      TP     0.3946
+      TPNB   0.3940
+      TPN    0.3891
+      ETB    0.3871
+      ET     0.3827
+      TB     0.3815
+      ETNB   0.3814
+      ETN    0.3804
+      T      0.3758
+      TNB    0.3708
+      TN     0.3680
+      E      0.3287
+      EB     0.3267
+      ENB    0.3149
+      NB     0.3058
+      EN     0.2973
+      N      0.2636
+      B      0.1638
+\end{verbatim}
+(maybe don't report all these numbers!)
+
+point: E+T $>$ T, E+TB $>$ TB, E+TP $>$ TP; information is available in
+syntax that is not captured in the other measures.
+
+point: syntax is *not* just an expensive way to get at n-grams;
+N-grams are not the same -- TP $>$ TP+N, but TP $<$ TP+E
 

### 2009-05-09 mtjournal.tex
comclusion from outline

@@
 
+\section{Conclusion}
+
+expected dependency pair matching gets at something new
+
+troubling area: speed of analysis -- current implementation requires
+running a complete n-best parser
+
+possible use: as a late-pass evaluation, to identify how systems
+perform overall
+
+future work:  explore ways to get at syntactic information without
+the expense of a full parser:
+\begin{itemize}
+\item this work moves from an LFG parser (Owczarzak) to the
+  substantially simpler and more robust Charniak system - keep going
+  and look at simpler partial-parsing approaches (forests?)
+\item conversely: how important is it that the parses be of
+  high-quality?
+\end{itemize}
+
 \section{References in the THEBIBLIOGRAPHY Environment}

### 2009-05-09 mtjournal.tex
no date included, a single reference now in place, dropped inline bibitems

@@
 \runningtitle{Expected Dependency Pair Match}
-\date{May 10, 2009}
+%s\date{May 10, 2009}
 
@@
 \item popular current
-  measures BLEU, TER
+  measures BLEU \cite{papineni02bleu}, TER
 
@@
 
-\section{References in the THEBIBLIOGRAPHY Environment}
-The first reference in the list below has the key \{BrownAndBurton\}.
-Together with the bibitem option $\backslash$citeauthoryear\{Brown and
-Burton\}\{1978\} the follwing cross-references can be used:\\
-$\backslash$cite\{BrownAndBurton\} produces: \cite{BrownAndBurton}\\
-$\backslash$shortcite\{BrownAndBurton\}  produces: \shortcite{BrownAndBurton}\\  
-$\backslash$citeauthor\{BrownAndBurton\}  produces: \citeauthor{BrownAndBurton}\\  
-$\backslash$citeyear\{BrownAndBurton\}  produces: \citeyear{BrownAndBurton}\\
-
 \appendix
 
-And this is my Appendix.
+[don't think we want an appendix]
 
 \acknowledgements
-And this is an acknowledgements section with a heading that was produced by the
-$\backslash$acknowledgements command. Thank you all for helping me writing this
-\LaTeX\ sample file.
 
-% The endnotes section will be placed here.
+[grant numbers for the DARPA/GALE grants here]
 
-\theendnotes
+% The endnotes section will be placed here.  But I don't think we have any
 
-\begin{thebibliography}{}
+% \theendnotes
 
-\bibitem[\protect\citeauthoryear{Brown and Burton}{1978}]{BrownAndBurton}
-J.~S. Brown and R.~R. Burton.
-\newblock {Diagnostic Models for Procedural Bugs in Basic Mathematical Skills}.
-\newblock {\em Cognitive Science}, 2(2):155--192, 1978.
+\bibliographystyle{klunamed}
+\bibliography{edpm-paper}
 
-\bibitem[\protect\citeauthoryear{Buchanan}{1984}]{BuchananRBES}
-Bruce~G. Buchanan and Edward~H. Shortliffe.
-\newblock {\em Rule-Based Expert Systems: The MYCIN Experiments of the Stanford
-  Heuristic Programming Project}.
-\newblock Addison-Wesley Publishing Company, 1984.
-
-\bibitem[\protect\citeauthoryear{Bunt}{1990}]{Bunt}
-H.~C. Bunt.
-\newblock {Modular Incremental Modelling of Belief and Intention}.
-\newblock In {\em Proceedings of the Second International Workshop on User
-  Modeling}, 1990.
-
-\bibitem[\protect\citeauthoryear{Cahour}{1990}]{Cahour-90-Montreal}
-B\'eatrice Cahour.
-\newblock {Competence Modelling in Consultation Dialogs}.
-\newblock In L.~Berlinguet and D.~Berthelette, editors, {\em Proceedings of the
-  International Congress, Work With Dispay Units' 89}, Montreal, Canada,
-  September 1990. North Holland, Amsterdam.
-
-\bibitem[\protect\citeauthoryear{Cahour}{1988}]{Cahour-thesis}
-B\'eatrice Cahour.
-\newblock {\em {La Mod\'elisation de l'Interlocuteur: Elaboration du Mod\`ele
-  et Effets au Cours de Dialogues de Consultation}}.
-\newblock PhD thesis, Universit\'e Paris 8, France, 1991.
-\newblock Cognitive psychology PhD.
-
-\bibitem[\protect\citeauthoryear{Carberry}{1988}]{CarberryCL88}
-Sandra~M. Carberry.
-\newblock {Modeling the User's Plans and Goals}.
-\newblock {\em Computational Linguistics}, 14(3):23--37, September 1988.
-
-\bibitem[\protect\citeauthoryear{Carbonell}{1970}]{CarbonellScholar}
-J.~R. Carbonell.
-\newblock {AI in CAI: An Artificial Intelligence Approach to Computer-Aided
-  Instruction}.
-\newblock {\em IEEE Transactions on Man-Machine Systems}, 11:190--202, 1970.
-
-\bibitem[\protect\citeauthoryear{Carr and Goldstein}{1977}]{Carr-Goldstein}
-B.~Carr and I.~Goldstein.
-\newblock {Overlays: A Theory of Modelling for Computed Aided Instruction}.
-\newblock AI Memo 406, 1977.
-
-\bibitem[\protect\citeauthoryear{Cawsey}{(in press)}]{CawseyIJMMS}
-Alison Cawsey.
-\newblock {Planning Interactive Explanations}.
-\newblock {\em International Journal of Man-Machine Studies}, in press.
-
-\bibitem[\protect\citeauthoryear{Chandrasekaran and Swartout}{1991}]{chandra-swartout}
-B.~Chandrasekaran and William Swartout.
-\newblock {Explanations in Knowledge Systems: The Role of Explicit
-  Representation of Design Knowledge}.
-\newblock {\em IEEE Expert}, 6(3):47--50, June 1991.
-
-\bibitem[\protect\citeauthoryear{Chappel and Cahour}{1991}]{MMI2-d3}
-Helen Chappel and B\'eatrice Cahour.
-\newblock {User Modeling for Multi-Modal Co-Operative Dialogue with KBS}.
-\newblock Deliverable D3, Esprit Project P2474, 1991.
-
-\bibitem[\protect\citeauthoryear{Chin}{1987}]{ChinThesis}
-David~N. Chin.
-\newblock {\em Intelligent Agents as a Basis for Natural Language Interfaces}.
-\newblock PhD thesis, University of California at Berkeley, 1987.
-
-\bibitem[\protect\citeauthoryear{Chin}{1989}]{Chin88-book}
-David~N. Chin.
-\newblock {KNOME}: {M}odeling {W}hat the {U}ser {K}nows in {UC}.
-\newblock In Alfred Kobsa and Wolfgang Wahlster, editors, {\em User Models in
-  Dialog Systems}. Springer-Verlag, Symbolic Computation Series, Berlin
-  Heidelberg New York Tokyo, 1989.
-
-\bibitem[\protect\citeauthoryear{Cohen and Jones}{1989}]{Cohen-Jones88-book}
-Robin Cohen and Marlene Jones.
-\newblock {Incorporating User Models into Expert Systems for Educational
-  Diagnosis}.
-\newblock In Alfred Kobsa and Wolfgang Wahlster, editors, {\em User Models in
-  Dialog Systems}, pages 35 -- 51. Springer-Verlag, Symbolic Computation
-  Series, Berlin Heidelberg New York Tokyo, 1989.
-
-\bibitem[\protect\citeauthoryear{Falzon}{1987}]{Falzon87}
-Pierre Falzon.
-\newblock {Les Dialogues de Diagnostic: L'\'evaluation des Connaissances de
-  l'Interlocuteur}.
-\newblock Technical Report 747, INRIA, Rocquencourt, France, 1987.
-
-\bibitem[\protect\citeauthoryear{Finin et al.}{1986}]{FininJoshiWebber}
-Timothy~W. Finin, Aravind~K. Joshi, and Bonnie~Lynn Webber.
-\newblock {Natural Language Interactions with Artificial Experts}.
-\newblock {\em Proceedings of the IEEE}, 74(7), July 1986.
-
-\end{thebibliography}
 \end{article}

### 2009-05-09 mtjournal.tex
minor title tweak
add macros for strikeout
introduction and some limited background

@@
 \newdisplay{guess}{Conjecture}
+
+% package provides \sout{} strikeout macro -- remove this before submission
+\usepackage{ulem}
 
@@
 \title{Expected Dependency Pair Match:\\
-Predicting improvements to {HTER} with expected syntactic structure} 
+Predicting {HTER} improvements with expected syntactic structure} 
 \author{Jeremy G. \surname{Kahn}\email{jgk@u.washington.edu}}  
@@
 
+% override ulem 's default change of \emph{} -- remove before ship
+\normalem
 \section{Introduction}
@@
 \begin{itemize}
-\item evaluation [esp. automatic evaluation]
-\item popular current
-  measures BLEU \cite{papineni02bleu}, TER
+\item \sout{ evaluation [esp. automatic evaluation]}
+\item \sout{popular current
+  measures BLEU \cite{papineni02bleu}, TER}
+\end{itemize}
+Machine translation (MT) evaluation is a challenge for research
+because the space of good translations is large, and two equally good
+translations may appear to be quite different at first glance. 
+%
+The challenges of choosing among translations are compounded when this
+evaluation is done automatically.
+%
+Human evaluation, however, is both time-consuming and difficult, so
+research has turned increasingly towards automatic measures of
+translation quality, usually by comparing the system translation to
+one or more reference (human) translations.
+%
+Automatic measures of this kind (e.g.\ BLEU \cite{papineni02bleu}) not
+only provide a well-defined evaluation standard but are also required
+for training on error criteria, e.g.\ with minimum error rate training
+\cite{och03mert}.
 
-\item challenges - motivate search for better measures
+The most popular evaluation measures are currently BLEU (based on
+$n$-gram precision) and the edit-distance measure Translation Edit
+Rate (TER) \cite{snover06ter}.  Recent research has found that these
+measures may not accurately track translation quality both empirically
+\cite{charniak03syntaxlmmt} and theoretically
+\cite{callisonburch06bleuproblems}.
 
-\item other extensions:
 \begin{itemize}
-  \item METEOR 
+\item \sout{challenges - motivate search for better measures}
+\item \sout{other extensions:}
   \begin{itemize}
-    \item stemming and synonym tables
-    \item tuned but not generally     tuneable
+  \item \sout{METEOR }
+    \begin{itemize}
+    \item \sout{stemming and synonym tables}
+    \item \sout{tuned but not generally     tuneable}
     \end{itemize}
-  \item TERp
+  \item \sout{TERp}
     \begin{itemize}
-    \item development-set tuning by user
-    \item stemming and synonym tables
-    \item paraphrase tables
+    \item \sout{development-set tuning by user}
+    \item \sout{stemming and synonym tables}
+    \item \sout{paraphrase tables}
     \end{itemize}
   \end{itemize}
+\end{itemize}
+
+These challenges have motivated a search for better measures that
+incorporate additional language knowledge sources.  METEOR
+\cite{banerjee05meteor}, for example, uses synonym tables and
+morphological stemming to do progressively more-forgiving matching.
+It can be tuned towards recall or precision, but is generally not
+tuned by users.  TERp \cite{snover09terp} is an extension of the
+previously-mentioned TER that also incorporates synonym sets, along
+with automatically-derived paraphrase tables.  TERp is explicitly
+intended to be tuned to a development set by users.
+
+\begin{itemize}
 \item prior work with (partial) syntactic locality:

### 2009-05-09 mtjournal.tex
include a paragraph about syntax-based measures

@@
 \begin{itemize}
-\item prior work with (partial) syntactic locality:
+\item \sout{prior work with (partial) syntactic locality:}
   \begin{itemize}
-  \item Liu \& Gildea 
-  \item Roark et al. Sparseval
-  \item Owczarzak et al. 2007
+  \item \sout{Liu \& Gildea }
+  \item \sout{Roark et al. Sparseval}
+  \item \sout{Owczarzak et al. 2007}
   \end{itemize}
+\end{itemize}
+
+As an alternative to these synonym- and paraphrase-based approaches,
+other metrics model syntactically-local (rather than string-local)
+word-sequences. \cite{liu05syntaxformteval} compared tree-local
+$n$-gram precision in various configurations of constituency and
+dependency trees.  The dependent-based SParseval measure
+\cite{roark06:sparseval}, originally designed as a parse-quality
+metric for speech, is a similar approach, in that it is an F-measure
+over a decomposition of reference and hypothesis trees.
+\cite{owczarzak07labelleddepseval}'s \textbf{d} and \textbf{d\_var}
+measures compare LFG-derived relational tuples from reference and
+hypothesis translations and report substantial improvement in
+correlation with human judgment relative to BLEU and TER.
+
+\begin{itemize}
 

### 2009-05-10 mtjournal.tex
last paragraph about syntax-oriented measures

@@
 
+These syntactically-oriented measures require a system for proposing
+dependency structure over the reference and hypothesis
+translations. Some \cite{liu05syntaxformteval,roark06:sparseval} use
+PCFG parsers with deterministic head-finding, while others
+\cite{owczarzak07labelleddepseval}extract the semantic dependency
+relations from an LFG parser \cite{cahill04lfg}.
+%
+This work extends the dependency-scoring strategies of
+\cite{roark06:sparseval,owczarzak07labelleddepseval} using a
+widely-used and publically available PCFG parser and deterministic
+head-finding rules.
+
 \begin{itemize}
-
 \item Evaluation of evaluation measures

### 2009-05-10 mtjournal.tex
paragraph regarding HTER measure

@@
 \begin{itemize}
-\item Evaluation of evaluation measures
+\item \sout{Evaluation of evaluation measures}
   \begin{itemize}
-  \item correlation vs. human fluency/adequacy 
-  \item correlation vs. human-in-the-loop measures (HTER)
+  \item \sout{correlation vs. human fluency/adequacy }
+  \item \sout{correlation vs. human-in-the-loop measures (HTER)}
   \end{itemize}
+\end{itemize}
+
+We may evaluate automatic MT measures in a variety of ways. Some
+\cite{banerjee05meteor,liu05syntaxformteval,owczarzak07labelleddepseval}
+have evaluated their success by comparing the measure to human
+judgments of fluency and adequacy.  In other work,
+e.g. \newcite{snover06ter}, measures are evaluated by comparison to
+human-targeted TER (HTER), a distance to a human-revised reference
+that uses wording closer to the MT system choices (keeping the
+original meaning) that is intended to measure the post-editing work
+required after translation.  In this paper, we pursue both kinds of
+evaluation.
+
+\begin{itemize}
 \item one challenge of metric-evaluation: underlying doc difficulty 

### 2009-05-10 mtjournal.tex
para on mean-subtraction

@@
 judgments of fluency and adequacy.  In other work,
-e.g. \newcite{snover06ter}, measures are evaluated by comparison to
+e.g. \cite{snover06ter}, measures are evaluated by comparison to
 human-targeted TER (HTER), a distance to a human-revised reference
@@
 original meaning) that is intended to measure the post-editing work
-required after translation.  In this paper, we pursue both kinds of
+required after translation.  In this paper, we explore both kinds of
 evaluation.
 
+\textbf{JGK: I think the next paragraph belongs elsewhere}
 \begin{itemize}
-\item one challenge of metric-evaluation: underlying doc difficulty 
+\item \sout{one challenge of metric-evaluation: underlying doc difficulty }
   \begin{itemize}
-  \item describe mean-subtraction
+  \item \sout{describe mean-subtraction}
   \end{itemize}
 \end{itemize}
+When considering a candidate measure $m$, reporting correlations of
+$m$ with absolute HTER scores can conflate $m$'s power in identifying
+which of two candidate translations is better with $m$'s (and HTER)'s
+ability to distinguish which of two source documents is more difficult
+to translate. To avoid this conflation, we consider the mean-removed
+$m$ and HTER scores; this lowers the correlations but ensures that the
+reported correlations are among differences in the translations rather
+than among differences in the underlying document.
 

### 2009-05-10 mtjournal.tex
moved some pieces from Experimental Paradigm into new Approach section per MO suggestion

@@
 
+OVERVIEW OF THE REST OF THE PAPER
+
 \textbf{JGK: I think the next paragraph belongs elsewhere}
@@
 
+\section{Approach}
+
+Outline of big idea here.
+
+\begin{itemize}
+\item comparison of various bags of components of parse tree
+  [ Figure describing Sparseval-ish variant ]
+
+  Like sparseval: basic unit is labeled dependency pair
+
+  Note: inbound links \& full pairs "overweight" head words; this is
+  a good thing.
+
+  [ figure demonstrating F-measure over d+l+h decomposition of toy
+  trees ]
+
+\item Tree features may be extracted from a 1-best tree, or from an
+  n-best tree list
+
+\item elements may be combined in various ways: single-bag F, combination of p,r
+  measures (evenly, on analogy with BLEU)... or we might weight
+  different combinations
+
+\end{itemize}
 \section{Experimental paradigm}
 
-Section explains space of measures to explore
+Section explains space of measures to explore, mean-normalization
 
 \begin{itemize}
-\item F-measure over various bags of components of parse tree
-
-  [ Figure describing Sparseval-ish variant ]
-
 \item Listing possible component valences
@@
   \end{itemize}
-
-  Note: inbound links \& full pairs "overweight" head words; this is
-  a good thing.
-
-  [ figure demonstrating F-measure over d+l+h decomposition of toy
-  trees ]
 

### 2009-05-10 mtjournal.tex
moved section on mean-removal to paradigm, added labels

@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 
-\textbf{JGK: I think the next paragraph belongs elsewhere}
-\begin{itemize}
-\item \sout{one challenge of metric-evaluation: underlying doc difficulty }
-  \begin{itemize}
-  \item \sout{describe mean-subtraction}
-  \end{itemize}
-\end{itemize}
-When considering a candidate measure $m$, reporting correlations of
-$m$ with absolute HTER scores can conflate $m$'s power in identifying
-which of two candidate translations is better with $m$'s (and HTER)'s
-ability to distinguish which of two source documents is more difficult
-to translate. To avoid this conflation, we consider the mean-removed
-$m$ and HTER scores; this lowers the correlations but ensures that the
-reported correlations are among differences in the translations rather
-than among differences in the underlying document.
-
 \section{Approach}
+\label{sec:approach}
 
@@
 \section{Experimental paradigm}
+\label{sec:paradigm}
 
@@
   \end{itemize}
+
+\begin{itemize}
+\item \sout{one challenge of metric-evaluation: underlying doc difficulty }
+  \begin{itemize}
+  \item \sout{describe mean-subtraction}
+  \end{itemize}
+\end{itemize}
+When considering a candidate measure $m$, reporting correlations of
+$m$ with absolute HTER scores can conflate $m$'s power in identifying
+which of two candidate translations is better with $m$'s (and HTER)'s
+ability to distinguish which of two source documents is more difficult
+to translate. To avoid this conflation, we consider the mean-removed
+$m$ and HTER scores; this lowers the correlations but ensures that the
+reported correlations are among differences in the translations rather
+than among differences in the underlying document.
+
 \end{itemize}
@@
 \section{Correlation with human judgments of fluency \& adequacy}
-
+\label{sec:faexpts}
 Corpus: LDC Multiple Translation Chinese corpus 3\&4 treat each
@@
 \section{Correlations with HTER}
-
+\label{sec:hter1}
   Corpus
@@
 \section{Weight-tuning to combine syntax and other knowledge sources}
+\label{sec:hter2}
 
@@
 \section{Conclusion}
-
+\label{sec:conclusion}
 expected dependency pair matching gets at something new

### 2009-05-10 mtjournal.tex
multiple changes outlining basics of approach section

@@
 \usepackage{ulem}
+
+\usepackage{graphicx}
+
+\newcommand{\arclabel}[1]{\ensuremath{\stackrel{#1}{\to}}}
+\newcommand{\precision}[1]{\ensuremath{\textrm{Prec}\left(#1\right)}}
+\newcommand{\recall}[1]{\ensuremath{\textrm{Recall}\left(#1\right)}}
 
@@
 
+TO DO: DISCUSS TUNING'S VIRTUES, FLAWS
+
 \begin{itemize}
@@
 
-OVERVIEW OF THE REST OF THE PAPER
+TO DO: OVERVIEW OF THE REST OF THE PAPER
 
@@
 \label{sec:approach}
+
+This work explores a family of dependency pair match (DPM) measures
+that are composed of precision and recall combinations over various
+decompositions of a syntactic dependency tree. These measures are
+extensions of the dependency-pair F measures found in
+\cite{roark06:sparseval,owczarzak07labelleddepseval}.  Rather than
+comparing string sequences, as BLEU does with its $n$-gram precision,
+this approach defers to a parser for an indication of the relevant
+words that indicate more about its meaning --- in these
+implementations, the head on which that word depends.  Each sentence
+(both reference and hypothesis) is converted to a labeled syntactic
+dependency tree and then relations from each tree are extracted and
+compared.
+
+A member of this family is defined by several parameters.  The first
+of these parameters is the nature of the decomposition of the
+dependency tree structure.  A \emph{decomposition list} is the list of
+ways in which the tree is reduced to a bag of tree-local
+tuples. Figure~\ref{fig:decompexample}
+\begin{figure}
+  \centering
+  \begin{tabular}{rcc}
+    & \textbf{Reference} & \textbf{Hypothesis} \\
+    \textbf{dependency tree}
+    & \includegraphics[scale=0.6]{dpm-example-ref.eps} & 
+    \includegraphics[scale=0.6]{dpm-example-hyp.eps}\\
+    \textbf{$dlh$ list} &
+    \begin{tabular}{@{$\langle$}c@{,~}c@{,~}c@{$\rangle$}}
+      the & \arclabel{\texttt{det}}  &  cat \\
+      red & \arclabel{\texttt{mod}}  &  cat \\
+      cat & \arclabel{\texttt{subj}} &  ate \\
+      ate & \arclabel{\texttt{root}} &  $<$root$>$ \\
+    \end{tabular} & 
+    \begin{tabular}{@{$\langle$}c@{,~}c@{,~}c@{$\rangle$}}
+      the &      \arclabel{\texttt{det}}  &  cat \\
+      cat &      \arclabel{\texttt{subj}} &  stumbled \\
+      stumbled & \arclabel{\texttt{root}} &  $<$root$>$ \\
+    \end{tabular}\\
+  \end{tabular}\\
+  Precision$_{dlh}$ is $\frac{1}{3}$ and Recall$_{dlh}$ is
+  $\frac{1}{4}$, so $F[dlh] = \frac{2}{7}$
+  \caption{Example hypothesis and reference dependency trees and the $dlh$ decomposition of each.}
+  \label{fig:decompexample}
+\end{figure}
+illustrates the \emph{dependency-link-head} decomposition of the
+dependency tree into a list of $\langle d, l, h \rangle$ tuples.  Some
+members of the DPM family may apply more than one decomposition; other
+good examples are the $dl$ decomposition, which generates a bag of
+dependent word with outbound links, and the $lh$ decomposition, which
+generates a bag of inbound link labels, with the head word for each
+included.
+
+It is worth noting here that the $dlh$ and $lh$ decompositions (but
+not the $dl$ decomposition) ``overweight'' the headwords, in that
+there are $n$ elements in the resulting bag, but if a word has no
+dependents it is found in the resulting bag exactly one time (in the
+$dlh$ case) or not at all (in the $lh$ case).  Conversely,
+syntactically ``key'' words, that are directly modified by many other
+words in the tree, are included multiple times in the decomposition
+(once for each inbound link).  We argue that this overcounting is a
+virtue; the syntactic structure indicates which words are more
+important to translate correctly.
+
+We may not completely trust the parser's best parse.  The parser
+itself, if we use a probabilistic parser, can provide an $n$-best
+parse list forest for the translation reference and translation
+hypothesis. We use the probability statistics of the forest to compute
+expected counts for each decomposition. Though this approach yields
+partial counts, standard comparisons like precision and recall are
+still valid.
+
+When multiple decomposition types used together, we may combine these
+subscores in a variety of ways.  We may compute precision and recall
+subscores for each decomposition separately, or, since the results of
+each decomposition are of different types entirely, we may compute
+them as members of one large bag for an even simpler F score.  These
+two approaches are equivalent when only one decomposition type is
+included.  For simplicity in presentation, we use the following
+notation, where $dl$ and $lh$ represent the two kinds of
+decompositions described above and $\mu_h$ represents a harmonic mean:
+\begin{eqnarray}
+  \label{eq:fprmeans}
+  F[dl,lh] & = &
+  \mu_h \left( \precision{dl \cup lh},
+    \recall{dl \cup lh} \right) \\
+  \mu_{PR}[dl,lh]  & = & \mu_h \left( \precision{dl},
+    \recall{dl}, \precision{lh}, \recall{lh} \right)    
+\end{eqnarray}
+Dependency-based SParseval and Owczarzak \textit{et al.}'s
+\cite*{owczarzak07labelleddepseval} \textbf{d} approach, then, may
+each be understood as $F[dlh]$, while the latter's \textbf{d\_var}
+method may be understood as something close to $F[dl,lh]$.
 

### 2009-05-10 mtjournal.tex
conclusion to approach section

@@
 method may be understood as something close to $F[dl,lh]$.
+
+Both of the combination methods above ($F[\cdot]$ and
+$\mu_{PR}[\cdot]$) are ``naive'' combinations, in that they treat each
+decomposition as equivalent to the next.  One further direction to
+explore involves tuning (presumably on held-out data) the precision
+and recall weights for each decomposition type for a later linear
+combination.
+
+The possible family of metrics outlined above is quite large. In the
+next section, we make explicit the range of these parameters that we
+explore in this article.
 

### 2009-05-10 mtjournal.tex
removed outline text that now is actual content

@@
 
-\begin{itemize}
-\item \sout{ evaluation [esp. automatic evaluation]}
-\item \sout{popular current
-  measures BLEU \cite{papineni02bleu}, TER}
-\end{itemize}
 Machine translation (MT) evaluation is a challenge for research
@@
 
-\begin{itemize}
-\item \sout{challenges - motivate search for better measures}
-\item \sout{other extensions:}
-  \begin{itemize}
-  \item \sout{METEOR }
-    \begin{itemize}
-    \item \sout{stemming and synonym tables}
-    \item \sout{tuned but not generally     tuneable}
-    \end{itemize}
-  \item \sout{TERp}
-    \begin{itemize}
-    \item \sout{development-set tuning by user}
-    \item \sout{stemming and synonym tables}
-    \item \sout{paraphrase tables}
-    \end{itemize}
-  \end{itemize}
-\end{itemize}
-
 These challenges have motivated a search for better measures that
@@
 TO DO: DISCUSS TUNING'S VIRTUES, FLAWS
-
-\begin{itemize}
-\item \sout{prior work with (partial) syntactic locality:}
-  \begin{itemize}
-  \item \sout{Liu \& Gildea }
-  \item \sout{Roark et al. Sparseval}
-  \item \sout{Owczarzak et al. 2007}
-  \end{itemize}
-\end{itemize}
 
@@
 head-finding rules.
-
-\begin{itemize}
-\item \sout{Evaluation of evaluation measures}
-  \begin{itemize}
-  \item \sout{correlation vs. human fluency/adequacy }
-  \item \sout{correlation vs. human-in-the-loop measures (HTER)}
-  \end{itemize}
-\end{itemize}
 
@@
 
-Outline of big idea here.
-
-\begin{itemize}
-\item comparison of various bags of components of parse tree
-  [ Figure describing Sparseval-ish variant ]
-
-  Like sparseval: basic unit is labeled dependency pair
-
-  Note: inbound links \& full pairs "overweight" head words; this is
-  a good thing.
-
-  [ figure demonstrating F-measure over d+l+h decomposition of toy
-  trees ]
-
-\item Tree features may be extracted from a 1-best tree, or from an
-  n-best tree list
-
-\item elements may be combined in various ways: single-bag F, combination of p,r
-  measures (evenly, on analogy with BLEU)... or we might weight
-  different combinations
-
-\end{itemize}
 \section{Experimental paradigm}

### 2009-05-10 mtjournal.tex
improved figure proportions
new figure re: dl,lh decomposition, also minor cleanups

@@
 
-% package provides \sout{} strikeout macro -- remove this before submission
+% TO DO: package provides \sout{} strikeout macro -- remove this
+% before submission
 \usepackage{ulem}
@@
     & \textbf{Reference} & \textbf{Hypothesis} \\
-    \textbf{dependency tree}
-    & \includegraphics[scale=0.6]{dpm-example-ref.eps} & 
-    \includegraphics[scale=0.6]{dpm-example-hyp.eps}\\
+    \textbf{tree}
+    & \includegraphics[scale=0.5]{dpm-example-ref.eps} & 
+    \includegraphics[scale=0.5]{dpm-example-hyp.eps}\\
     \textbf{$dlh$ list} &
@@
   Precision$_{dlh}$ is $\frac{1}{3}$ and Recall$_{dlh}$ is
-  $\frac{1}{4}$, so $F[dlh] = \frac{2}{7}$
+  $\frac{1}{4}$: thus $F[dlh] = \frac{2}{7}$
   \caption{Example hypothesis and reference dependency trees and the $dlh$ decomposition of each.}
@@
 \end{figure}
-illustrates the \emph{dependency-link-head} decomposition of the
+illustrates the \emph{dependency-link-head} decomposition of a toy
 dependency tree into a list of $\langle d, l, h \rangle$ tuples.  Some
@@
 included. 
+\begin{figure}
+  \centering
+    \begin{tabular}{cc}
+    $dl$ & $lh$ \\
+    \hline
+    \begin{tabular}{@{$\langle$}c@{,~}c@{$\rangle$}}
+      the & \arclabel{\texttt{det}}  \\
+      cat & \arclabel{\texttt{subj}} \\
+      stumbled & \arclabel{\texttt{root}} \\
+    \end{tabular} &
+    \begin{tabular}{@{$\langle$}c@{,~}c@{$\rangle$}}
+      \arclabel{\texttt{det} } &  cat \\
+      \arclabel{\texttt{subj}} &  stumbled \\
+      \arclabel{\texttt{root}} &  $<$root$>$ \\
+    \end{tabular}\\
+  \end{tabular}
+  \caption{The $dl$ and $lh$ decompositions of the hypothesis tree in
+    figure~\ref{fig:decompexample}. The items extracted here
+    are individually less restrictive in their ability to match
+    against reference tuples. 
+%     $\precision{dl} = \frac{2}{3}$,
+%     $\recall{dl} = \frac{2}{4}$, $\precision{lh} = \frac{2}{3}$,
+%     $\recall{lh} = \frac{2}{4}$, so $F[dl,lh] = \frac{4}{7}$
+  }
+  \label{fig:decompexample2}
+\end{figure}
+Figure~\ref{fig:decompexample2} demonstrates the $dl,lh$
+decomposition for the hypothesis tree in figure~\ref{fig:decompexample}.
 

### 2009-05-10 mtjournal.tex
new macros for tree diagrams
beginning of the paradigm section (with implementation details)

@@
 \usepackage{graphicx}
+\usepackage{qtree}
 
+%\newcommand{\headlabel}[2]{\ensuremath{\underset{\textrm{#2}}{\textrm{#1}}}}
+\newcommand{\headlabel}[2]{
+  \begin{tabular}{c} #1\\
+    {\small #2} \end{tabular}}
 \newcommand{\arclabel}[1]{\ensuremath{\stackrel{#1}{\to}}}
+
 \newcommand{\precision}[1]{\ensuremath{\textrm{Prec}\left(#1\right)}}
@@
 partial counts, standard comparisons like precision and recall are
-still valid.
+still valid, and using the expectations can help cope with both error
+in the parser and ambiguity in the translations (reference and
+hypothesis).
 
@@
 
+In principle, the family of DPM measures may be implemented with
+any parser that generates a dependency graph (a single labelled arc
+for each word, pointing to its head-word). Prior work
+\cite{owczarzak07labelleddepseval} on related
+measures has used an LFG parser \cite{cahill04lfg} or
+an unlabelled dependency tree \cite{liu05syntaxformteval}.
+
+%
+In this work, we use a state-of-the-art PCFG (the first stage of
+\cite{charniak-johnson:2005:ACL}) and context-free head-finding rules
+\cite{magerman95headfinding} to generate a 50-best list of dependency
+trees for each hypothesis and reference translation.  We use the
+parser's default (English-language) Wall Street Journal training
+parameters.  Head-finding uses the Charniak parser's rules, with three
+modifications: prepositional and complementizer phrases choose nominal
+and verbal heads respectively (rather than functional heads) and
+auxiliary verbs are modifiers of main verbs (rather than the
+converse).
+
+Having constructed the dependency tree, we label the arcs as $d
+\stackrel{A/B}{\to} h$, where the arc label $A/B$ between dependent
+$d$ and its head $h$ is composed of $A$ (the lowest constituent-label
+headed by $h$ and dominating $d$) and $B$ (the highest constituent
+label headed by $d$). 
+\begin{figure}
+  % \centering
+  \Tree
+  [.\headlabel{S}{stumbled}
+    [.\headlabel{NP}{cat}
+      [.\headlabel{DT}{the} the ]
+      [.\headlabel{NN}{cat} cat ]
+    ]
+    [.\headlabel{VP}{stumbled} 
+      [.\headlabel{VBD}{stumbled} stumbled ]
+    ]
+  ]
+  \\
+  \begin{center}
+    \includegraphics[scale=0.6]{dpm-example-depextract.eps}
+  \end{center}
+  \caption{An example constituent tree (heads of each constituent are
+    listed small below the label) and the labelled dependency tree
+    derived from it using the strategy described in
+    section~\ref{sec:paradigm}.}
+  \label{fig:depextract}
+\end{figure}
+For example, in figure~\ref{fig:depextract}, the S node is the
+lowest node headed by \emph{stumbled} that dominates \emph{cat}, and
+the NP node is the highest constituent label headed by \emph{cat}, so
+the arc between \emph{cat} and \emph{stumbled} is labelled \arclabel{S/NP}.
+%
+This strategy is very similar to one adopted in the reference
+implementation of labelled-dependency \textsc{SParseval}, and may be
+considered as an approximation of the rich semantics generated by
+\cite{cahill04lfg} or another heavily knowledge-engineered
+parser, but with much less knowledge-engineering required.
+
+The $A/B$ labels are not as descriptive as the LFG semantics, but they
+have a similar resolution, e.g.\ the $\stackrel{S/NP}{\to}$ arc label
+usually represents a subject dependent of a sentential verb.
+
+In the experiments presented later in this article, we explore the
+space of metrics outlined above along multiple dimensions:
+\begin{description}
+\item[Decompositions]: we consider syntactic decompositions of the
+  following types.
+  \begin{description}
+  \item[dlh] $\left\langle \textrm{Dependent}, \textrm{arc Label},
+      \textrm{Head}\right\rangle$ here, every tuple includes two words
+    and the dependency-label between them.
+  \item[dl] $\left\langle \textrm{Dependent}, \textrm{arc Label}
+      \right\rangle$ the tuple is the word plus the label of the arc
+      linking it to its head; this decomposition marks how the word
+      fits into its syntactic context (what it modifies).
+    \item[lh] $\left\langle \textrm{arc Label}, \textrm{Head}
+      \right\rangle$ each tuple here is the dependency relation plus
+      the word being modified; this decomposition implicitly marks how
+      key the word is to the sentence.
+    \item[dh] $\left\langle \textrm{Dependent}, \textrm{Head}
+      \right\rangle$ here, drops the arc-label. This decomposition
+      ignores the syntactic-role information implied in the arc labels.
+    \item[1g,2g] these decompositions are syntactically-null
+      they just includes each 1-gram (or 2-gram) without any
+      syntactic labeling at all.  They are thus simple measures of
+      unigram (bigram)
+      precision and recall.
+  \end{description}
+\item[Forest size]
+Although the parser can generate considerably more, we consider forests of size 1 and of size 50
+\item[Forest confidence]
+\end{description}
+
 Section explains space of measures to explore, mean-normalization
@@
 \begin{itemize}
-\item Listing possible component valences
-  \begin{itemize}
-  \item dependent + labeled link + head 
-  \item outbound only (dependent +
-    label)
-  \item inbound only (label + head)
-  \item word-word (skipping label)
-  \end{itemize}
-  \begin{itemize}
-  \item 1-grams
-  \item 2-grams
-  \item \ldots{}
-  \end{itemize}
 

### 2009-05-10 mtjournal.tex
major changes to make the paradigm section tick over

@@
       ignores the syntactic-role information implied in the arc labels.
-    \item[1g,2g] these decompositions are syntactically-null
-      they just includes each 1-gram (or 2-gram) without any
-      syntactic labeling at all.  They are thus simple measures of
-      unigram (bigram)
-      precision and recall.
+    \item[1g,2g] these decompositions are syntactically-null.  They
+      simply include each 1-gram (or 2-gram) without any syntactic
+      labeling at all.  They are thus simple measures of unigram
+      (bigram) precision and recall.
   \end{description}
 \item[Forest size]
-Although the parser can generate considerably more, we consider forests of size 1 and of size 50
-\item[Forest confidence]
+Although the parser can generate considerably more, we consider forests of size 1 and of size 50.
+\item[Forest confidence] An expectation requires a probability distribution over the $n$-best
+list, and we consider three options: uniform, the parser
+probabilities, and a flattened version of the parser probabilities
+such that $\tilde{p}(x) = \frac{p(x)^\gamma}{\sum_ip(i)^\gamma}$
+(where $\gamma$ is a free parameter) to account for the fact that the
+parser tends to be over-confident.  In all cases, the probabilities
+are normalized to sum to one over the the $n$-best list, where the
+maximum $n$ in this work is 50.  The uniform distribution ($\gamma =
+0$) is intended to be equivalent to the
+\cite{owczarzak07labelleddepseval} \textbf{d\_50} and
+\textbf{d\_50\_var} measures.\footnote{Since
+  \cite{owczarzak07labelleddepseval} report no use of parse
+  weights, \textbf{d\_50} and \textbf{d\_50\_var} may be using a sum
+  of counts over the 50-best list rather than expected-counts over a
+  uniform distribution. These two approaches are equivalent --- so
+  long as the $n$-best list is always the same length for hypothesis
+  and reference.  In our implementation, the $n$-best list does not
+  always reach 50 candidate parses on short sentences, so the
+  expectation matches our intent better than a sum of counts over the
+  $n$-best.}
+\item[score combination]: We consider three possible methods for
+  combining scores across decompositions, as suggested in
+  section~\ref{sec:approach}:
+  \begin{description}
+  \item[$F\lbrack\cdot\rbrack$] treats each decomposition as
+    contributing $n$ more items to the bag of tokens extracted from
+    the tree, ad then computes a global $F$ score.
+  \item[$\mu_{PR}\lbrack\cdot\rbrack$] constructs a precision and
+    recall score from each decomposition, and then does an
+    evenly-weighted harmonic mean of all the component precision and
+    recall scores.
+  \item Tuned, like the previous combination, calculates precision and
+    recall separately for each decomposition, but performs a linear
+    weighted combination of the resulting precision and recall scores.
+    For this tuning, we note that TERp \cite{snover09terp} provides a
+    tuneable system over a small number of parameters (they include,
+    among others, separate cost for insertions, deletions,
+    substitutions and synonyms) and an optimization algorithm that
+    assigns weights to each parameter. We use the various
+    decompositions' precision and recall as new parameters to the TERp
+    optimizer.
+  \end{description}
 \end{description}
@@
 Section explains space of measures to explore, mean-normalization
-
-\begin{itemize}
-
-\item Tree features may be extracted from a 1-best tree, or from an
-  n-best tree list
-
-  \begin{itemize}
-  \item  counts are an expectation over the forest
-  \item n parameter: how many trees in the forest
-  \item gamma parameter for flattening overconfident parse hyps
-  \end{itemize}
-  note intentional similarities to Owczarzak paper
-
-\item Components in the basic DPM measures combined in various ways:
-
-  \begin{itemize}
-  \item combination of precision, recall scores for different
-    components (on analogy with BLEU combination), naive assumption:
-    all weighted equally
-
-  \item simple F over items of all valences (naive assumption: all
-    weighted equally)
-  \end{itemize}
-\item Tuneable weights:
-  \begin{itemize}
-  \item recall that TERp has tuneable system over small number of parameters
-    \begin{itemize}
-    \item insertion weight
-    \item deletion weight
-    \item substitution weight
-    \item synonym weight (motivated with WordNet)
-    \end{itemize}
-  \item learn weights for separate precision, recall for various
-    components
-  \end{itemize}
-
-\item Implementation
-  \begin{itemize}
-  \item Charniak parser
-  \item n=1..50
-  \item head-finding, with some modifications
-  \item label-construction
-  \end{itemize}
 
@@
 than among differences in the underlying document.
-
-\end{itemize}
 

### 2009-05-10 mtjournal.tex
to-do items for paradigm section

@@
 
-Section explains space of measures to explore, mean-normalization
+\subsection{Reference evaluation measures}
 
-\begin{itemize}
-\item \sout{one challenge of metric-evaluation: underlying doc difficulty }
-  \begin{itemize}
-  \item \sout{describe mean-subtraction}
-  \end{itemize}
-\end{itemize}
+TO DO: TALK ABOUT HUMAN F/A JUDGMENTS
+
 When considering a candidate measure $m$, reporting correlations of
@@
 than among differences in the underlying document.
+
+TO DO: MENTION DELTAS IN A FOOTNOTE, AND CITE MetricsMATR SYSTEM PAPER
 

### 2009-05-10 mtjournal.tex
whitespace cleanup

--- 2009-05-10 mtjournal.tex
@@
       ignores the syntactic-role information implied in the arc labels.
-    \item[1g,2g] these decompositions are syntactically-null
-      they just includes each 1-gram (or 2-gram) without any
-      syntactic labeling at all.  They are thus simple measures of
-      unigram (bigram)
-      precision and recall.
+    \item[1g,2g] these decompositions are syntactically-null.  They
+      simply include each 1-gram (or 2-gram) without any syntactic
+      labeling at all.  They are thus simple measures of unigram
+      (bigram) precision and recall.
   \end{description}
-\item[Forest size]
-Although the parser can generate considerably more, we consider forests of size 1 and of size 50
-\item[Forest confidence]
+\item[Forest size] Although the parser can generate considerably more,
+  we consider forests of size 1 and of size 50.
+\item[Forest confidence] An expectation requires a probability
+  distribution over the $n$-best list, and we consider three options:
+  uniform, the parser probabilities, and a flattened version of the
+  parser probabilities such that $\tilde{p}(x) =
+  \frac{p(x)^\gamma}{\sum_ip(i)^\gamma}$ (where $\gamma$ is a free
+  parameter) to account for the fact that the parser tends to be
+  over-confident.  In all cases, the probabilities are normalized to
+  sum to one over the the $n$-best list, where the maximum $n$ in this
+  work is 50.  The uniform distribution ($\gamma = 0$) is intended to
+  be equivalent to the \cite{owczarzak07labelleddepseval}
+  \textbf{d\_50} and \textbf{d\_50\_var} measures.\footnote{Since
+    \cite{owczarzak07labelleddepseval} report no use of parse weights,
+    \textbf{d\_50} and \textbf{d\_50\_var} may be using a sum of
+    counts over the 50-best list rather than expected-counts over a
+    uniform distribution. These two approaches are equivalent --- so
+    long as the $n$-best list is always the same length for hypothesis
+    and reference.  In our implementation, the $n$-best list does not
+    always reach 50 candidate parses on short sentences, so the
+    expectation matches our intent better than a sum of counts over
+    the $n$-best.}
+\item[score combination]: We consider three possible methods for
+  combining scores across decompositions, as suggested in
+  section~\ref{sec:approach}:
+  \begin{description}
+  \item[$F\lbrack\cdot\rbrack$] treats each decomposition as
+    contributing $n$ more items to the bag of tokens extracted from
+    the tree, ad then computes a global $F$ score.
+  \item[$\mu_{PR}\lbrack\cdot\rbrack$] constructs a precision and
+    recall score from each decomposition, and then does an
+    evenly-weighted harmonic mean of all the component precision and
+    recall scores.
+  \item Tuned, like the previous combination, calculates precision and
+    recall separately for each decomposition, but performs a linear
+    weighted combination of the resulting precision and recall scores.
+    For this tuning, we note that TERp \cite{snover09terp} provides a
+    tuneable system over a small number of parameters (they include,
+    among others, separate cost for insertions, deletions,
+    substitutions and synonyms) and an optimization algorithm that
+    assigns weights to each parameter. We use the various
+    decompositions' precision and recall as new parameters to the TERp
+    optimizer.
+  \end{description}
 \end{description}
 
-Section explains space of measures to explore, mean-normalization
+\subsection{Reference evaluation measures}
 
-\begin{itemize}
+TO DO: TALK ABOUT HUMAN F/A JUDGMENTS
 
-\item Tree features may be extracted from a 1-best tree, or from an
-  n-best tree list
-
-  \begin{itemize}
-  \item  counts are an expectation over the forest
-  \item n parameter: how many trees in the forest
-  \item gamma parameter for flattening overconfident parse hyps
-  \end{itemize}
-  note intentional similarities to Owczarzak paper
-
-\item Components in the basic DPM measures combined in various ways:
-
-  \begin{itemize}
-  \item combination of precision, recall scores for different
-    components (on analogy with BLEU combination), naive assumption:
-    all weighted equally
-
-  \item simple F over items of all valences (naive assumption: all
-    weighted equally)
-  \end{itemize}
-\item Tuneable weights:
-  \begin{itemize}
-  \item recall that TERp has tuneable system over small number of parameters
-    \begin{itemize}
-    \item insertion weight
-    \item deletion weight
-    \item substitution weight
-    \item synonym weight (motivated with WordNet)
-    \end{itemize}
-  \item learn weights for separate precision, recall for various
-    components
-  \end{itemize}
-
-\item Implementation
-  \begin{itemize}
-  \item Charniak parser
-  \item n=1..50
-  \item head-finding, with some modifications
-  \item label-construction
-  \end{itemize}
-
-\begin{itemize}
-\item \sout{one challenge of metric-evaluation: underlying doc difficulty }
-  \begin{itemize}
-  \item \sout{describe mean-subtraction}
-  \end{itemize}
-\end{itemize}
 When considering a candidate measure $m$, reporting correlations of
@@
 
-\end{itemize}
+TO DO: MENTION DELTAS IN A FOOTNOTE, AND CITE MetricsMATR SYSTEM PAPER
 

### 2009-05-10 mtjournal.tex
subsections in paradigm section

@@
 \label{sec:paradigm}
+In this section, we outline the details of the implementation and
+experimental variables explored in the experiments presented here.
 
+\subsection{Parse tree implementation}
 In principle, the family of DPM measures may be implemented with
@@
 
+\subsection{Dimensions of metric family}
 In the experiments presented later in this article, we explore the

### 2009-05-10 mtjournal.tex
added corpus introduction to F/A section

@@
 \label{sec:faexpts}
-Corpus: LDC Multiple Translation Chinese corpus 3\&4 treat each
-segment and each judgment as independent data point
+
+TO DO: INTRODUCE THIS SECTION
+
+\subsection{Corpus}
+
+For these experiments, we use LDC Multiple Translation Chinese corpus
+parts 2~\cite{LDC03MTC2} and 4~\cite{LDC06MTC4}.  These corpora
+include multiple human judgments of fluency and adequacy for each
+sentence, with each judgment using a different human judge and a
+different reference translation.  For a rough\footnote{Our segment
+  count differs from \cite{owczarzak07labelleddepseval}, who report
+  16,800 segments over the same corpus. We find baseline correlations
+  (BLEU$_4$ and TER) lower than those reported there as well, so the
+  results presented here are not directly comparable with that paper,
+  though we demonstrate similar gains over those baselines in
+  essentially the same corpus. % (section~\ref{sec:facorr:newdeps}).}
+}
+comparison with \cite{owczarzak07labelleddepseval}, we treat each
+judgment as a separate segment.
+%
+This treatment of this corpus yields 16,815 tuples of
+$\langle$hypothesis, reference, fluency, adequacy$\rangle$.  In these
+experiments, we extend this tuple with automatic scores derived from
+$\langle$hypothesis, reference$\rangle$ and examine the
+correlations\footnote{The independence of each of these segments is
+  questionable, since the same hypothesis translations are used in
+  multiple items, but for the sake of methodological comparison with
+  prior work, this strategy is preserved.}  between those automatic
+scores and the arithmetic mean of the fluency and adequacy measures.
 
 \begin{itemize}
-\item EXPT
-
-  using n=1 parse for DPM variants
+\item EXPT  using n=1 parse for DPM variants
 
@@
 \item EXPT 
-
   combining precision, recall naively vs. combining items
@@
 \item  EXPT
-
   setting gamma

### 2009-05-10 mtjournal.tex
hacked in a summary paragraph for the FA correlations, and some labels

@@
 \item EXPT  using n=1 parse for DPM variants
-
+\label{tab:facorr:subgraphs}
 \begin{verbatim}
@@
 \item EXPT
+  \label{tab:facorr:multiparse}
   expanding n, gamma=0
@@
 \item  EXPT
+\label{tab:facorr:parseweight}
   setting gamma
@@
 
+\subsection{Summary}
+
+In this section, we have presented experiments exploring a number of
+variants of the DPM metric against an average fluency/adequacy judgment. The experiments suggest a
+best-case variant, where we set:
+\begin{displaymath}
+  EDPM = F[1g,2g,dl,lh], n=50, \gamma=0.25
+\end{displaymath}
+in which we choose a $1g,2g,dl,lh$ sub-graph decomposition based on
+the improvements from better sub-graphs
+(table~\ref{tab:facorr:subgraphs}), multiple parses ($n=50$) based on
+table~\ref{tab:facorr:multiparse}, and $\gamma=0.25$, hinted at by
+table~\ref{tab:facorr:parseweight}.  We use these EDPM
+parameter-settings in the experiments exploring document-level correlations with HTER
+(section~\ref{sec:hter1}).
+
 \end{itemize}
 
-\section{Correlations with HTER}
+\section{Per-document correlations with HTER}
 \label{sec:hter1}

### 2009-05-10 mtjournal.tex
first hacky attempts to turn verbatims into tables. looks bad

@@
 
+  \begin{table}
+    % \centering
+    \caption{$dl,lh$ decompositions are
+      more correlated with the fluency-adequacy average than BLEU$_4$
+      or $dlh$ decompositions.}
+    \begin{tabular}{lr}
+      metric  &    $r$ \\
+      \hline
+      $F[dl,lh]$ &   0.226 \\
+       % # +1 smoothing
+      BLEU$_4$ &   0.218 \\
+      $F[dlh]$ &     0.185 \\
+      TER &      -0.173 \\
+    \end{tabular}
+    \label{tab:facorr:subgraphs}
+\end{table}
+
+Table~\ref{tab:facorr:subgraphs} indicates that: using
+partial-labeling (dl,lh) much better than full link. Also, using
+Charniak parser as labeler seems to work, LFG not necessary
+Here, $n=1$ and only the $F[\cdot]$ score combination is shown.
+
 \begin{itemize}
-\item EXPT  using n=1 parse for DPM variants
-\label{tab:facorr:subgraphs}
-\begin{verbatim}
-   metric      r
-  --------   -----
-  F[dl;lh]   0.226
-  BLEU4      0.218  # +1 smoothing
-  F[dlh]     0.185
-  TER       -0.173
-\end{verbatim}
-
-  point: using partial-labeling (dl,lh) much better than full
-  link. Also, using Charniak parser as labeler seems to work, LFG not
-  necessary
 

### 2009-05-10 mtjournal.tex
corpus stats for GALE 2.5 corpus

@@
 \label{sec:hter1}
-  Corpus
-    Gale translation corpus 2.5; 3 sites, 2 source languages, 4 genres each
+TO DO: INTRODUCE THIS SECTION
+
+The GALE 2.5 translation corpus is made up of system translations into
+English from three sites.
+% MO added:
+The three sites all use system combination to integrate results from
+multiple systems, some of which are phrase-based and some which may
+use syntax on either the source or target side. No system provided
+system-generated parses.
+%
+The corpus being translated comes from
+Arabic and Chinese in four genres: \texttt{bc} (broadcast
+conversation), \texttt{bn} (broadcast news), \texttt{nw} (newswire),
+and \texttt{wb} (web text), with corpus sizes shown in
+table~\ref{tab:galestats}.
+\begin{table}
+%  \centering
+  \begin{tabular}{r|rr|rr|rr}
+     & \multicolumn{2}{c|}{Arabic} & \multicolumn{2}{c|}{Chinese}
+     & \multicolumn{2}{c}{Total}\\
+     & doc & sent & doc & sent & doc   & sent\\
+     \hline
+     \texttt{bc}    & 59  & 750 & 56 & 1061 & 115 & 1811\\
+     \texttt{bn}    & 63  & 666 & 63 & 620  & 126 & 1286\\
+     \texttt{nw}    & 68  & 494 & 70 & 440  & 138 & 934 \\
+     \texttt{wb}    & 69  & 683 & 68 & 588  & 137 & 1271\\
+     \hline
+     Total & 259 & 2593& 257& 2709 & 516 & 5302\\
+  \end{tabular}
+  \caption{Corpus statistics for the GALE 2.5 translation
+    corpus.}
+  \label{tab:galestats}
+\end{table}
+The corpus includes one English reference translation $r_i$
+\cite{gale08phase2_5references} for each sentence $i$ and a system
+translation $t_{i,z}$ for each of the three systems
+$z$. Additionally, each of the system translations of each segment $i$
+has a corresponding human-targeted reference aligned at the sentence
+level, so we have available the HTER score of each segment
+$s_{\textsc{hter}}(t_{i,z})$ at both the sentence and document level.
 
@@
 
-  (Go into details of per-genre differences here? or leave out, for
-  space reasons? -- leave out, I'd say)
+  include genre tables here as well -- cut out again if no story.
 

### 2009-05-10 mtjournal.tex
more text on human judgments in paradigm section

@@
 
-\subsection{Reference evaluation measures}
+\subsection{Reference evaluation measures over parallel translations}
 
-TO DO: TALK ABOUT HUMAN F/A JUDGMENTS
+For some of the experiments presented here (sections~\ref{sec:hter1}
+and~\ref{sec:hter2}), we have multiple translations of the same source
+segment, each from a different translation engine and each with its
+own HTER score.  For these translations, reporting correlations of
+metric $m$ with absolute HTER scores can conflate $m$'s power in
+identifying which of two candidate translations is better with $m$'s
+(and HTER)'s ability to distinguish which source segments are
+more difficult to translate.
 
-When considering a candidate measure $m$, reporting correlations of
-$m$ with absolute HTER scores can conflate $m$'s power in identifying
-which of two candidate translations is better with $m$'s (and HTER)'s
-ability to distinguish which of two source documents is more difficult
-to translate. To avoid this conflation, we consider the mean-removed
-$m$ and HTER scores; this lowers the correlations but ensures that the
-reported correlations are among differences in the translations rather
-than among differences in the underlying document.
+To avoid this conflation, on the HTER corpora we report correlations of
+the mean-removed $m$ and HTER scores. Mean-removal  lowers the
+correlation $r$ values reported but
+ensures that the reported correlations are among differences in the
+translations rather than among differences in the underlying segments.
+
+Section~\ref{sec:hter2} reports HTER correlations per-sentence. There,
+we use length-weighted correlation, since unweighted correlations
+effectively puts too much emphasis on short sentences.  We do not
+weight for length in section~\ref{sec:hter1}, which reports document
+correlations, because documents are assumed to be equally important
+(regardless of length).
 

### 2009-05-10 mtjournal.tex
matt's grant number, revised to-do lists
body text and tables for first half of fa expts

@@
 
-TO DO: INTRODUCE THIS SECTION
+To explore members of the DPM metric family, we explore the
+correlations of various configurations against a corpus of human
+judgments of fluency and adequacy.
+
+% TO DO: more to say here?
 
@@
 
+\subsection{Exploring decompositions}
+We begin by comparing some of the simplest DPM measures to baseline
+measures case-sensitive BLEU (4-grams, with add-one smoothing) and
+TER.  For these first experiments, we consider only the 1-best parse
+($n=1$).
 \begin{table}
-    % \centering
-    \caption{$dl,lh$ decompositions are
-      more correlated with the fluency-adequacy average than BLEU$_4$
-      or $dlh$ decompositions.}
-    \begin{tabular}{lr}
-      metric  &    $r$ \\
+  \caption{Comparing $dl,lh$ to $dlh$ and baselines.}
+  \label{tab:facorr:subgraphs}
+  \begin{tabular*}{2.5in}{lr}
+    \hline
+    metric  &    \multicolumn{1}{c}{$r$} \\
     \hline
@@
     TER &      -0.173 \\
-    \end{tabular}
-    \label{tab:facorr:subgraphs}
+    \hline
+  \end{tabular*}
 \end{table}
+In table~\ref{tab:facorr:subgraphs} we see that using the
+partial-label $dl,lh$ decomposition has a better correlation with the
+fluency/adequacy scores than TER or BLEU$_4$.  These results confirm,
+with a PCFG, what \inlinecite{owczarzak07labelleddepseval} found with
+an LFG parser: that partial-dependency matches are better correlated
+with human judgments than full-dependency links.
 
-Table~\ref{tab:facorr:subgraphs} indicates that: using
-partial-labeling (dl,lh) much better than full link. Also, using
-Charniak parser as labeler seems to work, LFG not necessary
-Here, $n=1$ and only the $F[\cdot]$ score combination is shown.
+In table~\ref{tab:facorr:combinations}, we compare among several other
+members of the DPM family. 
+\begin{table}
+  \begin{tabular*}{2.5in}{lr}
+    \hline
+    metric  &    \multicolumn{1}{c}{$r$} \\
+    \hline
+    $F[1g,2g,dl,lh]$      &  0.237 \\
+    $\mu_{PR}[1g,2g,dl,lh]$ &  0.217 \\
+%
+    $F[dl,lh]$            &  0.226 \\
+    $\mu_{PR}[dl,lh]$      &  0.208 \\
+    \hline
+  \end{tabular*}
+  \caption{Comparing combination methods $F[\cdot]$ and
+    $\mu_{PR}[\cdot]$, and comparing the $1g,2g,dl,lh$ decomposition
+    with the $dl,lh$ decomposition.}
+  \label{tab:facorr:combinations}
+\end{table}
+Here, we find that combining the decompositions naively in $F[\cdot]$
+(before computing precision and recall) has a better $r$ than the
+harmonic mean of precision and recall subscores in $\mu_{PR}[\cdot]$.
+We also see that including some simple surface-matching features (the
+$1g,2g$ n-gram decompositions) improves correlations further beyond
+the simple $dl,lh$ features.
+
+% (too many chances for zeroes?)
+
+\subsection{Exploring parse expectations}
 
 \begin{itemize}
-
-\item EXPT 
-  combining precision, recall naively vs. combining items
-
-\begin{verbatim}
-  F(1g,2g,dl,lh)        0.237
-  prmeans(1g,2g,dl,lh)  0.217
-
-  F(dl,lh)              0.226
-  prmeans(dl,lh)        0.208
-\end{verbatim}
-
-  point: combining individual items naively (before computing
-  precision and recall) seems better than naively combining precisions
-  and recalls (too many chances for zeroes?)
 
@@
 
+Matthew Snover is supported under the GALE Program, DARPA/IPTO
+Contract No. HR0011-06-C-0022.
+
 % The endnotes section will be placed here.  But I don't think we have any

### 2009-05-10 mtjournal.tex
finished body of fa correlations section

@@
     $\mu_{PR}[1g,2g,dl,lh]$ &  0.217 \\
-%
+    \\
+    $F[1g,2g]$            & 0.227 \\
+    $F[1g,dl,dlh]$        & 0.227 \\
+    \\
     $F[dl,lh]$            &  0.226 \\
@@
     $\mu_{PR}[\cdot]$, and comparing the $1g,2g,dl,lh$ decomposition
-    with the $dl,lh$ decomposition.}
+    with the $dl,lh$ decomposition and other decompositions.}
   \label{tab:facorr:combinations}
@@
 harmonic mean of precision and recall subscores in $\mu_{PR}[\cdot]$.
-We also see that including some simple surface-matching features (the
-$1g,2g$ n-gram decompositions) improves correlations further beyond
-the simple $dl,lh$ features.
+We also see that we can combine the benefits of string-local $n$-grams
+($F[1g,2g]$) with the benefits of dependency information ($F[dl,lh]$)
+for a further improved correlation with human judgment, with the best
+correlation in $F[1g,2g,dl,lh]$. Including progressively larger chunks
+of the dependency graph (as in $F[1g,dl,dlh]$, which is inspired by
+the BLEU$_k$ idea of progressively larger $n$-grams) does not seem to
+be an improvement over $F[dl,lh]$.
 
@@
 
-\begin{itemize}
-
-\item EXPT
+To explore the effect of extracting parse feature-counts from a forest
+instead of a single tree, we present in
+table~\ref{tab:facorr:multiparse} several variants, with $n=1$ and
+with $n=50$.  
+ \begin{table}
+  \begin{tabular*}{2.5in}{lrr}
+    \hline
+    metric  & \multicolumn{1}{c}{$n$} &  \multicolumn{1}{c}{$r$} \\
+    \hline
+    $F[1g,2g,dl,lh]$      & 50 &  0.239 \\
+    $F[1g,2g,dl,lh]$      &  1 &  0.237 \\
+    \\
+    $F[1g,dl,lh]$            &  50 & 0.237 \\
+    $F[1g, dl,lh]$            &  1 & 0.234 \\
+    \\
+    $F[dl,lh]$            &  50 & 0.234 \\
+    $F[dl,lh]$            &  1 & 0.226 \\
+    \hline
+  \end{tabular*}
+  \caption{Comparing $n=1$ to $n=50$ for several variants of the DPM measure.}
   \label{tab:facorr:multiparse}
-  expanding n, gamma=0
-
-\begin{verbatim}
-  F[1g,2g,dl,lh],n=50   0.239
-    ...          n=1    0.237
-  F[dl,lh],n=50         0.234
-    ...    n=1          0.226
-\end{verbatim}
-  point: larger n $\to$ better r
-  (holds for other comparisons too, but these are good examples)
-
-\item  EXPT
-\label{tab:facorr:parseweight}
-  setting gamma
-
-  Tuning experiments finds that gamma of 0.25 is slightly better
-  (especially when using F[dl,lh]) than gammas of other values in
-  range 0-1 (these values probably not significant)
+\end{table}
+For the $n=50$ cases, we set $\gamma=0$ to assign
+uniform probabilities to the parse forest, to compare as closely as
+possible to \inlinecite{owczarzak07labelleddepseval}, which includes a
+\textbf{d\_var\_50} measure with 50-best parses, with ranks but no
+weights.  While not all of these differences are significant, the
+trend is universally that the correlation $r$ improves as the number
+of parses ($n$) is increased.
+Tuning experiments find that increasing $\gamma$ to 0.25 can increase
+the $r$ reported here for $F[1g,2g,dl,lh]$ marginally (but insignificantly).
 
@@
 (table~\ref{tab:facorr:subgraphs}), multiple parses ($n=50$) based on
-table~\ref{tab:facorr:multiparse}, and $\gamma=0.25$, hinted at by
-table~\ref{tab:facorr:parseweight}.  We use these EDPM
-parameter-settings in the experiments exploring document-level correlations with HTER
-(section~\ref{sec:hter1}).
-
-\end{itemize}
+table~\ref{tab:facorr:multiparse}, and $\gamma=0.25$.
+%, hinted at by table~\ref{tab:facorr:parseweight}.  
+We use these EDPM parameter-settings in the experiments exploring
+document-level correlations with HTER (section~\ref{sec:hter1}).
 

### 2009-05-10 mtjournal.tex
incorporated Mari's draft abstract

@@
 \begin{abstract}
-Abstract to go here
+  Recent efforts aimed at improving over standard machine translation
+  evaluation methods (BLEU, TER) have investigated mechanisms for
+  accounting for allowable wording differences either in terms of
+  syntactic structure or synonyms/paraphrases. This paper explores an
+  approach for combining scores from partial syntactic dependency
+  matches with standard local n-gram matches using a statistical
+  parser, and taking advantage of parse probabilities in deriving
+  expected scores based on the N-best parses for the hypothesized
+  sentence translation.  The new scoring metric, Expected Dependency
+  Pair Match (EDPM), is shown to be superior to BLEU and TER in terms
+  of correlation to human judgements and as a per-document and
+  per-sentence predictor of HTER, using mean subtraction to account
+  for document difficulty. Further, we explore the potential benefit
+  of combining the n-gram and syntactic features of EDPM with the
+  alternative wording features of TERp, with experiments showing that
+  there is a benefit to accounting for syntactic structure on top of
+  the semantic equivalency features.
 \end{abstract}
@@
 
+\subsection{HTER corpus}
+
 The GALE 2.5 translation corpus is made up of system translations into

### 2009-05-10 mtjournal.tex
cleaned up inline citations

@@
 other metrics model syntactically-local (rather than string-local)
-word-sequences. \cite{liu05syntaxformteval} compared tree-local
+word-sequences. \inlinecite{liu05syntaxformteval} compared tree-local
 $n$-gram precision in various configurations of constituency and
@@
 over a decomposition of reference and hypothesis trees.
-\cite{owczarzak07labelleddepseval}'s \textbf{d} and \textbf{d\_var}
-measures compare LFG-derived relational tuples from reference and
-hypothesis translations and report substantial improvement in
-correlation with human judgment relative to BLEU and TER.
+\inlinecite{owczarzak07labelleddepseval} proposed \textbf{d} and
+\textbf{d\_var} measures that compare LFG-derived relational tuples
+from reference and hypothesis translations and reported substantial
+improvement in correlation with human judgment relative to BLEU and
+TER.
 
@@
 PCFG parsers with deterministic head-finding, while others
-\cite{owczarzak07labelleddepseval}extract the semantic dependency
+\cite{owczarzak07labelleddepseval} extract the semantic dependency
 relations from an LFG parser \cite{cahill04lfg}.
@@
 This work extends the dependency-scoring strategies of
-\cite{roark06:sparseval,owczarzak07labelleddepseval} using a
+\inlinecite{roark06:sparseval,owczarzak07labelleddepseval} using a
 widely-used and publically available PCFG parser and deterministic
@@
 judgments of fluency and adequacy.  In other work,
-e.g. \cite{snover06ter}, measures are evaluated by comparison to
+e.g. \inlinecite{snover06ter}, measures are evaluated by comparison to
 human-targeted TER (HTER), a distance to a human-revised reference
@@
 extensions of the dependency-pair F measures found in
-\cite{roark06:sparseval,owczarzak07labelleddepseval}.  Rather than
+\inlinecite{roark06:sparseval,owczarzak07labelleddepseval}.  Rather than
 comparing string sequences, as BLEU does with its $n$-gram precision,
@@
 \end{eqnarray}
-Dependency-based SParseval and Owczarzak \textit{et al.}'s
-\cite*{owczarzak07labelleddepseval} \textbf{d} approach, then, may
+Dependency-based SParseval and the \textbf{d} approach from
+\inlinecite{owczarzak07labelleddepseval}, then, may
 each be understood as $F[dlh]$, while the latter's \textbf{d\_var}
@@
 In this work, we use a state-of-the-art PCFG (the first stage of
-\cite{charniak-johnson:2005:ACL}) and context-free head-finding rules
+\inlinecite{charniak-johnson:2005:ACL}) and context-free head-finding rules
 \cite{magerman95headfinding} to generate a 50-best list of dependency
@@
 implementation of labelled-dependency \textsc{SParseval}, and may be
-considered as an approximation of the rich semantics generated by
-\cite{cahill04lfg} or another heavily knowledge-engineered
+considered as an approximation of the rich semantics generated by LFG
+parsers \cite{cahill04lfg} or another heavily knowledge-engineered
 parser, but with much less knowledge-engineering required.
@@
   \textbf{d\_50} and \textbf{d\_50\_var} measures.\footnote{Since
-    \cite{owczarzak07labelleddepseval} report no use of parse weights,
+    \inlinecite{owczarzak07labelleddepseval} report no use of parse weights,
     \textbf{d\_50} and \textbf{d\_50\_var} may be using a sum of
@@
 different reference translation.  For a rough\footnote{Our segment
-  count differs from \cite{owczarzak07labelleddepseval}, who report
+  count differs from \inlinecite{owczarzak07labelleddepseval}, who report
   16,800 segments over the same corpus. We find baseline correlations
@@
 }
-comparison with \cite{owczarzak07labelleddepseval}, we treat each
+comparison with \inlinecite{owczarzak07labelleddepseval}, we treat each
 judgment as a separate segment.

### 2009-05-11 mtjournal.tex
whitespace and obsolete comment cleanup
basic structure of hter1 section in place

@@
 
-%
 In this work, we use a state-of-the-art PCFG (the first stage of
@@
 \begin{figure}
-  % \centering
   \Tree
@@
 identifying which of two candidate translations is better with $m$'s
-(and HTER)'s ability to distinguish which source segments are
+(and HTER's) ability to distinguish which source segments are
 more difficult to translate.
 
-To avoid this conflation, on the HTER corpora we report correlations of
-the mean-removed $m$ and HTER scores. Mean-removal  lowers the
-correlation $r$ values reported but
-ensures that the reported correlations are among differences in the
+To avoid this conflation, on the HTER corpora we report correlations
+on $\bar{\mu}(m)$
+and $\bar{\mu}(\textrm{HTER})$ (rather than measure $m$ and HTER), where
+$\bar{\mu}$ represents mean-removal:
+\begin{equation}
+  \label{eq:meansub}
+  \bar{\mu}(m(t_i)) = m(t_i) - \sum_{j=1}^I\frac{m(t_j)}{I}
+\end{equation}
+Mean-removal lowers the correlation $r$ values reported but ensures
+that the reported correlations are among differences in the
 translations rather than among differences in the underlying segments.
@@
 In this section, we have presented experiments exploring a number of
-variants of the DPM metric against an average fluency/adequacy judgment. The experiments suggest a
-best-case variant, where we set:
+variants of the DPM metric against an average fluency/adequacy
+judgment. The experiments suggest a best-case variant, where we set:
 \begin{displaymath}
@@
 \label{sec:hter1}
-TO DO: INTRODUCE THIS SECTION
+In this section, we move from an exploration of the space of possible
+DPM metrics to a single chosen DPM-derived metric (EDPM).  Here, we
+compare EDPM to other metrics in correlation with document-level HTER
+performance.
 
@@
 \begin{table}
-%  \centering
   \begin{tabular}{r|rr|rr|rr}
@@
 \cite{gale08phase2_5references} for each sentence $i$ and a system
-translation $t_{i,z}$ for each of the three systems
-$z$. Additionally, each of the system translations of each segment $i$
-has a corresponding human-targeted reference aligned at the sentence
-level, so we have available the HTER score of each segment
-$s_{\textsc{hter}}(t_{i,z})$ at both the sentence and document level.
+translation $t_{i,z}$ for each of the three systems $z$. Additionally,
+each of the system translations of each segment $i$ has a
+corresponding human-targeted reference aligned at the sentence level,
+so we have available the HTER score of each segment
+$s_{\textrm{HTER}}(t_{i,z})$ at both the sentence and document level.
 
-  prediction:
-    Mean-subtracted HTER per-document
+As discussed in section~\ref{sec:paradigm}, we report correlations
+here between $\bar{\mu}(m)$ and $\bar{\mu}(\textrm{HTER})$, rather than
+$m$ and HTER directly, to abstract away from variations in the
+underlying documents.
 
-  base measure EDPM: F[1g,2g,dl,lh],n=50,gamma=0.25
+In table~\ref{tab:hterperdoc}, we show per-document Pearson's $r$
+between $\bar{\mu}$EDPM and $\bar{\mu}$HTER, as well as two other
+baselines: $\bar{\mu}$TER and $\bar{\mu}$BLEU$_4$.  
+\begin{table}
+  \begin{tabular}{r|rrrr|rr|r}
+    \hline
+    $r$ vs. $\bar{\mu}$HTER & \multicolumn{1}{c}{\texttt{bc}}
+    & \multicolumn{1}{c}{\texttt{bn}} &
+    \multicolumn{1}{c}{\texttt{nw}} & \multicolumn{1}{c}{\texttt{wb}}
+    & \multicolumn{1}{|c}{all Arabic} & \multicolumn{1}{c|}{all Chinese}
+    & \multicolumn{1}{c}{all} \\
+    \hline
+    $\bar{\mu}$TER
+    &  0.59 &  \textbf{0.35} &  \textbf{0.47}&  \textit{0.17}
+    & \textbf{0.54} & \textbf{0.32} &  0.44\\
+    $\bar{\mu}$BLEU
+    & -0.42 & \textbf{-0.32} & \textbf{-0.46}&  \textbf{-0.27}
+    & -0.42  &  \textbf{-0.33} & -0.37 \\
+    $\bar{\mu}$EDPM
+    & \textbf{-0.69} & \textbf{-0.39} & \textbf{-0.47} &
+    \textbf{-0.27}
+    & \textbf{-0.60} & \textbf{-0.39} & \textbf{-0.50} \\
+    \hline
+  \end{tabular}
+  \caption{Per-document correlations of $\bar{\mu}$EDPM and others to
+    $\bar{\mu}$HTER, by genre and by source language. Bold numbers are
+  within 95\% significance of the best per column; italics indicate
+  that the sign of the $r$ value has less than 95\% confidence.}
+  \label{tab:hterperdoc}
+\end{table}
+We see that the EDPM has the highest correlation in each of the
+subcorpora created by dividing by genre or by source language, as well
+as the corpus as a whole.  Not every difference is above 95\%
+confidence, but the trend is the same across all of the
+subcorpora. EDPM's strengths are particularly clear in the
+unstructured domains (\texttt{wb} and \texttt{bc}).
 
-  [EXPT]
-\begin{verbatim}
-            Ar     Zh    All
-   TER     0.51   0.32  0.44
-   BLEU_4 -0.42  -0.33 -0.37
-   EDPM   -0.60* -0.39 -0.50*
-\end{verbatim}
-  point: base EDPM does better on all docs, and on Arabic. difference
-  is not quite significant at the p=0.05 on the Chinese documents
+TO DO: NOTE RE: DELTAS
 
-  Note 1: also tried pairwise deltas per-document, with similar
-  results, as presented in the MetricsMATR competition submission.
+TO DO: NOTE RE: SEGMENT-LEVEL CORRS
 
-  Note 2: when using per-segment mean-subtracted, r values are smaller
-  and *Arabic* was the language where EDPM did not perform best.
-
-\begin{verbatim}
-               Ar     Zh     All
-      TER     0.38*  0.04   0.19
-      BLEU_4 -0.10  -0.16  -0.14
-      EDPM   -0.31  -0.19* -0.24*
-\end{verbatim}
-
-  include genre tables here as well -- cut out again if no story.
+%   Note 2: when using per-segment mean-subtracted, r values are smaller
+%   and *Arabic* was the language where EDPM did not perform best.
+% \begin{verbatim}
+%                Ar     Zh     All
+%       TER     0.38*  0.04   0.19
+%       BLEU_4 -0.10  -0.16  -0.14
+%       EDPM   -0.31  -0.19* -0.24*
+% \end{verbatim}
 

### 2009-05-11 mtjournal.tex
clean up tables to save some space
added table with per-seg HTER correlations

@@
     $\mu_{PR}[1g,2g,dl,lh]$ &  0.217 \\
-    \\
+    \rlcline{1-1} \rlcline{2-2}
     $F[1g,2g]$            & 0.227 \\
     $F[1g,dl,dlh]$        & 0.227 \\
-    \\
+    \rlcline{1-1} \rlcline{2-2}
     $F[dl,lh]$            &  0.226 \\
@@
     $F[1g,2g,dl,lh]$      &  1 &  0.237 \\
-    \\
+    \rlcline{1-2}\rlcline{3-3}
     $F[1g,dl,lh]$            &  50 & 0.237 \\
     $F[1g, dl,lh]$            &  1 & 0.234 \\
-    \\
+    \rlcline{1-2}\rlcline{3-3}
     $F[dl,lh]$            &  50 & 0.234 \\
@@
   \begin{tabular}{r|rr|rr|rr}
+    \hline
      & \multicolumn{2}{c|}{Arabic} & \multicolumn{2}{c|}{Chinese}
@@
      Total & 259 & 2593& 257& 2709 & 516 & 5302\\
+     \hline
   \end{tabular}
@@
 unstructured domains (\texttt{wb} and \texttt{bc}).
+
+In table~\ref{tab:hterpersent}, we see a similar story.
+\begin{table}
+  \begin{tabular}{r|rrrr|rr|r}
+    \hline
+    $r$ vs. $\bar{\mu}$HTER & \multicolumn{1}{c}{\texttt{bc}}
+    & \multicolumn{1}{c}{\texttt{bn}} &
+    \multicolumn{1}{c}{\texttt{nw}} & \multicolumn{1}{c}{\texttt{wb}}
+    & \multicolumn{1}{|c}{all Arabic} & \multicolumn{1}{c|}{all Chinese}
+    & \multicolumn{1}{c}{all} \\
+    \hline
+    $\bar{\mu}$TER
+    &  \textbf{0.44} &  \textbf{0.29} &  \textbf{0.33}&  0.25
+    & \textbf{0.44} & 0.25  &  \textbf{0.36}\\
+    $\bar{\mu}$BLEU
+    & -0.31 & -0.24  & -0.29 &  -0.25
+    & -0.31  &  -0.24  & -0.28 \\
+    $\bar{\mu}$EDPM
+    & \textbf{-0.46} & \textbf{-0.31} & \textbf{-0.34} &
+    \textbf{-0.30}
+    & \textbf{-0.44} & \textbf{-0.30} & \textbf{-0.37} \\
+    \hline
+  \end{tabular}
+  \caption{Per-sentence, length-weighted correlations of
+    $\bar{\mu}$EDPM and others to $\bar{\mu}$HTER, by genre and by
+    source language. Bold numbers are within 95\% significance of the
+    best per column.}
+  \label{tab:hterpersent}
+\end{table}
 

### 2009-05-11 mtjournal.tex
roughed out draft of HTER correlations section

@@
 
-\section{Per-document correlations with HTER}
+\section{Correlating EDPM with HTER}
 \label{sec:hter1}
@@
 DPM metrics to a single chosen DPM-derived metric (EDPM).  Here, we
-compare EDPM to other metrics in correlation with document-level HTER
-performance.
+compare EDPM to other metrics in correlation with document-level and
+segment-level HTER performance.
 
@@
 so we have available the HTER score of each segment
-$s_{\textrm{HTER}}(t_{i,z})$ at both the sentence and document level.
+HTER$(t_{i,z})$ at both the sentence and document level.
 
@@
 baselines: $\bar{\mu}$TER and $\bar{\mu}$BLEU$_4$.  
+\begin{subtable}
   \begin{table}
@@
   \end{table}
-We see that the EDPM has the highest correlation in each of the
-subcorpora created by dividing by genre or by source language, as well
-as the corpus as a whole.  Not every difference is above 95\%
-confidence, but the trend is the same across all of the
-subcorpora. EDPM's strengths are particularly clear in the
-unstructured domains (\texttt{wb} and \texttt{bc}).
-
-In table~\ref{tab:hterpersent}, we see a similar story.
   \begin{table}
@@
       $\bar{\mu}$EDPM and others to $\bar{\mu}$HTER, by genre and by
-    source language. Bold numbers are within 95\% significance of the
-    best per column.}
+      source language.}
     \label{tab:hterpersent}
   \end{table}
+\end{subtable}
+We see that the EDPM has the highest correlation in each of the
+subcorpora created by dividing by genre or by source language, as well
+as the corpus as a whole.  Not every difference is above 95\%
+confidence, but the trend is the same across all of the
+subcorpora. EDPM's strengths, relative to the other measures, are
+particularly clear in the unstructured domains (\texttt{wb} and
+\texttt{bc}).
+
+Table~\ref{tab:hterpersent} presents correlations weighted by sentence
+length.  Though many of the $r$ values are smaller magnitude in this
+correlation, EDPM again has the largest correlation in each category.
+TER, however, does quite well with length-weighted sentence
+correlation, with $r$ values within 95\% confidence of EDPM scores on
+nearly every breakdown.
 
@@
 
-TO DO: NOTE RE: SEGMENT-LEVEL CORRS
-
-%   Note 2: when using per-segment mean-subtracted, r values are smaller
-%   and *Arabic* was the language where EDPM did not perform best.
-% \begin{verbatim}
-%                Ar     Zh     All
-%       TER     0.38*  0.04   0.19
-%       BLEU_4 -0.10  -0.16  -0.14
-%       EDPM   -0.31  -0.19* -0.24*
-% \end{verbatim}
-
 \section{Weight-tuning to combine syntax and other knowledge sources}
 \label{sec:hter2}
+
+In the previous section, we observed that TER and EDPM each perform
+better than BLEU when considering mean-removed length-weighted
+correlations with HTER.  We observe that TER's basic operations
+(insertions, deletions, shifts and substitutions) are fundamentally
+different from those of EDPM, which is an $F$-style overlap measure
+and is seeking to make explicit comparison based on an inferred
+syntactic structure.
 

### 2009-05-11 mtjournal.tex
added some r confidence intervals, cleaned up intro to tuning

@@
 
-Table~\ref{tab:hterpersent} presents correlations weighted by sentence
-length.  Though many of the $r$ values are smaller magnitude in this
-correlation, EDPM again has the largest correlation in each category.
-TER, however, does quite well with length-weighted sentence
-correlation, with $r$ values within 95\% confidence of EDPM scores on
-nearly every breakdown.
+Table~\ref{tab:hterpersent} presents per-sentence correlations,
+weighted by sentence length.  Though many of the $r$ values are
+smaller magnitude in this correlation, EDPM again has the largest
+correlation in each category.  TER, however, does quite well with
+length-weighted sentence correlation, with $r$ values within 95\%
+confidence of EDPM scores on nearly every breakdown.
 
@@
 and is seeking to make explicit comparison based on an inferred
-syntactic structure.
+syntactic structure.  These two different approaches (word-level and
+syntactic features) are complementary, and we seek in this section to
+explore mechanisms for combining them.
 
-TERp optimizer [Matt, a brief description here]
+As discussed in section~\ref{sec:intro}, the TER Plus (TERp) research
+tools \cite{snover09terp} provide an optimizer for weighting multiple
+simple subscores, including selecting weights for the TER basic
+operations. TERp's feature list extends TER to additionally include
+synonym tables and an empirically-derived paraphrase table.
 
-TERp comes with a variety of features (described earlier)
+TO DO: TERp optimizer [Matt, a brief description here]
 
-include optionally new features
+In the experiments presented in this section, we extend this feature
+list by including the syntactic overlap features, and we use the TERp
+optimizer tools to tune the relative weight of each feature.
 
-Corpus: GALE 2.5 data, again documents randomly split into two groups,
-evenly split across language and genre
+\subsection{Features}
 
-tuning target: weighted correlation with mean-removed segment HTER
-
-[we use weighted correlation to avoid emphasis on short segments
-(which hurts document-level and system-level utility) ]
-
-Select features from set:
-
+We train a model using features from one or more of the following
+sets:
 \begin{description}
-\item[E]: features from EDPM, specifically, P,R for inbound, outbound,
-  full-links, and unlabeled links  (8 features)
-\item[T]: features from basic TERp
-  (inserts, deletes, substitutions, shifts)  (7 features)
-\item[P]: features from TERp paraphraser
-  (4 features)
-\item[N]: precision and recall for 1-grams, 2-grams
-\item[B]: brevity/prolixity (2 features: one for longer-than-ref, one for
+\item[E]: the fully syntactic features from the DPM family.
+  Specifically, this featureset includes precision and recall scores
+  for the $dl$, $lh$, $dlh$, and $dh$ decompositions of the syntactic
+  structure, using $n=50$ and $\gamma=0.25$.  The \textbf{E} set
+  includes 8 features.
+\item[N]: the non-syntactic features from the DPM family;
+  specifically, precision and recall for $1g$ and $2g$
+  decompositions.  The \textbf{N} set includes 4 features.
+\item[T]: the features from basic TERp, which include a separate cost
+  for each of insertions, deletions, substitutions, shifts, and
+  synonym-alignment. The \textbf{T} set includes 7 features.
+  % TO DO: what are the other two?
+\item[P]: features from the automatically-derived TERp paraphrase table
+  (4 features)  % TO DO: Matt, what are these
+\item[B]: two ``length penalties'' one is assigned the count of
+  hypothesis words beyond the count of reference words, and one
+  assigned the count of reference words beyond the count of hypothesis
+  wordsbrevity/prolixity (2 features: one for longer-than-ref, one for
   shorter-than-ref)
@@
 
-tune weights on one set, test on the other (and vice versa).
-reporting average r between the two.
+\subsection{Corpus}
 
+For these tuning experiments we use the same GALE 2.5 documents and
+sentences from the previous section.  Since this approach requires
+tuning, we assign documents randomly into two groups, such that each
+group has the same document distribution across source-language and
+genre.
+%
+These two corpus halves are used in two-fold cross-validation, testing
+on one and training on the other.  We tune on (and report correlation with)
+length-weighted per-sentence correlation with HTER, using mean-removed
+scores as before.
+
+% tuning target: weighted correlation with mean-removed segment HTER
+
+% [we use weighted correlation to avoid emphasis on short segments
+% (which hurts document-level and system-level utility) ]
+In table~\ref{tab:tuneresults}, we report the average $r$ between the
+two.   
 \begin{verbatim}
@@
       TPB    0.4010
-      TP     0.3946
+      TP     0.3946  > 0.416
       TPNB   0.3940
@@
       ETB    0.3871
-      ET     0.3827
-      TB     0.3815
+      ET     0.3827  > 0.404
+      TB     0.3815  > 0.403
       ETNB   0.3814
       ETN    0.3804
-      T      0.3758
+      T      0.3758  > 0.398
       TNB    0.3708
-      TN     0.3680
-      E      0.3287
+      TN     0.3680  > 0.390
+      TER    0.353   > 0.375
+      E      0.3287  > 0.352
       EB     0.3267
       ENB    0.3149
-      NB     0.3058
+      NB     0.3058  > 0.329
       EN     0.2973
       N      0.2636
+      BLEU   0.244
       B      0.1638
@@
 (maybe don't report all these numbers!)
+
+TO DO: discuss this as a trend, despite lack of significance? plot charts?
 
@@
 
-\appendix
+% \appendix
 
-[don't think we want an appendix]
+% [don't think we want an appendix]
 

### 2009-05-11 mtjournal.tex
framework for a figure for tuneable weights

@@
 % (which hurts document-level and system-level utility) ]
-In table~\ref{tab:tuneresults}, we report the average $r$ between the
-two.   
+In figure~\ref{fig:tuneresults}, we plot the Spearman's $r$ (with 95\%
+confidence interval) for the results on the two test sets (each
+trained on one and tested on the other).
+\begin{figure}
+  \centerline{{one figure}\qquad{}{two figure}}
+  \caption{Figure with two tune-set plots will go here.}
+  \label{fig:tuneresults}
+\end{figure}
+
 \begin{verbatim}

### 2009-05-11 mtjournal.tex
partial draft of conclusion

@@
 
+question: this is kind of like the $\mu_{PR}[\cdot]$ --- seems like it
+also doesn't work so well. compare to EDPM as-is
+
+talk about why EN is < EDPM (is it?)
+
 \section{Conclusion}
 \label{sec:conclusion}
-expected dependency pair matching gets at something new
+In this research, we explore dependency pair match (DPM) a family of
+syntactically-oriented measures.  Through a corpus of human judgments,
+we settle on EDPM, a member of that family with promising predictive
+power.  When tested against correlations with HTER, we find that EDPM
+is as good a predictor of HTER as current measures BLEU$_4$ and TER,
+and --- depending on the language and genre --- often better.
 
-troubling area: speed of analysis -- current implementation requires
-running a complete n-best parser
+Several areas of future work are open to us. Using these syntactic
+techniques may be possible as a late-pass evaluation, to identify how
+candidate translation systems are performing overall.  Alternatively,
+the various tree-decompositions of the DPM family could be used as
+system diagnostics, by looking at relative quality of these component
+scores compared to those of an alternative candidate system.
 
-possible use: as a late-pass evaluation, to identify how systems
-perform overall
+Another interesting research area is in the quality of the parser.  In
+this work, we demonstrated that the
+\inlinecite{charniak-johnson:2005:ACL} PCFG parser could serve in
+place of the LFG parser used in
+\inlinecite{owczarzak07labelleddepseval}. PCFG parsers, however, are
+fairly easily adaptable to another domain. One interesting research
+direction explores the intersection of genre (and genre-mismatch),
+parser quality, and prediction of human performance. Another
+consideration is the speed of the parser: parse time currently takes
+much, much longer than BLEU, TER, or other word-based measures. We are
+interested in the trade-offs --- if any --- among parse quality,
+speed, and predictive power as a translation metric.  One possibility,
+for example, is to store packed forests \cite{huang08packedforests}
+rather than generating an $n$-best list only to sum across it again in
+calculating the expectation.
 
-future work:  explore ways to get at syntactic information without
-the expense of a full parser:
-\begin{itemize}
-\item this work moves from an LFG parser (Owczarzak) to the
-  substantially simpler and more robust Charniak system - keep going
-  and look at simpler partial-parsing approaches (forests?)
-\item conversely: how important is it that the parses be of
-  high-quality?
-\end{itemize}
+TO DO: parting words
 

### 2009-05-11 mtjournal.tex
removed old edpm-paper files, move things to mtjournal name
whoops, correct Spearman's => Pearson's

@@
 % (which hurts document-level and system-level utility) ]
-In figure~\ref{fig:tuneresults}, we plot the Spearman's $r$ (with 95\%
+In figure~\ref{fig:tuneresults}, we plot the Pearson's $r$ (with 95\%
 confidence interval) for the results on the two test sets (each
@@
 \bibliographystyle{klunamed}
-\bibliography{edpm-paper}
+\bibliography{mtjournal}
 

### 2009-05-11 mtjournal.tex
removed call to ulem package -- strikeout not needed
overview paragraph at end of intro roughed in.

@@
 \newdisplay{guess}{Conjecture}
-
-% TO DO: package provides \sout{} strikeout macro -- remove this
-% before submission
-\usepackage{ulem}
 
@@
 
-% override ulem 's default change of \emph{} -- remove before ship
-\normalem
 \section{Introduction}
@@
 
-TO DO: OVERVIEW OF THE REST OF THE PAPER
+In the following section (section~\ref{sec:approach}), we outline our
+approach to including syntactic information in MT evaluation, by
+describing a family of metrics. In section~\ref{sec:paradigm}, we
+describe the implementation details and make explicit which dimensions
+of this family of metrics we explore in this paper.
+Section~\ref{sec:faexpts} examines the correlation of several members
+of this family with human judgments of fluency and adequacy over a
+corpus of sentence translations and selects a favorable member of the
+family (EDPM) based on this comparison. Section~\ref{sec:hter1} compares
+the correlation of EDPM with HTER over a corpus of document- and
+sentence-level comparisons, and demonstrates that its correlation is
+favorable to that of competitor measures BLEU and TER at both sentence
+and document levels, regardless of source-language or genre.
+Section~\ref{sec:hter2} uses the tuning components of the TERp tools
+to explore the combination of these syntactic measures with other
+TERp-based word-based measures. Finally, section~\ref{sec:conclusion}
+concludes with some remarks on challenges and future work.
 
@@
 \label{sec:paradigm}
-In this section, we outline the details of the implementation and
+In this section, we describe the details of the implementation and
 experimental variables explored in the experiments presented here.

### 2009-05-11 mtjournal.tex
footnote on deltas

@@
 translations rather than among differences in the underlying segments.
+\footnote{ Previous work \cite{kahn08metricsmatr} reported HTER
+  correlations against pairwise differences among translation segments
+  derived from the same source segment. Those results are similar to
+  those reported here. However, we believe that pairwise differences
+  introduce problems with the independence assumptions in
+  the Pearson's $r$ tests, and we use mean-removal
+  throughout this work instead.}
 
-Section~\ref{sec:hter2} reports HTER correlations per-sentence. There,
-we use length-weighted correlation, since unweighted correlations
-effectively puts too much emphasis on short sentences.  We do not
-weight for length in section~\ref{sec:hter1}, which reports document
-correlations, because documents are assumed to be equally important
-(regardless of length).
-
-TO DO: MENTION DELTAS IN A FOOTNOTE, AND CITE MetricsMATR SYSTEM PAPER
+When reporting HTER correlations per-sentence, we use length-weighted
+correlation, since unweighted correlations effectively puts too much
+emphasis on short sentences.  We do not weight for length for the
+per-document correlations in section~\ref{sec:hter1}, because we
+assume documents to be equally important (regardless of length).
 

### 2009-05-11 mtjournal.tex
grant numbers straightened out

@@
 
-[grant numbers for the DARPA/GALE grants here]
+This material is based upon work supported by the National Science
+Foundation under Grant No. 0741585 and the Defense Advanced Research
+Projects Agency under Contract Nos. HR0011-06-C-0022 and HR0011-06-C-0023.
 
-Matthew Snover is supported under the GALE Program, DARPA/IPTO
-Contract No. HR0011-06-C-0022.
+% Re: DARPA contract numbers
+% -0022 is Matt's [Agile] grant.
+% -0023 is Mari & Jeremy's [Nightingale] grant
+
+Any opinions, findings, and conclusions or recommendations expressed
+in this material are those of the author(s) and do not necessarily
+reflect the views of the National Science Foundation or the Defense
+Advanced Research Projects Agency.
 

### 2009-05-12 mtjournal.tex
whitespace and formatting cleanups [small acknowledgments]
incorporated Mari's suggested sentence on tuning

@@
 intended to be tuned to a development set by users.
-
-TO DO: DISCUSS TUNING'S VIRTUES, FLAWS
+%
+Tuning has the advantage that the weight of different types of errors
+can be adjusted to match the needs of the task, though it makes it
+more difficult to compare results across tasks, particularly when
+there is little data for tuning.
 
@@
 English from three sites.
-% MO added:
+%
 The three sites all use system combination to integrate results from
@@
 confidence of EDPM scores on nearly every breakdown.
-
-TO DO: NOTE RE: DELTAS
 
@@
 % \appendix
-
 % [don't think we want an appendix]
@@
 
+{\small
 This material is based upon work supported by the National Science
@@
 Advanced Research Projects Agency.
+}
 

### 2009-05-12 mtjournal.tex
edits from mari.

@@
 %
-Automatic measures of this kind (e.g.\ BLEU \cite{papineni02bleu}) not
+Automatic measures of this kind not
 only provide a well-defined evaluation standard but are also required
@@
 
-The most popular evaluation measures are currently BLEU (based on
-$n$-gram precision) and the edit-distance measure Translation Edit
+Currently, the most popular evaluation measures include a measure
+based on $n$-gram precision known as BLEU \cite{papineni02bleu}) and the edit-distance measure Translation Edit
 Rate (TER) \cite{snover06ter}.  Recent research has found that these
@@
 dependency trees.  The dependent-based SParseval measure
-\cite{roark06:sparseval}, originally designed as a parse-quality
+\cite{roark06:sparseval}, designed as a parse-quality
 metric for speech, is a similar approach, in that it is an F-measure
@@
 This work extends the dependency-scoring strategies of
-\inlinecite{roark06:sparseval,owczarzak07labelleddepseval} using a
+\inlinecite{roark06:sparseval} and \inlinecite{owczarzak07labelleddepseval} using a
 widely-used and publically available PCFG parser and deterministic

### 2009-05-13 mtjournal.tex
mari's suggested changes on intro and approach sections

@@
 
-In the following section (section~\ref{sec:approach}), we outline our
+In section~\ref{sec:approach}), we outline our
 approach to including syntactic information in MT evaluation, by
@@
 corpus of sentence translations and selects a favorable member of the
-family (EDPM) based on this comparison. Section~\ref{sec:hter1} compares
+family, Expected Dependency Pair Match (EDPM), based on this comparison. Section~\ref{sec:hter1} compares
 the correlation of EDPM with HTER over a corpus of document- and
@@
 to explore the combination of these syntactic measures with other
-TERp-based word-based measures. Finally, section~\ref{sec:conclusion}
+TERp word-based measures. Finally, section~\ref{sec:conclusion}
 concludes with some remarks on challenges and future work.
@@
 extensions of the dependency-pair F measures found in
-\inlinecite{roark06:sparseval,owczarzak07labelleddepseval}.  Rather than
+\inlinecite{roark06:sparseval} and \inlinecite{owczarzak07labelleddepseval}.  Rather than
 comparing string sequences, as BLEU does with its $n$-gram precision,
 this approach defers to a parser for an indication of the relevant
-words that indicate more about its meaning --- in these
+word tuples associated with meaning --- in these
 implementations, the head on which that word depends.  Each sentence
@@
 itself, if we use a probabilistic parser, can provide an $n$-best
-parse list forest for the translation reference and translation
-hypothesis. We use the probability statistics of the forest to compute
+parse list for the translation reference and translation
+hypothesis. We use the probability statistics of the list to compute
 expected counts for each decomposition. Though this approach yields
@@
 
-When multiple decomposition types used together, we may combine these
+When multiple decomposition types are used together, we may combine these
 subscores in a variety of ways.  We may compute precision and recall
@@
 Dependency-based SParseval and the \textbf{d} approach from
-\inlinecite{owczarzak07labelleddepseval}, then, may
+\inlinecite{owczarzak07labelleddepseval} may
 each be understood as $F[dlh]$, while the latter's \textbf{d\_var}
@@
 Both of the combination methods above ($F[\cdot]$ and
-$\mu_{PR}[\cdot]$) are ``naive'' combinations, in that they treat each
-decomposition as equivalent to the next.  One further direction to
+$\mu_{PR}[\cdot]$) are ``naive'' in that they treat each
+component score as equivalent to the next.  One further direction to
 explore involves tuning (presumably on held-out data) the precision
@@
 \label{sec:paradigm}
-In this section, we describe the details of the implementation and
-experimental variables explored in the experiments presented here.
+% In this section, we describe the details of the implementation and
+% experimental variables explored in the experiments presented here.
 

### 2009-05-13 mtjournal.tex
Mari's comments on paradigm section

@@
 \begin{description}
-\item[Decompositions]: we consider syntactic decompositions of the
-  following types.
+\item[Decompositions.] We consider syntactic decompositions of the
+  following types:
   \begin{description}
   \item[dlh] $\left\langle \textrm{Dependent}, \textrm{arc Label},
-      \textrm{Head}\right\rangle$ here, every tuple includes two words
+      \textrm{Head}\right\rangle$ Here, every tuple includes two words
     and the dependency-label between them.
   \item[dl] $\left\langle \textrm{Dependent}, \textrm{arc Label}
-      \right\rangle$ the tuple is the word plus the label of the arc
+      \right\rangle$ The tuple is the word plus the label of the arc
       linking it to its head; this decomposition marks how the word
@@
     \item[lh] $\left\langle \textrm{arc Label}, \textrm{Head}
-      \right\rangle$ each tuple here is the dependency relation plus
+      \right\rangle$ Each tuple here is the dependency relation plus
       the word being modified; this decomposition implicitly marks how
@@
     \item[dh] $\left\langle \textrm{Dependent}, \textrm{Head}
-      \right\rangle$ here, drops the arc-label. This decomposition
+      \right\rangle$ Here, drops the arc-label. This decomposition
       ignores the syntactic-role information implied in the arc labels.
-    \item[1g,2g] these decompositions are syntactically-null.  They
+    \item[1g,2g] These decompositions are syntactically-null.  They
       simply include each 1-gram (or 2-gram) without any syntactic
@@
   \end{description}
-\item[Forest size] Although the parser can generate considerably more,
-  we consider forests of size 1 and of size 50.
-\item[Forest confidence] An expectation requires a probability
+\item[Size of $n$-best list.] Although the parser can generate considerably more,
+  we consider $n$-best lists  of size 1 and of size 50.
+\item[Confidence of $n$-best list.] An expectation requires a probability
   distribution over the $n$-best list, and we consider three options:
@@
     the $n$-best.}
-\item[score combination]: We consider three possible methods for
+\item[Score combination.]: We consider three possible methods for
   combining scores across decompositions, as suggested in
@@
     contributing $n$ more items to the bag of tokens extracted from
-    the tree, ad then computes a global $F$ score.
+    the tree, and then computes a global $F$ score.
   \item[$\mu_{PR}\lbrack\cdot\rbrack$] constructs a precision and
@@
   correlations against pairwise differences among translation segments
-  derived from the same source segment. Those results are similar to
+  derived from the same source segment. The results are similar to
   those reported here. However, we believe that pairwise differences
@@
 
+%% MO: Okay, but a little hard to figure out.
 When reporting HTER correlations per-sentence, we use length-weighted

### 2009-05-13 mtjournal.tex
Mari's notes on HTER comparison sections now included

@@
 
+%% MO: I still think this is a bit weird given that what you are
+%% adding is structure. If time, you might want to look at some
+%% example sentences
 Table~\ref{tab:hterpersent} presents per-sentence correlations,
@@
 
+%% MO: important to describe basic model before listing features.
 TO DO: TERp optimizer [Matt, a brief description here]
@@
   synonym-alignment. The \textbf{T} set includes 7 features.
+  % MO  (to Matt): is this basic TER or basic TERp?
   % TO DO: what are the other two?
@@
   wordsbrevity/prolixity (2 features: one for longer-than-ref, one for
-  shorter-than-ref)
+  shorter-than-ref).
+  % MO: out of curiosity, why do you need 2 features?
+  We assign two features here to let the system separately penalize
+  ``too short'' translations and ``too long''.
 \end{description}
@@
 
+% MO suggestions:
+
+%% do n-fold with >2 N -- better robustness to variation?
+
+%% look at variation in weights on 2-fold case   
 \begin{verbatim}

### 2009-05-13 mtjournal.tex
mari's notes on conclusion

@@
 we settle on EDPM, a member of that family with promising predictive
-power.  When tested against correlations with HTER, we find that EDPM
-is as good a predictor of HTER as current measures BLEU$_4$ and TER,
-and --- depending on the language and genre --- often better.
+power.  We find that EDPM is superior to BLEU and TER in terms of
+correlation with human judgments and as a per-document and
+per-sentence predictor of HTER (using mean-subtraction to account for
+underlying document- and segment-level difficulty).
+% TO DO:  add something here regarding summary of tuning?
 
-Several areas of future work are open to us. Using these syntactic
+Several areas of future work are open to us. One obstacle to using
+this method widely is the computational expense of the system. Parsing
+is many times more computationally costly than word-based measures
+such as BLEU or TER.  Using these syntactic
 techniques may be possible as a late-pass evaluation, to identify how
@@
 direction explores the intersection of genre (and genre-mismatch),
-parser quality, and prediction of human performance. Another
-consideration is the speed of the parser: parse time currently takes
-much, much longer than BLEU, TER, or other word-based measures. We are
+parser quality, and prediction of human performance. 
+% Another
+% consideration is the speed of the parser: parse time currently takes
+% much, much longer than BLEU, TER, or other word-based measures. 
+We are
 interested in the trade-offs --- if any --- among parse quality,

### 2009-05-13 mtjournal.tex
whitespace and linewrap cleanup

@@
 Currently, the most popular evaluation measures include a measure
-based on $n$-gram precision known as BLEU \cite{papineni02bleu}) and the edit-distance measure Translation Edit
-Rate (TER) \cite{snover06ter}.  Recent research has found that these
-measures may not accurately track translation quality both empirically
+based on $n$-gram precision known as BLEU \cite{papineni02bleu}) and
+the edit-distance measure Translation Edit Rate (TER)
+\cite{snover06ter}.  Recent research has found that these measures may
+not accurately track translation quality both empirically
 \cite{charniak03syntaxlmmt} and theoretically
@@
 dependency trees.  The dependent-based SParseval measure
-\cite{roark06:sparseval}, designed as a parse-quality
-metric for speech, is a similar approach, in that it is an F-measure
-over a decomposition of reference and hypothesis trees.
+\cite{roark06:sparseval}, designed as a parse-quality metric for
+speech, is a similar approach, in that it is an F-measure over a
+decomposition of reference and hypothesis trees.
 \inlinecite{owczarzak07labelleddepseval} proposed \textbf{d} and
@@
 This work extends the dependency-scoring strategies of
-\inlinecite{roark06:sparseval} and \inlinecite{owczarzak07labelleddepseval} using a
-widely-used and publically available PCFG parser and deterministic
-head-finding rules.
+\inlinecite{roark06:sparseval} and
+\inlinecite{owczarzak07labelleddepseval} using a widely-used and
+publically available PCFG parser and deterministic head-finding rules.
 
@@
 have evaluated their success by comparing the measure to human
-judgments of fluency and adequacy.  In other work,
-e.g. \inlinecite{snover06ter}, measures are evaluated by comparison to
+judgments of fluency and adequacy.  In other work, e.g.\
+\inlinecite{snover06ter}, measures are evaluated by comparison to
 human-targeted TER (HTER), a distance to a human-revised reference
@@
 
-In section~\ref{sec:approach}), we outline our
-approach to including syntactic information in MT evaluation, by
-describing a family of metrics. In section~\ref{sec:paradigm}, we
-describe the implementation details and make explicit which dimensions
-of this family of metrics we explore in this paper.
-Section~\ref{sec:faexpts} examines the correlation of several members
-of this family with human judgments of fluency and adequacy over a
-corpus of sentence translations and selects a favorable member of the
-family, Expected Dependency Pair Match (EDPM), based on this comparison. Section~\ref{sec:hter1} compares
-the correlation of EDPM with HTER over a corpus of document- and
-sentence-level comparisons, and demonstrates that its correlation is
-favorable to that of competitor measures BLEU and TER at both sentence
-and document levels, regardless of source-language or genre.
+In section~\ref{sec:approach}), we outline our approach to including
+syntactic information in MT evaluation, by describing a family of
+metrics. In section~\ref{sec:paradigm}, we describe the implementation
+details and make explicit which dimensions of this family of metrics
+we explore in this paper.  Section~\ref{sec:faexpts} examines the
+correlation of several members of this family with human judgments of
+fluency and adequacy over a corpus of sentence translations and
+selects a favorable member of the family, Expected Dependency Pair
+Match (EDPM), based on this comparison. Section~\ref{sec:hter1}
+compares the correlation of EDPM with HTER over a corpus of document-
+and sentence-level comparisons, and demonstrates that its correlation
+is favorable to that of competitor measures BLEU and TER at both
+sentence and document levels, regardless of source-language or genre.
 Section~\ref{sec:hter2} uses the tuning components of the TERp tools
-to explore the combination of these syntactic measures with other
-TERp word-based measures. Finally, section~\ref{sec:conclusion}
-concludes with some remarks on challenges and future work.
+to explore the combination of these syntactic measures with other TERp
+word-based measures. Finally, section~\ref{sec:conclusion} concludes
+with some remarks on challenges and future work.
 
@@
 extensions of the dependency-pair F measures found in
-\inlinecite{roark06:sparseval} and \inlinecite{owczarzak07labelleddepseval}.  Rather than
-comparing string sequences, as BLEU does with its $n$-gram precision,
-this approach defers to a parser for an indication of the relevant
-word tuples associated with meaning --- in these
-implementations, the head on which that word depends.  Each sentence
-(both reference and hypothesis) is converted to a labeled syntactic
-dependency tree and then relations from each tree are extracted and
-compared.
+\inlinecite{roark06:sparseval} and
+\inlinecite{owczarzak07labelleddepseval}.  Rather than comparing
+string sequences, as BLEU does with its $n$-gram precision, this
+approach defers to a parser for an indication of the relevant word
+tuples associated with meaning --- in these implementations, the head
+on which that word depends.  Each sentence (both reference and
+hypothesis) is converted to a labeled syntactic dependency tree and
+then relations from each tree are extracted and compared.
 
@@
   $\frac{1}{4}$: thus $F[dlh] = \frac{2}{7}$
-  \caption{Example hypothesis and reference dependency trees and the $dlh$ decomposition of each.}
+  \caption{Example hypothesis and reference dependency trees and the
+    $dlh$ decomposition of each.}
   \label{fig:decompexample}
@@
 
-When multiple decomposition types are used together, we may combine these
-subscores in a variety of ways.  We may compute precision and recall
-subscores for each decomposition separately, or, since the results of
-each decomposition are of different types entirely, we may compute
-them as members of one large bag for an even simpler F score.  These
-two approaches are equivalent when only one decomposition type is
-included.  For simplicity in presentation, we use the following
+When multiple decomposition types are used together, we may combine
+these subscores in a variety of ways.  We may compute precision and
+recall subscores for each decomposition separately, or, since the
+results of each decomposition are of different types entirely, we may
+compute them as members of one large bag for an even simpler F score.
+These two approaches are equivalent when only one decomposition type
+is included.  For simplicity in presentation, we use the following
 notation, where $dl$ and $lh$ represent the two kinds of
@@
 Dependency-based SParseval and the \textbf{d} approach from
-\inlinecite{owczarzak07labelleddepseval} may
-each be understood as $F[dlh]$, while the latter's \textbf{d\_var}
-method may be understood as something close to $F[dl,lh]$.
+\inlinecite{owczarzak07labelleddepseval} may each be understood as
+$F[dlh]$, while the latter's \textbf{d\_var} method may be understood
+as something close to $F[dl,lh]$.
 
 Both of the combination methods above ($F[\cdot]$ and
-$\mu_{PR}[\cdot]$) are ``naive'' in that they treat each
-component score as equivalent to the next.  One further direction to
-explore involves tuning (presumably on held-out data) the precision
-and recall weights for each decomposition type for a later linear
-combination.
+$\mu_{PR}[\cdot]$) are ``naive'' in that they treat each component
+score as equivalent to the next.  One further direction to explore
+involves tuning (presumably on held-out data) the precision and recall
+weights for each decomposition type for a later linear combination.
 
@@
 In this work, we use a state-of-the-art PCFG (the first stage of
-\inlinecite{charniak-johnson:2005:ACL}) and context-free head-finding rules
-\cite{magerman95headfinding} to generate a 50-best list of dependency
-trees for each hypothesis and reference translation.  We use the
-parser's default (English-language) Wall Street Journal training
-parameters.  Head-finding uses the Charniak parser's rules, with three
-modifications: prepositional and complementizer phrases choose nominal
-and verbal heads respectively (rather than functional heads) and
-auxiliary verbs are modifiers of main verbs (rather than the
-converse).
+\inlinecite{charniak-johnson:2005:ACL}) and context-free head-finding
+rules \cite{magerman95headfinding} to generate a 50-best list of
+dependency trees for each hypothesis and reference translation.  We
+use the parser's default (English-language) Wall Street Journal
+training parameters.  Head-finding uses the Charniak parser's rules,
+with three modifications: prepositional and complementizer phrases
+choose nominal and verbal heads respectively (rather than functional
+heads) and auxiliary verbs are modifiers of main verbs (rather than
+the converse).
 
@@
 \end{figure}
-For example, in figure~\ref{fig:depextract}, the S node is the
-lowest node headed by \emph{stumbled} that dominates \emph{cat}, and
-the NP node is the highest constituent label headed by \emph{cat}, so
-the arc between \emph{cat} and \emph{stumbled} is labelled \arclabel{S/NP}.
+For example, in figure~\ref{fig:depextract}, the S node is the lowest
+node headed by \emph{stumbled} that dominates \emph{cat}, and the NP
+node is the highest constituent label headed by \emph{cat}, so the arc
+between \emph{cat} and \emph{stumbled} is labelled \arclabel{S/NP}.
 %
@@
   \end{description}
-\item[Size of $n$-best list.] Although the parser can generate considerably more,
-  we consider $n$-best lists  of size 1 and of size 50.
-\item[Confidence of $n$-best list.] An expectation requires a probability
-  distribution over the $n$-best list, and we consider three options:
-  uniform, the parser probabilities, and a flattened version of the
-  parser probabilities such that $\tilde{p}(x) =
+\item[Size of $n$-best list.] Although the parser can generate
+  considerably more, we consider $n$-best lists of size 1 and of size
+  50.
+\item[Confidence of $n$-best list.] An expectation requires a
+  probability distribution over the $n$-best list, and we consider
+  three options: uniform, the parser probabilities, and a flattened
+  version of the parser probabilities such that $\tilde{p}(x) =
   \frac{p(x)^\gamma}{\sum_ip(i)^\gamma}$ (where $\gamma$ is a free
@@
   \textbf{d\_50} and \textbf{d\_50\_var} measures.\footnote{Since
-    \inlinecite{owczarzak07labelleddepseval} report no use of parse weights,
-    \textbf{d\_50} and \textbf{d\_50\_var} may be using a sum of
-    counts over the 50-best list rather than expected-counts over a
+    \inlinecite{owczarzak07labelleddepseval} report no use of parse
+    weights, \textbf{d\_50} and \textbf{d\_50\_var} may be using a sum
+    of counts over the 50-best list rather than expected-counts over a
     uniform distribution. These two approaches are equivalent --- so
@@
 identifying which of two candidate translations is better with $m$'s
-(and HTER's) ability to distinguish which source segments are
-more difficult to translate.
+(and HTER's) ability to distinguish which source segments are more
+difficult to translate.
 
 To avoid this conflation, on the HTER corpora we report correlations
-on $\bar{\mu}(m)$
-and $\bar{\mu}(\textrm{HTER})$ (rather than measure $m$ and HTER), where
-$\bar{\mu}$ represents mean-removal:
+on $\bar{\mu}(m)$ and $\bar{\mu}(\textrm{HTER})$ (rather than measure
+$m$ and HTER), where $\bar{\mu}$ represents mean-removal:
 \begin{equation}
@@
   those reported here. However, we believe that pairwise differences
-  introduce problems with the independence assumptions in
-  the Pearson's $r$ tests, and we use mean-removal
-  throughout this work instead.}
+  introduce problems with the independence assumptions in the
+  Pearson's $r$ tests, and we use mean-removal throughout this work
+  instead.}
 
@@
 different reference translation.  For a rough\footnote{Our segment
-  count differs from \inlinecite{owczarzak07labelleddepseval}, who report
-  16,800 segments over the same corpus. We find baseline correlations
-  (BLEU$_4$ and TER) lower than those reported there as well, so the
-  results presented here are not directly comparable with that paper,
-  though we demonstrate similar gains over those baselines in
-  essentially the same corpus. % (section~\ref{sec:facorr:newdeps}).}
-}
-comparison with \inlinecite{owczarzak07labelleddepseval}, we treat each
-judgment as a separate segment.
+  count differs from \inlinecite{owczarzak07labelleddepseval}, who
+  report 16,800 segments over the same corpus. We find baseline
+  correlations (BLEU$_4$ and TER) lower than those reported there as
+  well, so the results presented here are not directly comparable with
+  that paper, though we demonstrate similar gains over those baselines
+  in essentially the same
+  corpus. % (section~\ref{sec:facorr:newdeps}).}
+} comparison with \inlinecite{owczarzak07labelleddepseval}, we treat
+each judgment as a separate segment.
 %
@@
 \end{table}
-For the $n=50$ cases, we set $\gamma=0$ to assign
-uniform probabilities to the parse forest, to compare as closely as
-possible to \inlinecite{owczarzak07labelleddepseval}, which includes a
+For the $n=50$ cases, we set $\gamma=0$ to assign uniform
+probabilities to the parse forest, to compare as closely as possible
+to \inlinecite{owczarzak07labelleddepseval}, which includes a
 \textbf{d\_var\_50} measure with 50-best parses, with ranks but no
@@
 trend is universally that the correlation $r$ improves as the number
-of parses ($n$) is increased.
-Tuning experiments find that increasing $\gamma$ to 0.25 can increase
-the $r$ reported here for $F[1g,2g,dl,lh]$ marginally (but insignificantly).
+of parses ($n$) is increased.  Tuning experiments find that increasing
+$\gamma$ to 0.25 can increase the $r$ reported here for
+$F[1g,2g,dl,lh]$ marginally (but insignificantly).
 
@@
 %
-The corpus being translated comes from
-Arabic and Chinese in four genres: \texttt{bc} (broadcast
-conversation), \texttt{bn} (broadcast news), \texttt{nw} (newswire),
-and \texttt{wb} (web text), with corpus sizes shown in
-table~\ref{tab:galestats}.
+The corpus being translated comes from Arabic and Chinese in four
+genres: \texttt{bc} (broadcast conversation), \texttt{bn} (broadcast
+news), \texttt{nw} (newswire), and \texttt{wb} (web text), with corpus
+sizes shown in table~\ref{tab:galestats}.
 \begin{table}
@@
 corresponding human-targeted reference aligned at the sentence level,
-so we have available the HTER score of each segment
-HTER$(t_{i,z})$ at both the sentence and document level.
+so we have available the HTER score of each segment HTER$(t_{i,z})$ at
+both the sentence and document level.
 
 As discussed in section~\ref{sec:paradigm}, we report correlations
-here between $\bar{\mu}(m)$ and $\bar{\mu}(\textrm{HTER})$, rather than
-$m$ and HTER directly, to abstract away from variations in the
+here between $\bar{\mu}(m)$ and $\bar{\mu}(\textrm{HTER})$, rather
+than $m$ and HTER directly, to abstract away from variations in the
 underlying documents.
@@
 \item[N]: the non-syntactic features from the DPM family;
-  specifically, precision and recall for $1g$ and $2g$
-  decompositions.  The \textbf{N} set includes 4 features.
+  specifically, precision and recall for $1g$ and $2g$ decompositions.
+  The \textbf{N} set includes 4 features.
 \item[T]: the features from basic TERp, which include a separate cost
@@
   % TO DO: what are the other two?
-\item[P]: features from the automatically-derived TERp paraphrase table
-  (4 features)  % TO DO: Matt, what are these
+\item[P]: features from the automatically-derived TERp paraphrase
+  table (4 features) % TO DO: Matt, what are these
 \item[B]: two ``length penalties'' one is assigned the count of
@@
 These two corpus halves are used in two-fold cross-validation, testing
-on one and training on the other.  We tune on (and report correlation with)
-length-weighted per-sentence correlation with HTER, using mean-removed
-scores as before.
+on one and training on the other.  We tune on (and report correlation
+with) length-weighted per-sentence correlation with HTER, using
+mean-removed scores as before.
 
@@
 
-point: E+T $>$ T, E+TB $>$ TB, E+TP $>$ TP; information is available in
-syntax that is not captured in the other measures.
+point: E+T $>$ T, E+TB $>$ TB, E+TP $>$ TP; information is available
+in syntax that is not captured in the other measures.
 
@@
 
-talk about why EN is < EDPM (is it?)
+talk about why EN is $<$ EDPM (is it?)
 
@@
 is many times more computationally costly than word-based measures
-such as BLEU or TER.  Using these syntactic
-techniques may be possible as a late-pass evaluation, to identify how
-candidate translation systems are performing overall.  Alternatively,
-the various tree-decompositions of the DPM family could be used as
-system diagnostics, by looking at relative quality of these component
-scores compared to those of an alternative candidate system.
+such as BLEU or TER.  Using these syntactic techniques may be possible
+as a late-pass evaluation, to identify how candidate translation
+systems are performing overall.  Alternatively, the various
+tree-decompositions of the DPM family could be used as system
+diagnostics, by looking at relative quality of these component scores
+compared to those of an alternative candidate system.
 
@@
 % much, much longer than BLEU, TER, or other word-based measures. 
-We are
-interested in the trade-offs --- if any --- among parse quality,
-speed, and predictive power as a translation metric.  One possibility,
-for example, is to store packed forests \cite{huang08packedforests}
-rather than generating an $n$-best list only to sum across it again in
-calculating the expectation.
+We are interested in the trade-offs --- if any --- among parse
+quality, speed, and predictive power as a translation metric.  One
+possibility, for example, is to store packed forests
+\cite{huang08packedforests} rather than generating an $n$-best list
+only to sum across it again in calculating the expectation.
 
@@
 
-{\small
-This material is based upon work supported by the National Science
-Foundation under Grant No. 0741585 and the Defense Advanced Research
-Projects Agency under Contract Nos. HR0011-06-C-0022 and HR0011-06-C-0023.
+{\small This material is based upon work supported by the National
+  Science Foundation under Grant No. 0741585 and the Defense Advanced
+  Research Projects Agency under Contract Nos. HR0011-06-C-0022 and
+  HR0011-06-C-0023.
 
@@
   reflect the views of the National Science Foundation or the Defense
-Advanced Research Projects Agency.
-}
+  Advanced Research Projects Agency.  }
 

### 2009-05-14 mtjournal.tex
small changes from Mari's e-mail

@@
 \title{Expected Dependency Pair Match:\\
-Predicting {HTER} improvements with expected syntactic structure} 
+Predicting translation quality with expected syntactic structure} 
 \author{Jeremy G. \surname{Kahn}\email{jgk@u.washington.edu}}  
@@
 Currently, the most popular evaluation measures include a measure
-based on $n$-gram precision known as BLEU \cite{papineni02bleu}) and
+based on $n$-gram precision known as BLEU \cite{papineni02bleu} and
 the edit-distance measure Translation Edit Rate (TER)
@@
   50.
-\item[Confidence of $n$-best list.] An expectation requires a
+\item[Confidence of $n$-best hypotheses.] An expectation requires a
   probability distribution over the $n$-best list, and we consider
@@
     the $n$-best.}
-\item[Score combination.]: We consider three possible methods for
+\item[Score combination.] We consider three possible methods for
   combining scores across decompositions, as suggested in
@@
 When reporting HTER correlations per-sentence, we use length-weighted
-correlation, since unweighted correlations effectively puts too much
+correlation, since unweighted correlations effectively put too much
 emphasis on short sentences.  We do not weight for length for the
@@
 \item[E]: the fully syntactic features from the DPM family.
-  Specifically, this featureset includes precision and recall scores
+  Specifically, this feature set includes precision and recall scores
   for the $dl$, $lh$, $dlh$, and $dh$ decompositions of the syntactic

### 2009-05-14 mtjournal.tex
changes from Matt.

@@
 operations. TERp's feature list extends TER to additionally include
-synonym tables and an empirically-derived paraphrase table.
+synonymy, stemming, and automatically-derived paraphrases.
 
 %% MO: important to describe basic model before listing features.
-TO DO: TERp optimizer [Matt, a brief description here]
+The TERp optimizer performs a hill-climbing search, with randomized
+restarts, to maximize the correlation of a linear combination of the
+features with a set of human judgments.  Within the TERp framework the
+features are the count of the various edit types normalized for the
+length of the reference so that the dot product of the features with a
+vector of edit costs will give the TERp score for a given segment.  In
+these experiments, we follow~\inlinecite{snover09terp} by first
+aligning the MT outputs with the reference, using TERp with a set of
+default edit costs, to find the counts of the various edit types.
+These alignments and edit costs are then held constant to optimize the
+edit costs. In the case of comparing TERp without the paraphrasing
+feature, a separate alignment was performed without that feature
+enabled.  Additional features can be added to the TERp optimization
+framework and optimized in the same manner as the edit costs.
 
@@
   The \textbf{N} set includes 4 features.
-\item[T]: the features from basic TERp, which include a separate cost
-  for each of insertions, deletions, substitutions, shifts, and
-  synonym-alignment. The \textbf{T} set includes 7 features.
-  % MO  (to Matt): is this basic TER or basic TERp?
-  % TO DO: what are the other two?
+\item[T]: the features from basic TERp, excluding paraphrases, which include a separate cost
+  for each of matches, insertions, deletions, substitutions, shifts, and
+  synonym matches and stem matches. The \textbf{T} set includes 7 features.
 \item[P]: features from the automatically-derived TERp paraphrase
-  table (4 features) % TO DO: Matt, what are these
+  table. These are four features which are a function of of the edit
+  distance between the two phrases in each paraphrase and the
+  probability of the paraphrase.
 \item[B]: two ``length penalties'' one is assigned the count of

### 2009-05-15 mtjournal.tex
changes to include tuning plot in text. much text written.

@@
 \item[E]: the fully syntactic features from the DPM family.
-  Specifically, this feature set includes precision and recall scores
+  Specifically, this feature set includes error-counts
+  ($\frac{|\textrm{hyp}| - |\textrm{correct}|}{|\textrm{ref}|}$ and
+  $\frac{|\textrm{ref}| - |\textrm{correct}|}{|\textrm{ref}|}$) 
   for the $dl$, $lh$, $dlh$, and $dh$ decompositions of the syntactic
@@
 \item[N]: the non-syntactic features from the DPM family;
-  specifically, precision and recall for $1g$ and $2g$ decompositions.
+  specifically, error-counts for the $1g$ and $2g$ decompositions.
   The \textbf{N} set includes 4 features.
@@
 % tuning target: weighted correlation with mean-removed segment HTER
-
+\subsection{Effects of features on tuning}
 % [we use weighted correlation to avoid emphasis on short segments
@@
 \begin{figure}
-  \centerline{{one figure}\qquad{}{two figure}}
-  \caption{Figure with two tune-set plots will go here.}
+  \begin{center}
+    \includegraphics[scale=0.45]{tuning.eps}
+  \end{center}
+  \caption{Pearson's $r$ values for the two halves of the training
+    set, with 95\% confidence intervals. }
   \label{fig:tuneresults}
 \end{figure}
-
+%
+For each set of features, we plot the $r$ (and confidence) on both
+halves of the corpus (for each half, tuned on the other half). In the
+leftmost group of the figure, we include baselines: EDPM and BLEU and
+TER.  Next, the tuned E feature, which (by itself) does not perform as
+well as TER or EDPM.  
+%
+The remaining groups of three show that for each set of features, as E
+(and EB) are included with others, correlation improves (though not
+significantly).  These all move in the same direction. By the same
+measurement, P features help beyond T -- as reported elsewhere -- but
+N features do not seem to find a better correlation than TP features.
 % MO suggestions:
 
+Overall, we find that this approach for combining the underlying idea
+of EDPM with the underlying ideas of TERp is promising, in that the E
+features offer small (though non-significant) improvements to the
+features already within TERp. However, this combination approach still
+has more in common with the $\mu_{PR}$ combination method (explored in
+section~\ref{sec:faexpts}) than with EDPM, which suggests a reason
+that the EN features --- though tuned --- do not outperform EDPM.
 %% do n-fold with >2 N -- better robustness to variation?
-
-%% look at variation in weights on 2-fold case   
-\begin{verbatim}
-   == Tuned For Weighted Mean Removed Pearson
-     - Examining WGT_MEAN_RM_SEG_PEAR ==
-            Average
-      Cond   Test  
-     ---------------
-      ETPB   0.4096
-      ETP    0.4038
-      TPB    0.4010
-      TP     0.3946  > 0.416
-      TPNB   0.3940
-      TPN    0.3891
-      ETB    0.3871
-      ET     0.3827  > 0.404
-      TB     0.3815  > 0.403
-      ETNB   0.3814
-      ETN    0.3804
-      T      0.3758  > 0.398
-      TNB    0.3708
-      TN     0.3680  > 0.390
-      TER    0.353   > 0.375
-      E      0.3287  > 0.352
-      EB     0.3267
-      ENB    0.3149
-      NB     0.3058  > 0.329
-      EN     0.2973
-      N      0.2636
-      BLEU   0.244
-      B      0.1638
-\end{verbatim}
-(maybe don't report all these numbers!)
-
-TO DO: discuss this as a trend, despite lack of significance? plot charts?
-
-point: E+T $>$ T, E+TB $>$ TB, E+TP $>$ TP; information is available
-in syntax that is not captured in the other measures.
-
-point: syntax is *not* just an expensive way to get at n-grams;
-N-grams are not the same -- TP $>$ TP+N, but TP $<$ TP+E
-
-question: this is kind of like the $\mu_{PR}[\cdot]$ --- seems like it
-also doesn't work so well. compare to EDPM as-is
-
-talk about why EN is $<$ EDPM (is it?)
 

### 2009-05-15 mtjournal.tex
converted footnotes to endnotes.
still to do: endnotes must read sensibly out-of-context.

@@
   be equivalent to the \cite{owczarzak07labelleddepseval}
-  \textbf{d\_50} and \textbf{d\_50\_var} measures.\footnote{Since
+  \textbf{d\_50} and \textbf{d\_50\_var} measures.\endnote{Since
     \inlinecite{owczarzak07labelleddepseval} report no use of parse
@@
 translations rather than among differences in the underlying segments.
-\footnote{ Previous work \cite{kahn08metricsmatr} reported HTER
+\endnote{ Previous work \cite{kahn08metricsmatr} reported HTER
   correlations against pairwise differences among translation segments
@@
 
-% TO DO: more to say here?
-
 \subsection{Corpus}
@@
 sentence, with each judgment using a different human judge and a
-different reference translation.  For a rough\footnote{Our segment
+different reference translation.  For a rough\endnote{Our segment
   count differs from \inlinecite{owczarzak07labelleddepseval}, who
@@
 $\langle$hypothesis, reference$\rangle$ and examine the
-correlations\footnote{The independence of each of these segments is
+correlations\endnote{The independence of each of these segments is
   questionable, since the same hypothesis translations are used in
@@
 
-% \theendnotes
+\theendnotes
 

### 2009-05-15 mtjournal.tex
\bar{\mu} now always used without following parens
per Mari's suggestion

@@
 To avoid this conflation, on the HTER corpora we report correlations
-on $\bar{\mu}(m)$ and $\bar{\mu}(\textrm{HTER})$ (rather than measure
+on $\bar{\mu}m$ and $\bar{\mu}\textrm{HTER}$ (rather than measure
 $m$ and HTER), where $\bar{\mu}$ represents mean-removal:
@@
   \label{eq:meansub}
-  \bar{\mu}(m(t_i)) = m(t_i) - \sum_{j=1}^I\frac{m(t_j)}{I}
+  \bar{\mu}m(t_i) = m(t_i) - \sum_{j=1}^I\frac{m(t_j)}{I}
 \end{equation}
@@
 As discussed in section~\ref{sec:paradigm}, we report correlations
-here between $\bar{\mu}(m)$ and $\bar{\mu}(\textrm{HTER})$, rather
+here between $\bar{\mu}m$ and $\bar{\mu}\textrm{HTER}$, rather
 than $m$ and HTER directly, to abstract away from variations in the

### 2009-05-15 mtjournal.tex
Mari's last pass of changes.

@@
 leftmost group of the figure, we include baselines: EDPM and BLEU and
-TER.  Next, the tuned E feature, which (by itself) does not perform as
-well as TER or EDPM.  
+TER (which are not tuned).  Next, the tuned E feature, which (by
+itself) does not perform as well as TER or EDPM.
 %
@@
 In this research, we explore dependency pair match (DPM) a family of
-syntactically-oriented measures.  Through a corpus of human judgments,
+syntactically-oriented measures.  Through a corpus of human fluency
+and adequacy judgments,
 we settle on EDPM, a member of that family with promising predictive
@@
 underlying document- and segment-level difficulty).
-% TO DO:  add something here regarding summary of tuning?
+We experiment with including syntactic features in TERp-style tuning,
+with linear weight combination, and find that including these features
+improves correlation with mean-subtracted HTER, although these
+improvements do not reach significance.
 
 Several areas of future work are open to us. One obstacle to using
-this method widely is the computational expense of the system. Parsing
-is many times more computationally costly than word-based measures
+this method widely is the computational cost of parsing compared to
+word-based measures
 such as BLEU or TER.  Using these syntactic techniques may be possible
@@
 compared to those of an alternative candidate system.
+%
+Another possible approach is to store packed forests
+\cite{huang08packedforests} rather than generating an $n$-best list
+only to sum across it again in calculating the expectation.
 
-Another interesting research area is in the quality of the parser.  In
+Another research area relates to the quality of the parser.  In
 this work, we demonstrated that the
@@
 \inlinecite{owczarzak07labelleddepseval}. PCFG parsers, however, are
-fairly easily adaptable to another domain. One interesting research
+fairly easily adaptable to another domain. One research
 direction explores the intersection of genre (and genre-mismatch),
@@
 We are interested in the trade-offs --- if any --- among parse
-quality, speed, and predictive power as a translation metric.  One
-possibility, for example, is to store packed forests
-\cite{huang08packedforests} rather than generating an $n$-best list
-only to sum across it again in calculating the expectation.
-
-TO DO: parting words
+quality, genre adaptation, and predictive power as a translation metric.
 

### 2009-05-15 mtjournal.tex
changes from Matt (suggested by phone)

@@
 $n$-gram precision in various configurations of constituency and
-dependency trees.  The dependent-based SParseval measure
+dependency trees.  The dependency-based SParseval measure
 \cite{roark06:sparseval}, designed as a parse-quality metric for
@@
     \item[dh] $\left\langle \textrm{Dependent}, \textrm{Head}
-      \right\rangle$ Here, drops the arc-label. This decomposition
-      ignores the syntactic-role information implied in the arc labels.
+      \right\rangle$ This decomposition drops the arc from the
+      dependency-pair tuple, ignoring the syntactic-role information
+      implied in the arc labels.
     \item[1g,2g] These decompositions are syntactically-null.  They
@@
     \hline
-    metric  &    \multicolumn{1}{c}{$r$} \\
+    metric  &    \multicolumn{1}{c}{$|r|$} \\
     \hline
@@
     $F[dlh]$ &     0.185 \\
-    TER &      -0.173 \\
+    TER &      0.173 \\
     \hline
@@
     $F[1g,2g]$            & 0.227 \\
-    $F[1g,dl,dlh]$        & 0.227 \\
+    $\mu_{PR}[1g,2g]$            & 0.215 \\
+%    $F[1g,dl,dlh]$        & 0.227 \\
     \rlcline{1-1} \rlcline{2-2}
@@
 correlation in $F[1g,2g,dl,lh]$. Including progressively larger chunks
-of the dependency graph (as in $F[1g,dl,dlh]$, which is inspired by
+of the dependency graph with $F[1g,dl,dlh]$ (not shown in table, but inspired by
 the BLEU$_k$ idea of progressively larger $n$-grams) does not seem to

### 2009-05-15 mtjournal.tex
made endnotes more readable as endnotes, updated to-do

@@
 \subsection{Dimensions of metric family}
+\label{sec:metricdimensions}
 In the experiments presented later in this article, we explore the
@@
   be equivalent to the \cite{owczarzak07labelleddepseval}
-  \textbf{d\_50} and \textbf{d\_50\_var} measures.\endnote{Since
+  \textbf{d\_50} and \textbf{d\_50\_var} measures.\endnote{
     \inlinecite{owczarzak07labelleddepseval} report no use of parse
-    weights, \textbf{d\_50} and \textbf{d\_50\_var} may be using a sum
-    of counts over the 50-best list rather than expected-counts over a
-    uniform distribution. These two approaches are equivalent --- so
-    long as the $n$-best list is always the same length for hypothesis
-    and reference.  In our implementation, the $n$-best list does not
-    always reach 50 candidate parses on short sentences, so the
-    expectation matches our intent better than a sum of counts over
-    the $n$-best.}
+    confidence weights, thus \textbf{d\_50} and \textbf{d\_50\_var}
+    may be using a sum of counts over the 50-best list rather than
+    expected-counts over a uniform distribution, as we suggest in
+    section~\ref{sec:metricdimensions}. These two approaches are
+    equivalent --- so long as the $n$-best list is always the same
+    length for hypothesis and reference.  In our implementation, the
+    $n$-best list does not always reach 50 candidate parses on short
+    sentences, so the expectation matches our intent better than a sum
+    of counts over the $n$-best.}
 \item[Score combination.] We consider three possible methods for
@@
 \subsection{Reference evaluation measures over parallel translations}
-
+\label{sec:metricsreference}
 For some of the experiments presented here (sections~\ref{sec:hter1}
@@
 translations rather than among differences in the underlying segments.
-\endnote{ Previous work \cite{kahn08metricsmatr} reported HTER
+\endnote{Previous work \cite{kahn08metricsmatr} reported HTER
   correlations against pairwise differences among translation segments
-  derived from the same source segment. The results are similar to
-  those reported here. However, we believe that pairwise differences
+  derived from the same source segment, unlike the mean-removal
+  suggested in section~\ref{sec:metricsreference}. Those results are similar to
+  those reported in section~\ref{sec:hter1}. However, we believe that pairwise differences
   introduce problems with the independence assumptions in the
@@
 \subsection{Corpus}
-
+\label{sec:facorpus}
 For these experiments, we use LDC Multiple Translation Chinese corpus
@@
 different reference translation.  For a rough\endnote{Our segment
-  count differs from \inlinecite{owczarzak07labelleddepseval}, who
-  report 16,800 segments over the same corpus. We find baseline
-  correlations (BLEU$_4$ and TER) lower than those reported there as
-  well, so the results presented here are not directly comparable with
-  that paper, though we demonstrate similar gains over those baselines
-  in essentially the same
-  corpus. % (section~\ref{sec:facorr:newdeps}).}
+  count reported in section~\ref{sec:facorpus} differs from
+  \inlinecite{owczarzak07labelleddepseval}, who report 16,800 segments
+  over the same corpus. We find baseline correlations (BLEU$_4$ and
+  TER) lower than those reported there as well, so the results
+  presented here are not directly comparable with that paper, though
+  we demonstrate similar gains over those baselines in essentially the
+  same corpus.
 } comparison with \inlinecite{owczarzak07labelleddepseval}, we treat
@@
 $\langle$hypothesis, reference$\rangle$ and examine the
-correlations\endnote{The independence of each of these segments is
-  questionable, since the same hypothesis translations are used in
-  multiple items, but for the sake of methodological comparison with
-  prior work, this strategy is preserved.}  between those automatic
-scores and the arithmetic mean of the fluency and adequacy measures.
+correlations\endnote{The independence of the translation segments in
+  the Multiple Translation Corpus described in
+  section~\ref{sec:facorpus} is questionable, since the same
+  hypothesis translations are used in multiple items, but for the sake
+  of methodological comparison with prior work, this strategy is
+  preserved.}  between those automatic scores and the arithmetic mean
+of the fluency and adequacy measures.
 

### 2009-05-16 mtjournal.tex
Matt's edits

@@
 research has turned increasingly towards automatic measures of
-translation quality, usually by comparing the system translation to
+translation quality, usually by comparing the system (hypothesis) translation to
 one or more reference (human) translations.
@@
 tuned by users.  TERp \cite{snover09terp} is an extension of the
-previously-mentioned TER that also incorporates synonym sets, along
+previously-mentioned TER that also incorporates synonym sets and stemming, along
 with automatically-derived paraphrase tables.  TERp is explicitly
@@
 
-In section~\ref{sec:approach}), we outline our approach to including
+In section~\ref{sec:approach}, we outline our approach to including
 syntactic information in MT evaluation, by describing a family of
@@
 sentence and document levels, regardless of source-language or genre.
-Section~\ref{sec:hter2} uses the tuning components of the TERp tools
-to explore the combination of these syntactic measures with other TERp
+Section~\ref{sec:hter2} uses the tuning components of TERp
+to explore the combination of these syntactic measures with TERp's
 word-based measures. Finally, section~\ref{sec:conclusion} concludes
@@
 \end{figure}
-Figure~\ref{fig:decompexample2} demonstrates the $dl,lh$
-decomposition for the hypothesis tree in figure~\ref{fig:decompexample}.
+Figure~\ref{fig:decompexample2} demonstrates the $dl$ and $lh$
+decompositions for the hypothesis tree in figure~\ref{fig:decompexample}.
 
@@
 \begin{displaymath}
-  EDPM = F[1g,2g,dl,lh], n=50, \gamma=0.25
+  \mbox{EDPM} = F[1g,2g,dl,lh], n=50, \gamma=0.25
 \end{displaymath}
@@
 features with a set of human judgments.  Within the TERp framework the
-features are the count of the various edit types normalized for the
+features are the counts of the various edit types normalized for the
 length of the reference so that the dot product of the features with a
-vector of edit costs will give the TERp score for a given segment.  In
+vector of edit costs will give the TERp score for a given sentence.  In
 these experiments, we follow~\inlinecite{snover09terp} by first
@@
 list by including the syntactic overlap features, and we use the TERp
-optimizer tools to tune the relative weight of each feature.
+optimizer to tune the relative weight of each feature.
 
@@
 \item[P]: features from the automatically-derived TERp paraphrase
-  table. These are four features which are a function of of the edit
+  table. These are four features which are a function of the edit
   distance between the two phrases in each paraphrase and the
   probability of the paraphrase.
-\item[B]: two ``length penalties'' one is assigned the count of
+\item[B]: two ``length penalties'': brevity and prolixity. One is assigned the count of
   hypothesis words beyond the count of reference words, and one
   assigned the count of reference words beyond the count of hypothesis
-  wordsbrevity/prolixity (2 features: one for longer-than-ref, one for
-  shorter-than-ref).
+  words.
   % MO: out of curiosity, why do you need 2 features?

### 2009-07-29 mtjournal.tex
fixed spelling
expanded some acronyms, corrected spelling issues

@@
 \inlinecite{owczarzak07labelleddepseval} proposed \textbf{d} and
-\textbf{d\_var} measures that compare LFG-derived relational tuples
-from reference and hypothesis translations and reported substantial
+\textbf{d\_var} measures that compare relational tuples derived from a
+lexical functional grammar (LFG)
+over reference and hypothesis translations and reported substantial
 improvement in correlation with human judgment relative to BLEU and
@@
 \inlinecite{owczarzak07labelleddepseval} using a widely-used and
-publically available PCFG parser and deterministic head-finding rules.
+publicly-available probabilistic context-free grammar (PCFG) parser and deterministic head-finding rules.
 
@@
 good examples are the $dl$ decomposition, which generates a bag of
-dependent word with outbound links, and the $lh$ decomposition, which
+dependent words with outbound links, and the $lh$ decomposition, which
 generates a bag of inbound link labels, with the head word for each

### 2009-07-29 mtjournal.tex
added justification subclause to sentence about modification to head-finding rules

@@
 training parameters.  Head-finding uses the Charniak parser's rules,
-with three modifications: prepositional and complementizer phrases
-choose nominal and verbal heads respectively (rather than functional
-heads) and auxiliary verbs are modifiers of main verbs (rather than
-the converse).
+with three modifications to make the semantic (rather than syntactic)
+relations more dominant in the dependency tree: prepositional and
+complementizer phrases choose nominal and verbal heads respectively
+(rather than functional heads) and auxiliary verbs are modifiers of
+main verbs (rather than the converse).
 

### 2009-07-29 mtjournal.tex
added fluency/adequacy mention to tables 1-3

@@
 \begin{table}
-  \caption{Comparing $dl,lh$ to $dlh$ and baselines.}
+  \caption{Comparing $dl,lh$ to $dlh$ and baseline correlations with
+    human fluency and adequacy judgments.}
   \label{tab:facorr:subgraphs}
@@
   \end{tabular*}
-  \caption{Comparing combination methods $F[\cdot]$ and
+  \caption{Correlation with human fluency/adequacy judgments. Comparing combination methods $F[\cdot]$ and
     $\mu_{PR}[\cdot]$, and comparing the $1g,2g,dl,lh$ decomposition
@@
   \end{tabular*}
-  \caption{Comparing $n=1$ to $n=50$ for several variants of the DPM measure.}
+  \caption{Correlation with human fluency/adequacy judgments. Comparing $n=1$ to $n=50$ for several variants of the DPM measure.}
   \label{tab:facorr:multiparse}

### 2009-07-29 mtjournal.tex
more info on MTC, and correlations with it

@@
   Pair Match (EDPM), is shown to be superior to BLEU and TER in terms
-  of correlation to human judgements and as a per-document and
+  of correlation to human judgments and as a per-document and
   per-sentence predictor of HTER, using mean subtraction to account
@@
 For these experiments, we use LDC Multiple Translation Chinese corpus
-parts 2~\cite{LDC03MTC2} and 4~\cite{LDC06MTC4}.  These corpora
+parts 2~\cite{LDC03MTC2} and 4~\cite{LDC06MTC4}, which are composed of
+translations of written Chinese news stories.  These corpora
 include multiple human judgments of fluency and adequacy for each
 sentence, with each judgment using a different human judge and a
-different reference translation.  For a rough\endnote{Our segment
+different reference translation.  Both fluency and adequacy
+assessments are assigned on a five-point scale. For a rough\endnote{Our segment
   count reported in section~\ref{sec:facorpus} differs from
@@
 experiments, we extend this tuple with automatic scores derived from
-$\langle$hypothesis, reference$\rangle$ and examine the
+$\langle$hypothesis, reference$\rangle$ and examine the per-segment
 correlations\endnote{The independence of the translation segments in
@@
 \begin{table}
-  \caption{Comparing $dl,lh$ to $dlh$ and baseline correlations with
+  \caption{Comparing $dl,lh$ to $dlh$ and baseline per-segment correlations with
     human fluency and adequacy judgments.}
@@
 In table~\ref{tab:facorr:subgraphs} we see that using the
-partial-label $dl,lh$ decomposition has a better correlation with the
+partial-label $dl,lh$ decomposition has a better per-segment correlation with the
 fluency/adequacy scores than TER or BLEU$_4$.  These results confirm,

### 2009-07-29 mtjournal.tex
reframed discussion of EDPM vs others in HTER correlation discussion

@@
 confidence, but the trend is the same across all of the
-subcorpora. EDPM's strengths, relative to the other measures, are
-particularly clear in the unstructured domains (\texttt{wb} and
-\texttt{bc}).
+subcorpora.
 
@@
 confidence of EDPM scores on nearly every breakdown.
+%
+In all cases in tables~\ref{tab:hterperdoc} and~\ref{tab:hterpersent},
+EDPM is the most strongly correlated with HTER. In structured data
+(\texttt{bn} and \texttt{nw}), these differences are not always
+significant, but in the unstructured domains (\texttt{wb} and
+\texttt{bc}), EDPM is always significantly better than at least one of
+the comparison baselines.
 

### 2009-07-29 mtjournal.tex
updated a bunch of refs to owczarzak mt journal paper

@@
 It can be tuned towards recall or precision, but is generally not
-tuned by users.  TERp \cite{snover09terp} is an extension of the
+tuned by users.  TER Plus (TERp) \cite{snover09terp} is an extension of the
 previously-mentioned TER that also incorporates synonym sets and stemming, along
@@
 decomposition of reference and hypothesis trees.
-\inlinecite{owczarzak07labelleddepseval} proposed \textbf{d} and
+\inlinecite{owczarzak07evaluatingmt} proposed \textbf{d} and
 \textbf{d\_var} measures that compare relational tuples derived from a
@@
 PCFG parsers with deterministic head-finding, while others
-\cite{owczarzak07labelleddepseval} extract the semantic dependency
+\cite{owczarzak07evaluatingmt} extract the semantic dependency
 relations from an LFG parser \cite{cahill04lfg}.
@@
 \inlinecite{roark06:sparseval} and
-\inlinecite{owczarzak07labelleddepseval} using a widely-used and
+\inlinecite{owczarzak07evaluatingmt} using a widely-used and
 publicly-available probabilistic context-free grammar (PCFG) parser and deterministic head-finding rules.
@@
 We may evaluate automatic MT measures in a variety of ways. Some
-\cite{banerjee05meteor,liu05syntaxformteval,owczarzak07labelleddepseval}
+\cite{banerjee05meteor,liu05syntaxformteval,owczarzak07evaluatingmt}
 have evaluated their success by comparing the measure to human
@@
 \inlinecite{roark06:sparseval} and
-\inlinecite{owczarzak07labelleddepseval}.  Rather than comparing
+\inlinecite{owczarzak07labelleddepseval}\endnote{
+  \inlinecite{owczarzak07evaluatingmt} extends their previous line of
+  research \cite{owczarzak07labelleddepseval} by variably-weighting
+  dependencies and by including synonym matching, two directions not
+  pursued here.
+}.  Rather than comparing
 string sequences, as BLEU does with its $n$-gram precision, this
@@
 
+% Reviewer 2 suggests reducing/removing the use of \mu here to a basic
+% explanation.
+
 When multiple decomposition types are used together, we may combine
@@
 Dependency-based SParseval and the \textbf{d} approach from
-\inlinecite{owczarzak07labelleddepseval} may each be understood as
+\inlinecite{owczarzak07evaluatingmt} may each be understood as
 $F[dlh]$, while the latter's \textbf{d\_var} method may be understood
@@
 for each word, pointing to its head-word). Prior work
-\cite{owczarzak07labelleddepseval} on related
+\cite{owczarzak07evaluatingmt} on related
 measures has used an LFG parser \cite{cahill04lfg} or
@@
   work is 50.  The uniform distribution ($\gamma = 0$) is intended to
-  be equivalent to the \cite{owczarzak07labelleddepseval}
-  \textbf{d\_50} and \textbf{d\_50\_var} measures.\endnote{
-    \inlinecite{owczarzak07labelleddepseval} report no use of parse
-    confidence weights, thus \textbf{d\_50} and \textbf{d\_50\_var}
-    may be using a sum of counts over the 50-best list rather than
-    expected-counts over a uniform distribution, as we suggest in
-    section~\ref{sec:metricdimensions}. These two approaches are
-    equivalent --- so long as the $n$-best list is always the same
-    length for hypothesis and reference.  In our implementation, the
-    $n$-best list does not always reach 50 candidate parses on short
-    sentences, so the expectation matches our intent better than a sum
-    of counts over the $n$-best.}
+  be equivalent to the \cite{owczarzak07evaluatingmt}
+  \textbf{d\_50} and \textbf{d\_50\_pm} measures.\endnote{
+    \inlinecite{owczarzak07evaluatingmt} report no use of parse
+    confidence weights, and \textbf{d\_50} and \textbf{d\_50\_pm}
+    return the best-matching pair of trees rather than an aggregate
+    over the tree sets.}
 \item[Score combination.] We consider three possible methods for
@@
   count reported in section~\ref{sec:facorpus} differs from
-  \inlinecite{owczarzak07labelleddepseval}, who report 16,800 segments
-  over the same corpus. We find baseline correlations (BLEU$_4$ and
-  TER) lower than those reported there as well, so the results
+  \inlinecite{owczarzak07evaluatingmt}, who report 16,807 segments
+  over the same corpus. We find baseline (BLEU and TER) per-segment correlations slightly
+  different (BLEU$_4$ is higher here, while TER here is lower) than
+  reported there as well, so the results
   presented here are not directly comparable with that paper, though
@@
   same corpus.
-} comparison with \inlinecite{owczarzak07labelleddepseval}, we treat
+} comparison with \inlinecite{owczarzak07evaluatingmt}, we treat
 each judgment as a separate segment.
@@
 fluency/adequacy scores than TER or BLEU$_4$.  These results confirm,
-with a PCFG, what \inlinecite{owczarzak07labelleddepseval} found with
+with a PCFG, what \inlinecite{owczarzak07evaluatingmt} found with
 an LFG parser: that partial-dependency matches are better correlated
@@
 probabilities to the parse forest, to compare as closely as possible
-to \inlinecite{owczarzak07labelleddepseval}, which includes a
+to \inlinecite{owczarzak07evaluatingmt}, which includes a
 \textbf{d\_var\_50} measure with 50-best parses, with ranks but no
@@
 place of the LFG parser used in
-\inlinecite{owczarzak07labelleddepseval}. PCFG parsers, however, are
+\inlinecite{owczarzak07evaluatingmt}. PCFG parsers, however, are
 fairly easily adaptable to another domain. One research

### 2009-07-29 mtjournal.tex
small clarifications to address some reviewer #2 concerns

@@
     recall scores.
-  \item Tuned, like the previous combination, calculates precision and
-    recall separately for each decomposition, but performs a linear
+  \item Tuned weight combination, like the previous $\mu_{PR}$
+    combination, calculates precision and
+    recall separately for each decomposition but performs a linear
     weighted combination of the resulting precision and recall scores.
@@
 harmonic mean of precision and recall subscores in $\mu_{PR}[\cdot]$.
+In fact, $F[\cdot]$ measures always outperform the BLEU baseline (from
+table~\ref{tab:facorr:subgraphs}), and $\mu_{PR}[\cdot]$ measures are
+never better than BLEU.
 We also see that we can combine the benefits of string-local $n$-grams

### 2009-07-29 mtjournal.tex
addressed "heavily-knowledge-engineered" by saying surface vs deep representation

@@
 considered as an approximation of the rich semantics generated by LFG
-parsers \cite{cahill04lfg} or another heavily knowledge-engineered
-parser, but with much less knowledge-engineering required.
+parsers \cite{cahill04lfg} or another parser with a rich semantic
+representation, but using a much shallower representation.
 

### 2009-07-30 mtjournal.tex
expanded a definition per R2 request

@@
 on $\bar{\mu}m$ and $\bar{\mu}\textrm{HTER}$ (rather than measure
-$m$ and HTER), where $\bar{\mu}$ represents mean-removal:
+$m$ and HTER), where $\bar{\mu}$ represents mean-removal and $t_i$
+represents the $i$th translation of segment $t$:
 \begin{equation}

### 2009-08-02 reviewers/review-response.tex
basics of review-response in place

% -*- LaTeX -*-
\documentclass[letterpaper,12pt]{article}

\newenvironment{response}
{\begin{quotation} \bf}
  {\end{quotation}}

\begin{document}

\section{Comments for the Author: Reviewer \#1}
\label{sec:reviewer1}

This submission on Dependency Pair Match MT metrics is technically and
structurally sound, thorough, and well-written.  Some specific
comments/questions below, in order of appearance in the paper.
\begin{response}
  Thank you for this feedback. Our responses are below.
\end{response}
As per the submission guidelines, submissions must be limited to 12
pages of content, with an additional page for references.  The current
version has 16.5 pages plus one page of references, so significant
shortening will be necessary.  I find the overall balance of sections
adequate, so I would recommend shortening across the board rather than
in one particular section.
\begin{response}
  We have done revisions to compress and simplify the writing
  throughout the paper to address the length concerns.
\end{response}
While likely known by most of the community, I would recommend
expanding all acronyms the first time (LFG, PCFG, ...).

In section 2, ``bag of dependent word'' should be ``bag of dependent
word*s*''.
\begin{response}
  We believe that we have emended these.
\end{response}
Can you expand on the reasoning behind the modifications to the
Charniak parser rules described in section 3.1?
\begin{response}
  We hope that our examples (and some additional explanatory sentences
  in section 3) have addressed this concern.
\end{response}
In section 4.1, while the corpus is referenced and the information can
be looked up, it would still be helpful to include directly in the
paper a brief description of the adequacy and fluency measures.  In
particular, what kind (granularity) of scale was used for the
judgments needs to be known in order to be able to assess the reported
correlations.  A brief description of the data genre(s) would be
useful to have at hand as well.  Also, I assume you correlate at the
*segment* level, but this is not explicitly stated.
\begin{response}
  Section 4 has now included several additional sentences to clarify
  these points.
\end{response}
The correlations reported in section 4 (tables I-III) are overall
quite low (also compared to EDPM's r correlations for Chinese source
data in NIST's MetricsMATR challenge).  Could you comment on this?
\begin{response} [TO DO]
\end{response}
For ease of lookup, please include in the table headings for tables
I-III that the correlations are with human judgments of adequacy and
fluency

In section 5.1, ``EDPM's strengths, relative to the other measures, are
particularly clear in the unstructured domains (wb and bc).''  This
statement does not seem to hold for EDPM vs. BLEU for the wb genre in
table Va; the correlation reported is the same for the two metrics?

\section{Comments for the Author: Reviewer \#2}
\label{sec:reviewer2}

This article presents a method for evaluating MT system hypotheses
that is based on complete and partial dependency triple matches
between hypothesis and reference translation.

The paper is well motivated, clearly written, and presents interesting
results. On my opinion, it is suitable for publication with minor
revisions. My main points of criticism are:

- It is considerably too long. Papers are supposed to be 12 pages +
 references; this paper has more than 16 pages. I recommend
 (a) shortening the introduction; (b) shortening the discussion
 of the "model space" in Sections 2/3 somewhat; (c) removing the
 third experiment (Section 6), whose contribution is unclear to
 me, altogether. (see below for details)

- On the other hand, the paper does not motivate at all what the
 intuition behind the use of dependency triple match is. Why
 should we assume in the first place that dependency triples
 predict human judgments better than (e.g.) n-grams?

- Finally, the paper does not give a single example. In a paper
 about a novel evaluation metric, I would like to get a feeling
 for what it is that the new metric does better than existing
 metrics. (This can be addressed together with the previous point.)

Detail comments:

- The discussion on page 2 reads nicely, but is probably too
 extensive -- in particular for a paper aimed at a special
 issue where there will presumably be an overview article
 that discusses relevant previous work.

- p.2, publically -> publicly?

- p.3, "In section 2 etc." I'm not sure that a 12-page paper requires
 a table of contents.

- p.3, reference to Owczarzak et al.'s work. In particular since their
 work has also appeared in the MT journal (please update the
 reference, btw!), it would be good to expand on the similarities and
 differences of your work and theirs.

-p.4 I'm not so sure that I believe in the benefit of overweighting.
 In standard dependency graphs, for example, modifications are usually
 dependendents of the head and are thus "underweighted" by your scheme,
 but modifiers like negations etc. can considerably change the meaning.
 (You write "we argue" but I don't see an argument, just a claim --
  details of counts for an example might be helpful.)

-p.5 "When multiple decompositions etc." As far as I can see, you
 get consistently better scores for F[] than for $\mu_{PR}[]$ throughout
 your experiments.  I'd suggest reducing the discussion of $\mu$ here
 and in Section 3.2 with a short explanation that $\mu$ has been found to
 work worse than F.

- p.6 "LFG...heavily knowledge-engineered parser". I think you're being
 overly harsh on the LFG-based work here. As the title of the paper you
 cite says, you're talking about an "automatically acquired PCFG-based
 LFG approximation" that's learned from a treebank, so you are really
 very much in the same boat.

- p.7 "Score combination" - see comment above
     "Tuned, like the ..." - ungrammatical sentence?

- Evaluation (p. 8) In principle I like the idea of teasing apart
 sentence difficulty and translation quality. However, I have two
 reservations: (1), you are losing comparability of scores across
 sentences; and (2), you are committing to a scenario where you
 always have access to a sufficient number of translations in order
 to estimate the mean score. In other words, in the limit of one
 system, you do not predict anything.  These issues should be
 mentioned and discussed.  (Note about presentation: you don't
 introduce the variable t in Equation 3 -- I guess this is the
 segment?)

 More generally, I would like to see a motivation for the evaluation
 you perform. You seem to largely follow Owczarzak et al's
 evaluation, e.g. in the choice of data (Section 4.1), but you depart
 from it in some crucial respects (e.g. the definition of the
 dependent variable by mean subtraction). Can you provide a clearer
 description of why each experiment is set up in the way it is?

- p.10 Am I right in reading Table II to say that F always outperforms
 BLEU, but $\mu$ never does? Maybe it would be good to re-mention the
 BLEU results in Table II.

- p.12 Can you elaborate on why you want to weight per-sentence
 correlations by sentence length? I don't feel that this
 appropriate. After all, I'd think that you are not \emph{more} interested
 in longer sentences, but you want each sentence to have the same
 contribution to the analysis. I assume the problem is that there is
 considerably more variance for shorter sentence, which is alleviated
 by length-based weighting, but to me this reads like a cover-up. If
 you have good arguments for wanting long sentences to count more in
 the analysis, I'd be interested to hear them.

- Section 6. To my mind, this is the least convincing section of the
 paper. The experiments in Section 4 have already shown that a simple
 F-score-based combination of the different matches does better than
 considering individual components. While I agree that this might be
 a weighting problem, this experiment shows basically a null result.
 From what I can gather from Figure 4, TERp on its own does
 (significantly?) better than EDPM, and their combination doesn't
 really improve a lot. I acknowledge the positive result that there
 is no \emph{decrease} in performance either, but then you're using quite
 a lot of data (50\% = 2500 sentences) for tuning so that is to be
 expected. I would very much like to see how robust the tuning
 process is to smaller development sets, but that would require even
 more space.

 In sum, I'd suggest replacing Section 6 with some actual examples
 and more discussion of the methodical points I have suggested.
\end{document}

### 2009-08-02 reviewers/review-response.tex
revising reviewer comments into LaTeX-ish for a response

@@
 data in NIST's MetricsMATR challenge).  Could you comment on this?
-\begin{response} [TO DO]
+\begin{response}
+ [TO DO]
 \end{response}
@@
 
-- It is considerably too long. Papers are supposed to be 12 pages +
- references; this paper has more than 16 pages. I recommend
- (a) shortening the introduction; (b) shortening the discussion
- of the "model space" in Sections 2/3 somewhat; (c) removing the
- third experiment (Section 6), whose contribution is unclear to
- me, altogether. (see below for details)
+\begin{itemize}
+\item It is considerably too long. Papers are supposed to be 12 pages
+  + references; this paper has more than 16 pages. I recommend (a)
+  shortening the introduction; (b) shortening the discussion of the
+  ``model space'' in Sections 2/3 somewhat; (c) removing the third
+  experiment (Section 6), whose contribution is unclear to me,
+  altogether. (see below for details)
 
-- On the other hand, the paper does not motivate at all what the
- intuition behind the use of dependency triple match is. Why
- should we assume in the first place that dependency triples
- predict human judgments better than (e.g.) n-grams?
+\item On the other hand, the paper does not motivate at all what the
+  intuition behind the use of dependency triple match is. Why should
+  we assume in the first place that dependency triples predict human
+  judgments better than (e.g.) n-grams?
 
-- Finally, the paper does not give a single example. In a paper
- about a novel evaluation metric, I would like to get a feeling
- for what it is that the new metric does better than existing
+\item Finally, the paper does not give a single example. In a paper
+  about a novel evaluation metric, I would like to get a feeling for
+  what it is that the new metric does better than existing
   metrics. (This can be addressed together with the previous point.)
+\end{itemize}
 
@@
 
-- The discussion on page 2 reads nicely, but is probably too
- extensive -- in particular for a paper aimed at a special
- issue where there will presumably be an overview article
- that discusses relevant previous work.
+\begin{itemize}
+\item The discussion on page 2 reads nicely, but is probably too
+  extensive -- in particular for a paper aimed at a special issue
+  where there will presumably be an overview article that discusses
+  relevant previous work.
 
-- p.2, publically -> publicly?
+\item  p.2, publically $\to$ publicly?
 
-- p.3, "In section 2 etc." I'm not sure that a 12-page paper requires
- a table of contents.
+\item  p.3, ``In section 2 etc.'' I'm not sure that a 12-page paper
+  requires a table of contents.
 
-- p.3, reference to Owczarzak et al.'s work. In particular since their
- work has also appeared in the MT journal (please update the
+\item  p.3, reference to Owczarzak et al.'s work. In particular since
+  their work has also appeared in the MT journal (please update the
   reference, btw!), it would be good to expand on the similarities and
@@
 
--p.4 I'm not so sure that I believe in the benefit of overweighting.
- In standard dependency graphs, for example, modifications are usually
- dependendents of the head and are thus "underweighted" by your scheme,
- but modifiers like negations etc. can considerably change the meaning.
- (You write "we argue" but I don't see an argument, just a claim --
-  details of counts for an example might be helpful.)
+\item p.4 I'm not so sure that I believe in the benefit of overweighting.
+  In standard dependency graphs, for example, modifications are
+  usually dependendents of the head and are thus ``underweighted'' by
+  your scheme, but modifiers like negations etc. can considerably
+  change the meaning.  (You write ``we argue'' but I don't see an
+  argument, just a claim -- details of counts for an example might be
+  helpful.)
 
--p.5 "When multiple decompositions etc." As far as I can see, you
- get consistently better scores for F[] than for $\mu_{PR}[]$ throughout
- your experiments.  I'd suggest reducing the discussion of $\mu$ here
- and in Section 3.2 with a short explanation that $\mu$ has been found to
- work worse than F.
+\item p.5 ``When multiple decompositions etc.'' As far as I can see, you
+  get consistently better scores for F[] than for $\mu_{PR}[]$
+  throughout your experiments.  I'd suggest reducing the discussion of
+  $\mu$ here and in Section 3.2 with a short explanation that $\mu$
+  has been found to work worse than F.
 
-- p.6 "LFG...heavily knowledge-engineered parser". I think you're being
- overly harsh on the LFG-based work here. As the title of the paper you
- cite says, you're talking about an "automatically acquired PCFG-based
- LFG approximation" that's learned from a treebank, so you are really
- very much in the same boat.
+\item  p.6 ``LFG...heavily knowledge-engineered parser''. I think you're
+  being overly harsh on the LFG-based work here. As the title of the
+  paper you cite says, you're talking about an ``automatically acquired
+  PCFG-based LFG approximation'' that's learned from a treebank, so you
+  are really very much in the same boat.
 
-- p.7 "Score combination" - see comment above
-     "Tuned, like the ..." - ungrammatical sentence?
+\item p.7 ``Score combination'' --- see comment above
 
-- Evaluation (p. 8) In principle I like the idea of teasing apart
+\item ``Tuned, like the ...''  --- ungrammatical sentence?
+
+\item  Evaluation (p. 8) In principle I like the idea of teasing apart
   sentence difficulty and translation quality. However, I have two
@@
 
-- p.10 Am I right in reading Table II to say that F always outperforms
- BLEU, but $\mu$ never does? Maybe it would be good to re-mention the
- BLEU results in Table II.
+\item  p.10 Am I right in reading Table II to say that F always
+  outperforms BLEU, but $\mu$ never does? Maybe it would be good to
+  re-mention the BLEU results in Table II.
 
-- p.12 Can you elaborate on why you want to weight per-sentence
+\item  p.12 Can you elaborate on why you want to weight per-sentence
   correlations by sentence length? I don't feel that this
- appropriate. After all, I'd think that you are not \emph{more} interested
- in longer sentences, but you want each sentence to have the same
- contribution to the analysis. I assume the problem is that there is
- considerably more variance for shorter sentence, which is alleviated
- by length-based weighting, but to me this reads like a cover-up. If
- you have good arguments for wanting long sentences to count more in
- the analysis, I'd be interested to hear them.
+  appropriate. After all, I'd think that you are not \emph{more}
+  interested in longer sentences, but you want each sentence to have
+  the same contribution to the analysis. I assume the problem is that
+  there is considerably more variance for shorter sentence, which is
+  alleviated by length-based weighting, but to me this reads like a
+  cover-up. If you have good arguments for wanting long sentences to
+  count more in the analysis, I'd be interested to hear them.
 
-- Section 6. To my mind, this is the least convincing section of the
+\item  Section 6. To my mind, this is the least convincing section of the
   paper. The experiments in Section 4 have already shown that a simple
@@
   really improve a lot. I acknowledge the positive result that there
- is no \emph{decrease} in performance either, but then you're using quite
- a lot of data (50\% = 2500 sentences) for tuning so that is to be
- expected. I would very much like to see how robust the tuning
+  is no \emph{decrease} in performance either, but then you're using
+  quite a lot of data (50\% = 2500 sentences) for tuning so that is to
+  be expected. I would very much like to see how robust the tuning
   process is to smaller development sets, but that would require even
   more space.
-
+\end{itemize}
  In sum, I'd suggest replacing Section 6 with some actual examples

### 2009-08-04 reviewers/review-response.tex
added another response.
gentle the responses by using italics rather than bold
inserted more to-do items and addressed the easier concerns. I'll fill in more as the drafts come in.

@@
 \newenvironment{response}
-{\begin{quotation} \bf}
+{\begin{quotation} \it}
   {\end{quotation}}
@@
 fluency
+\begin{response}
+  This suggestion has been included in the revised table headings.
+\end{response}
 
@@
 table Va; the correlation reported is the same for the two metrics?
+\begin{response}
+  [TO DO]
+\end{response}
 
@@
   altogether. (see below for details)
+  \begin{response}
+    Our revisions incorporate many of these suggestions. Rather than
+    eliminate section 6, we have chosen to compress it, but we have
+    adopted the rest of the suggestions above.
+  \end{response}
 
@@
   judgments better than (e.g.) n-grams?
+  \begin{response}
+    [TO DO -- using motivating examples?]
+  \end{response}
 
@@
   metrics. (This can be addressed together with the previous point.)
+  \begin{response}
+    [TO DO -- again using motivating examples?]
+  \end{response}
 \end{itemize}
@@
   relevant previous work.
+  \begin{response}
+    Thank you for pointing this out. We have squeezed this discussion
+    considerably.
+  \end{response}
 
 \item  p.2, publically $\to$ publicly?
+  \begin{response}
+    This spelling error has been corrected.
+  \end{response}
 
@@
   requires a table of contents.
+  \begin{response}
+    This section has also been compressed.
+  \end{response}
 
@@
   differences of your work and theirs.
+  \begin{response}
+    [TO DO].  The reference has been updated.
+  \end{response}
 
@@
   helpful.)
+  \begin{response}
+    [TO DO: motivating examples?]
+  \end{response}
 
@@
   has been found to work worse than F.
+  \begin{response}
+    [TO DO] reduce the discussion of $\mu_{PR}$ further?
+  \end{response}
 
@@
   are really very much in the same boat.
+  \begin{response}
+    Both are indeed built off treebanks, so the knowledge-engineering
+    is similar. The contrast is in fact between a more ``surface''
+    syntactic representation (the PCFG model) or a ``deeper'', more
+    semantic one, and the discussion has been changed to reflect this.
+  \end{response}
 
 \item p.7 ``Score combination'' --- see comment above 
+  \begin{response}
+    [TO DO]
+  \end{response}
 
 \item ``Tuned, like the ...''  --- ungrammatical sentence?
+  \begin{response}
+    This bullet's wording has been clarified.
+  \end{response}
 
@@
   system, you do not predict anything.  These issues should be
-  mentioned and discussed.  (Note about presentation: you don't
-  introduce the variable t in Equation 3 -- I guess this is the
+  mentioned and discussed.  
+  \begin{response}
+    [TO DO]
+  \end{response}
+  (Note about presentation: you don't
+  introduce the variable $t$ in Equation 3 -- I guess this is the
   segment?)
+  \begin{response}
+    Thank you! this oversight is now repaired and $t$ is explicit.
+  \end{response}
 
@@
   description of why each experiment is set up in the way it is?
+  \begin{response}
+    Section 4 is intended to match the Owczarzak \textit{et al.} data
+    selection, but sections 5 and 6 do not; in sections 5 and 6 the
+    data available \emph{does} allow multiple translations of the same
+    segment and thus we can use mean-subtraction. Mean subtraction is
+    only used in sections 5 and 6, and the discussion of
+    mean-subtraction has been moved to the beginning of section 5 for
+    clarity.
+  \end{response}
 
@@
   re-mention the BLEU results in Table II.
+  \begin{response}
+    [TO DO]
+  \end{response}
 
@@
   count more in the analysis, I'd be interested to hear them.
+  \begin{response}
+    [TO DO]
+  \end{response}
 
@@
 \end{itemize}
- In sum, I'd suggest replacing Section 6 with some actual examples
- and more discussion of the methodical points I have suggested.
+In sum, I'd suggest replacing Section 6 with some actual examples and
+more discussion of the methodical points I have suggested.
+\begin{response}
+  [TO DO]: We have made a number of changes to section 6.  We have
+  revisited the two-fold validation and provided a new statistical
+  analysis that simplifies both the presentation (figure 4) and the
+  discussion (elsewhere in section 6).
+\end{response}
 \end{document}

### 2009-08-05 reviewers/review-response.tex
added an extra response at the top of reviewer-2's text.
many small response revisions encoding Mari's handwritten notes

@@
 expanding all acronyms the first time (LFG, PCFG, ...).
+\begin{response}
+  Done.
+\end{response}
 
@@
 \begin{response}
-  We believe that we have emended these.
+  Done.
 \end{response}
@@
 \begin{response}
-  We hope that our examples (and some additional explanatory sentences
-  in section 3) have addressed this concern.
+  We added discussion in section 3 and refer back to the example in
+  section two, which illustrates this point. [TO DO -- emend the discussion].
 \end{response}
@@
 \begin{response}
- [TO DO]
+ [TO DO -- add a footnote]
 \end{response}
@@
 \begin{response}
-  This suggestion has been included in the revised table headings.
+  Done.
 \end{response}
@@
 \begin{response}
-  [TO DO]
+  The sentence at the end of the section is revised to indicate that
+  ``in the unstructured domains (\texttt{wb} and \texttt{bc}), EDPM is
+  always significantly better than at least one of the comparison
+  baselines.''
 \end{response}
@@
 results. On my opinion, it is suitable for publication with minor
-revisions. My main points of criticism are:
+revisions. 
+\begin{response}
+  Thank you for your detailed critique and many useful
+  suggestions. Our responses are below:
+\end{response}
+My main points of criticism are:
 

### 2009-08-05 reviewers/review-response.tex
more notes transcribed from mari's notes.

@@
   \begin{response}
-    [TO DO -- using motivating examples?]
+    We include an example in section 2 to address this point.  [TO DO
+    -- using motivating examples?]
   \end{response}
@@
   \begin{response}
-    [TO DO -- again using motivating examples?]
+    Done.
   \end{response}
@@
   \begin{response}
-    Thank you for pointing this out. We have squeezed this discussion
-    considerably.
+    We have compressed this discussion considerably.
   \end{response}
@@
   \begin{response}
-    This spelling error has been corrected.
+    Done.
   \end{response}
@@
   \begin{response}
-    [TO DO].  The reference has been updated.
+    [TO DO -- expand to rebuttal on expanding similarities and
+    differences -- note that sec. 4 is essentially this].  The
+    reference has been updated.
   \end{response}
@@
   \begin{response}
-    [TO DO] reduce the discussion of $\mu_{PR}$ further?
+    [TO DO] reduce the discussion of $\mu_{PR}$ further? [TO DO:
+    revisit after compression]
   \end{response}

### 2009-08-05 reviewers/review-response.tex
more responses coded in from Mari's hand notes

@@
   \begin{response}
-    [TO DO]
+    [TO DO -- revisit after compression]
   \end{response}
@@
   \begin{response}
-    This bullet's wording has been clarified.
+    Clarified.
   \end{response}
@@
   \begin{response}
-    [TO DO]
+    [TO DO] We agree that length normalization is not possible in all
+    scenarios and in fact did not use it in section four.  Hence, we
+    moved the length normalization description to section 5, since it
+    is only used whene we have multiple translations.  
   \end{response}
@@
   \begin{response}
-    Thank you! this oversight is now repaired and $t$ is explicit.
+    fixed.
   \end{response}
@@
   \begin{response}
+    [TO DO: consolidate with other responses?]
+
     Section 4 is intended to match the Owczarzak \textit{et al.} data
@@
   \begin{response}
-    [TO DO]
+    [TO DO revisit after compression]
   \end{response}
@@
   \begin{response}
-    [TO DO]
+    [TO DO --- include rebuttal.]
   \end{response}
@@
 \begin{response}
-  [TO DO]: We have made a number of changes to section 6.  We have
+  [TO DO -- list out other changes to section 6]: We have made a number of changes to section 6.  We have
   revisited the two-fold validation and provided a new statistical
   analysis that simplifies both the presentation (figure 4) and the
-  discussion (elsewhere in section 6).
+  discussion (elsewhere in section 6). In particular, the changes
+  include:
+  \begin{enumerate}
+  \item dropping the discussion of $B$ brevity penalties
+  \item collapsing the $T$ (edit and synonymy) and $P$ (paraphrasing)
+    categories together (now presented together as $T$)
+  \item folding together and normalizing the values produced for the
+    cross-tuned combination measures to provide a single value and
+    confidence interval rather than presenting both splits.
+  \end{enumerate}
 \end{response}

### 2009-08-10 mtjournal.tex
revised figure to use only T,N,E features
Mari comments
Mari's revisions to title spacing and abstract

@@
+%%MO: changes other than compression
+%% - changed notation for mean normalization in sect 5 to avoid confusion
+%%   of two notions of mu
+%% - moved tables out of the middle of paragraphs since that's visually
+%%   easier for me to deal with in editing
+
 \documentclass{kluwer}    % Specifies the document style.
@@
 \begin{opening}         
-\title{Expected Dependency Pair Match:\\
+\title{Expected Dependency Pair Match:
 Predicting translation quality with expected syntactic structure} 
@@
 \begin{abstract}
-  Recent efforts aimed at improving over standard machine translation
-  evaluation methods (BLEU, TER) have investigated mechanisms for
-  accounting for allowable wording differences either in terms of
-  syntactic structure or synonyms/paraphrases. This paper explores an
-  approach for combining scores from partial syntactic dependency
+  Recent efforts to develop new machine translation
+  evaluation methods have tried to 
+  account for allowable wording differences either in terms of
+  syntactic structure or synonyms/paraphrases. This paper primarily considers syntactic structure, combining scores from partial syntactic dependency
   matches with standard local n-gram matches using a statistical
-  parser, and taking advantage of parse probabilities in deriving
-  expected scores based on the N-best parses for the hypothesized
-  sentence translation.  The new scoring metric, Expected Dependency
+  parser, and taking advantage of N-best parse probabilities in deriving
+  scores for the hypothesized
+  translation.  The new scoring metric, Expected Dependency
   Pair Match (EDPM), is shown to be superior to BLEU and TER in terms
   of correlation to human judgments and as a per-document and
-  per-sentence predictor of HTER, using mean subtraction to account
-  for document difficulty. Further, we explore the potential benefit
-  of combining the n-gram and syntactic features of EDPM with the
-  alternative wording features of TERp, with experiments showing that
-  there is a benefit to accounting for syntactic structure on top of
-  the semantic equivalency features.
+  per-sentence predictor of HTER. Further, we combine the syntactic features of EDPM with the
+  alternative wording features of TERp, showing a benefit to accounting for syntactic structure on top of
+  semantic equivalency features.
 \end{abstract}
@@
   \end{center}
-  \caption{Pearson's $r$ values for the two halves of the training
-    set, with 95\% confidence intervals. }
+  \caption{Pearson's $r$ values for various feature tunings, with 95\%
+    confidence intervals. BLEU$_4$ (add-one smoothed), TER, and EDPM
+    scores are provided on the left for comparison.}
   \label{fig:tuneresults}

### 2009-08-10 mtjournal.tex
Mari's changes to introduction

@@
 
-Machine translation (MT) evaluation is a challenge for research
-because the space of good translations is large, and two equally good
-translations may appear to be quite different at first glance. 
-%
-The challenges of choosing among translations are compounded when this
-evaluation is done automatically.
-%
-Human evaluation, however, is both time-consuming and difficult, so
-research has turned increasingly towards automatic measures of
-translation quality, usually by comparing the system (hypothesis) translation to
-one or more reference (human) translations.
-%
-Automatic measures of this kind not
-only provide a well-defined evaluation standard but are also required
-for training on error criteria, e.g.\ with minimum error rate training
-\cite{och03mert}.
-
+A challenge in automatic machine translation (MT) evaluation is accounting
+for allowable variability: two equally good
+translations may be quite different in surface form. 
 Currently, the most popular evaluation measures include a measure
@@
 the edit-distance measure Translation Edit Rate (TER)
-\cite{snover06ter}.  Recent research has found that these measures may
+\cite{snover06ter}.  These measures can only account for variability when 
+given multiple translations, and studies show that they may
 not accurately track translation quality both empirically
@@
 
-These challenges have motivated a search for better measures that
-incorporate additional language knowledge sources.  METEOR
-\cite{banerjee05meteor}, for example, uses synonym tables and
-morphological stemming to do progressively more-forgiving matching.
-It can be tuned towards recall or precision, but is generally not
-tuned by users.  TER Plus (TERp) \cite{snover09terp} is an extension of the
+Alternative measures that 
+incorporate synonym knowledge sources include: METEOR
+\cite{banerjee05meteor}, which uses synonym tables and
+morphological stemming to do progressively more forgiving matching;
+%%MO: I am taking out the tuning stuff here since it isn't critical
+%It can be tuned towards recall or precision, but is generally not
+%tuned by users.  
+TER Plus (TERp) \cite{snover09terp}, which is an extension of the
 previously-mentioned TER that also incorporates synonym sets and stemming, along
-with automatically-derived paraphrase tables.  TERp is explicitly
-intended to be tuned to a development set by users.
+with automatically-derived paraphrase tables.  
+%TERp is explicitly
+%intended to be tuned to a development set by users.
 %
-Tuning has the advantage that the weight of different types of errors
-can be adjusted to match the needs of the task, though it makes it
-more difficult to compare results across tasks, particularly when
-there is little data for tuning.
-
-As an alternative to these synonym- and paraphrase-based approaches,
-other metrics model syntactically-local (rather than string-local)
-word-sequences. \inlinecite{liu05syntaxformteval} compared tree-local
+%Tuning has the advantage that the weight of different types of errors
+%can be adjusted to match the needs of the task, though it makes it
+%more difficult to compare results across tasks, particularly when
+%there is little data for tuning.
+Other metrics model syntactically-local (rather than string-local)
+word-sequences include: tree-local
 $n$-gram precision in various configurations of constituency and
-dependency trees.  The dependency-based SParseval measure
+dependency trees \cite{liu05syntaxformteval};\endnote{The dependency-based SParseval measure
 \cite{roark06:sparseval}, designed as a parse-quality metric for
 speech, is a similar approach, in that it is an F-measure over a
-decomposition of reference and hypothesis trees.
-\inlinecite{owczarzak07evaluatingmt} proposed \textbf{d} and
-\textbf{d\_var} measures that compare relational tuples derived from a
+decomposition of reference and hypothesis trees.} and the \textbf{d} and
+\textbf{d\_var} measures proposed by \inlinecite{owczarzak07evaluatingmt} that compare relational tuples derived from a
 lexical functional grammar (LFG)
-over reference and hypothesis translations and reported substantial
-improvement in correlation with human judgment relative to BLEU and
-TER.
-
+over reference and hypothesis translations.
+%
 These syntactically-oriented measures require a system for proposing
 dependency structure over the reference and hypothesis
-translations. Some \cite{liu05syntaxformteval,roark06:sparseval} use
-PCFG parsers with deterministic head-finding, while others
+translations. \inlinecite{liu05syntaxformteval} use a
+PCFG parser with deterministic head-finding, while 
 \cite{owczarzak07evaluatingmt} extract the semantic dependency
@@
 This work extends the dependency-scoring strategies of
-\inlinecite{roark06:sparseval} and
-\inlinecite{owczarzak07evaluatingmt} using a widely-used and
+\inlinecite{owczarzak07evaluatingmt}, which reported substantial
+improvement in correlation with human judgment relative to BLEU and
+TER, by using a 
 publicly-available probabilistic context-free grammar (PCFG) parser and deterministic head-finding rules.
+In addition, we consider more types of constituents and different score combinations, as well as combination with of synonym-type scores.
 
-We may evaluate automatic MT measures in a variety of ways. Some
+MT measures are evaluated in a variety of ways. Some
 \cite{banerjee05meteor,liu05syntaxformteval,owczarzak07evaluatingmt}
-have evaluated their success by comparing the measure to human
+comparing the measure to human
 judgments of fluency and adequacy.  In other work, e.g.\
-\inlinecite{snover06ter}, measures are evaluated by comparison to
+\cite{snover06ter}, measures are compared to
 human-targeted TER (HTER), a distance to a human-revised reference
@@
 
-In section~\ref{sec:approach}, we outline our approach to including
-syntactic information in MT evaluation, by describing a family of
-metrics. In section~\ref{sec:paradigm}, we describe the implementation
-details and make explicit which dimensions of this family of metrics
-we explore in this paper.  Section~\ref{sec:faexpts} examines the
+We describe our approach to including syntactic information in MT
+evaluation by outlining a family of metrics in
+section~\ref{sec:approach} and provide implementation details in
+section~\ref{sec:paradigm}.  Section~\ref{sec:faexpts} examines the
 correlation of several members of this family with human judgments of
-fluency and adequacy over a corpus of sentence translations and
-selects a favorable member of the family, Expected Dependency Pair
-Match (EDPM), based on this comparison. Section~\ref{sec:hter1}
-compares the correlation of EDPM with HTER over a corpus of document-
-and sentence-level comparisons, and demonstrates that its correlation
-is favorable to that of competitor measures BLEU and TER at both
-sentence and document levels, regardless of source-language or genre.
-Section~\ref{sec:hter2} uses the tuning components of TERp
-to explore the combination of these syntactic measures with TERp's
-word-based measures. Finally, section~\ref{sec:conclusion} concludes
-with some remarks on challenges and future work.
+fluency and adequacy, using the paradigm from
+\cite{owczarzak07evaluatingmt} to provide comparisons to that work and
+select a best case configuration, Expected Dependency Pair Match
+(EDPM). The EDPM measure is then compared to BLEU and TER in terms of
+correlation with HTER, exploring language/genre effects in
+section~\ref{sec:hter1} and combination with TERp's synonym/paraphrase
+features in section~\ref{sec:hter2}. Finally, findings and future work
+are summarized in section~\ref{sec:conclusion}.
 

### 2009-08-10 mtjournal.tex
made figures agnostic between eps and other pdflatex-able formats
Mari's major changes to Approach section

@@
 
-This work explores a family of dependency pair match (DPM) measures
-that are composed of precision and recall combinations over various
+The specific family of dependency pair match (DPM) measures explored here
+combines precision and recall scores of various
 decompositions of a syntactic dependency tree. These measures are
 extensions of the dependency-pair F measures found in
-\inlinecite{roark06:sparseval} and
 \inlinecite{owczarzak07labelleddepseval}\endnote{
-  \inlinecite{owczarzak07evaluatingmt} extends their previous line of
+  \inlinecite{owczarzak07evaluatingmt} extend their previous line of
   research \cite{owczarzak07labelleddepseval} by variably-weighting
+%%MO: you actually do variable weighting using parse probabilities, so
+%% would be good to mention what they use.  If you can say this in a few
+%% words, then you won't add a line
   dependencies and by including synonym matching, two directions not
@@
 
-A member of this family is defined by several parameters.  The first
-of these parameters is the nature of the decomposition of the
-dependency tree structure.  A \emph{decomposition list} is the list of
-ways in which the tree is reduced to a bag of tree-local
-tuples. Figure~\ref{fig:decompexample}
+We motivate the use of dependencies with actual translations:\\[4pt]
+{\sl
+Ref: Authorities have also closed southern Basra's airport and seaport.\\
+S1: The authorities also closed the airport and seaport in the
+southern port of Basra.\\
+S2: Authorities closed the airport and the port of.\\[4pt]
+}
+A human judged the system 1 result (S1) as equivalent to the reference, but the system 2 (S2) result as
+having errors.  BLEU gives both a similar score (0.199
+vs.\ 0.203). TER scores S2 as better (errors
+of 0.9 vs.  0.7, respectively), since a simple deletion requires fewer
+edits that rephrasing.  By representing matches of dependencies, we
+obtain a score for S1 from the new EDPM measure that is higher than
+that for S2 (0.414 vs.\ 0.356).  The two phrases ``southern Basra's
+airport and seaport'' and ``the airport and seaport in the southern
+port of Basra'' have more in similarities in terms of dependencies
+than word order.
+
+The particular relations that are extracted from the dependency tree
+are referred to here as \emph{decompositions}.
+Figure~\ref{fig:decompexample}
+illustrates the \emph{dependency-link-head} decomposition of a toy
+dependency tree into a list of $\langle d, l, h \rangle$ tuples.  Some
+members of the DPM family may apply more than one decomposition; other
+good examples are the $dl$ decomposition, which generates a bag of
+dependent words with outbound links, and the $lh$ decomposition, which
+generates a bag of inbound link labels, with the head word for each
+included. Figure~\ref{fig:decompexample2} shows the $dl$ and $lh$
+decompositions for the same hypothesis tree.
+
 \begin{figure}
@@
     \textbf{tree}
-    & \includegraphics[scale=0.5]{dpm-example-ref.eps} & 
-    \includegraphics[scale=0.5]{dpm-example-hyp.eps}\\
+    & \includegraphics[scale=0.5]{dpm-example-ref} & 
+    \includegraphics[scale=0.5]{dpm-example-hyp}\\
     \textbf{$dlh$ list} &
@@
   \end{tabular}\\
-  Precision$_{dlh}$ is $\frac{1}{3}$ and Recall$_{dlh}$ is
-  $\frac{1}{4}$: thus $F[dlh] = \frac{2}{7}$
+%%MO: I don't think we need the P/R/F at this point in the narrative
+%  Precision$_{dlh}$ is $\frac{1}{3}$ and Recall$_{dlh}$ is
+%  $\frac{1}{4}$: thus $F[dlh] = \frac{2}{7}$
   \caption{Example hypothesis and reference dependency trees and the
@@
 \end{figure}
-illustrates the \emph{dependency-link-head} decomposition of a toy
-dependency tree into a list of $\langle d, l, h \rangle$ tuples.  Some
-members of the DPM family may apply more than one decomposition; other
-good examples are the $dl$ decomposition, which generates a bag of
-dependent words with outbound links, and the $lh$ decomposition, which
-generates a bag of inbound link labels, with the head word for each
-included. 
+
 \begin{figure}
@@
   \begin{center}
-    \includegraphics[scale=0.6]{dpm-example-depextract.eps}
+    \includegraphics[scale=0.6]{dpm-example-depextract}
   \end{center}

### 2009-08-10 mtjournal.tex
lots of changes from mu-overline Score to \overline only

@@
   \end{tabular}\\
-%%MO: I don't think we need the P/R/F at this point in the narrative
-%  Precision$_{dlh}$ is $\frac{1}{3}$ and Recall$_{dlh}$ is
-%  $\frac{1}{4}$: thus $F[dlh] = \frac{2}{7}$
+  Precision$_{dlh}$ is $\frac{1}{3}$ and Recall$_{dlh}$ is
+  $\frac{1}{4}$: thus $F[dlh] = \frac{2}{7}$
   \caption{Example hypothesis and reference dependency trees and the
@@
 \end{figure}
-
+illustrates the \emph{dependency-link-head} decomposition of a toy
+dependency tree into a list of $\langle d, l, h \rangle$ tuples.  Some
+members of the DPM family may apply more than one decomposition; other
+good examples are the $dl$ decomposition, which generates a bag of
+dependent words with outbound links, and the $lh$ decomposition, which
+generates a bag of inbound link labels, with the head word for each
+included. 
 \begin{figure}
@@
       \hline
-      $r$ vs. $\bar{\mu}$HTER & \multicolumn{1}{c}{\texttt{bc}}
+      $r$ vs. $\overline{\textrm{HTER}}$ & \multicolumn{1}{c}{\texttt{bc}}
       & \multicolumn{1}{c}{\texttt{bn}} &
@@
       \hline
-      $\bar{\mu}$TER
+      $\overline{\textrm{TER}}$
       &  0.59 &  \textbf{0.35} &  \textbf{0.47}&  \textit{0.17}
       & \textbf{0.54} & \textbf{0.32} &  0.44\\
-      $\bar{\mu}$BLEU
+      $\overline{\textrm{BLEU}}$
       & -0.42 & \textbf{-0.32} & \textbf{-0.46}&  \textbf{-0.27}
       & -0.42  &  \textbf{-0.33} & -0.37 \\
-      $\bar{\mu}$EDPM
+      $\overline{\textrm{EDPM}}$
       & \textbf{-0.69} & \textbf{-0.39} & \textbf{-0.47} &
@@
     \end{tabular}
-    \caption{Per-document correlations of $\bar{\mu}$EDPM and others to
-      $\bar{\mu}$HTER, by genre and by source language. Bold numbers are
+    \caption{Per-document correlations of $\overline{\textrm{EDPM}}$ and others to
+      $\overline{\textrm{HTER}}$, by genre and by source language. Bold numbers are
       within 95\% significance of the best per column; italics indicate
@@
       \hline
-      $\bar{\mu}$TER
+      $\overline{\textrm{TER}}$
       &  \textbf{0.44} &  \textbf{0.29} &  \textbf{0.33}&  0.25
       & \textbf{0.44} & 0.25  &  \textbf{0.36}\\
-      $\bar{\mu}$BLEU
+      $\overline{\textrm{BLEU}}$
       & -0.31 & -0.24  & -0.29 &  -0.25
       & -0.31  &  -0.24  & -0.28 \\
-      $\bar{\mu}$EDPM
+      $\overline{\textrm{EDPM}}$
       & \textbf{-0.46} & \textbf{-0.31} & \textbf{-0.34} &
@@
     \caption{Per-sentence, length-weighted correlations of
-      $\bar{\mu}$EDPM and others to $\bar{\mu}$HTER, by genre and by
-      source language.}
+      $\overline{\textrm{EDPM}}$ and others to $\overline{\textrm{HTER}}$, by genre and by
+      source language. Bold numbers indicate significance as above.}
     \label{tab:hterpersent}

### 2009-08-10 mtjournal.tex
small wording and title changes from Mari

@@
 \end{eqnarray}
-Dependency-based SParseval and the \textbf{d} approach from
+where $\mu_h$ represents a harmonic mean.
+Dependency-based SParseval \inlinecite{roark06:sparseval} 
+and the \textbf{d} approach from
 \inlinecite{owczarzak07evaluatingmt} may each be understood as
@@
 
-\subsection{Parse tree implementation}
-In principle, the family of DPM measures may be implemented with
+\subsection{Parsing and dependency extraction}
+
+The family of DPM measures may be implemented with
 any parser that generates a dependency graph (a single labelled arc
@@
 \inlinecite{charniak-johnson:2005:ACL}) and context-free head-finding
-rules \cite{magerman95headfinding} to generate a 50-best list of
+rules \cite{magerman95headfinding} to generate an N-best list of
 dependency trees for each hypothesis and reference translation.  We
-use the parser's default (English-language) Wall Street Journal
+use the parser's default (English) Wall Street Journal
 training parameters.  Head-finding uses the Charniak parser's rules,
@@
 
-{\small This material is based upon work supported by the National
+{\small This material is based on work supported by the National
   Science Foundation under Grant No. 0741585 and the Defense Advanced
@@
   HR0011-06-C-0023.
-
 % Re: DARPA contract numbers
@@
 % -0023 is Mari & Jeremy's [Nightingale] grant
-
-  Any opinions, findings, and conclusions or recommendations expressed
-  in this material are those of the author(s) and do not necessarily
-  reflect the views of the National Science Foundation or the Defense
-  Advanced Research Projects Agency.  }
+%%MO: slight rewording to save lines
+  Any opinions, findings, conclusions or recommendations expressed
+  herein are those of the authors and do not necessarily
+  reflect the views of the funding agencies.  }
 

### 2009-08-10 mtjournal.tex
Mari's major compression and reorganization for compact.

@@
 \end{figure}
-Figure~\ref{fig:decompexample2} demonstrates the $dl$ and $lh$
-decompositions for the hypothesis tree in figure~\ref{fig:decompexample}.
 
@@
 words in the tree, are included multiple times in the decomposition
-(once for each inbound link).  We argue that this overcounting is a
-virtue; the syntactic structure indicates which words are more
-important to translate correctly.
+(once for each inbound link).  This ``overcounting'' leverages
+syntactic indications of which words are more
+important to translate correctly (e.g., ``Basra'' in the
+example).
 
-We may not completely trust the parser's best parse.  The parser
-itself, if we use a probabilistic parser, can provide an $n$-best
-parse list for the translation reference and translation
-hypothesis. We use the probability statistics of the list to compute
-expected counts for each decomposition. Though this approach yields
-partial counts, standard comparisons like precision and recall are
-still valid, and using the expectations can help cope with both error
-in the parser and ambiguity in the translations (reference and
+A probabilistic parser provides confidences associated with parses
+in an $n$-best list, which we can use to compute
+expected counts for each decomposition in both the reference and
+hypothesized translations. The expected counts lead to partial matches
+which are used to compute precision and recall as weighted counts.
+Using the expectations can help cope with both error
+in the best parse and ambiguity in the translations (reference and
 hypothesis).
@@
 
+%%MO: I think that it is important to keep because of some of the comparisons
+%% you make, so I will try to just compress
+
 When multiple decomposition types are used together, we may combine
-these subscores in a variety of ways.  We may compute precision and
-recall subscores for each decomposition separately, or, since the
-results of each decomposition are of different types entirely, we may
-compute them as members of one large bag for an even simpler F score.
-These two approaches are equivalent when only one decomposition type
-is included.  For simplicity in presentation, we use the following
-notation, where $dl$ and $lh$ represent the two kinds of
-decompositions described above and $\mu_h$ represents a harmonic mean:
+these subscores in a variety of ways. Here, we experiment with using
+two variations of a harmonic mean:
+computing precision and recall over all
+decompositions as a group (giving a single precision and recall number)
+vs.\ computing precision and recall separately for each decomposition.
+We distinguish between these using the notation:
 \begin{eqnarray}
@@
 as something close to $F[dl,lh]$.
+%
+Both the combination methods $F$ and
+$\mu_{PR}$ are ``naive'' in that they treat each component
+score as equivalent to the next.  When we introduce syntactic/paraphrasing
+features, we will move to a weighted combination.
 
-Both of the combination methods above ($F[\cdot]$ and
-$\mu_{PR}[\cdot]$) are ``naive'' in that they treat each component
-score as equivalent to the next.  One further direction to explore
-involves tuning (presumably on held-out data) the precision and recall
-weights for each decomposition type for a later linear combination.
+%%MO: actually the next section does more than this, and this sentence is
+%% essentially repeated later, so I deleted it here
+%The possible family of metrics outlined above is quite large. In the
+%next section, we make explicit the range of these parameters that we
+%explore in this article.
 
-The possible family of metrics outlined above is quite large. In the
-next section, we make explicit the range of these parameters that we
-explore in this article.
+%%MO: I changed the title, since I usually put corpus in ``paradigm''. You 
+%% are really just talking about implementation here
 
-\section{Experimental paradigm}
+\section{Experimental implementation}
 \label{sec:paradigm}
@@
 (rather than functional heads) and auxiliary verbs are modifiers of
-main verbs (rather than the converse).
+main verbs (rather than the converse). These changes capture the
+fact that main verbs are more important for adequacy in translation, as
+illustrated by the functional equivalence of ``have also closed'' vs.\ 
+``also closed'' in the example in section~\ref{sec:approach}.
 
-Having constructed the dependency tree, we label the arcs as $d
-\stackrel{A/B}{\to} h$, where the arc label $A/B$ between dependent
-$d$ and its head $h$ is composed of $A$ (the lowest constituent-label
-headed by $h$ and dominating $d$) and $B$ (the highest constituent
+Having constructed the dependency tree, we label the arc between dependent
+$d$ and its head $h$ as $A/B$ when $A$ is the lowest constituent-label
+headed by $h$ and dominating $d$ and $B$ is the highest constituent
 label headed by $d$).
+%
+For example, in figure~\ref{fig:depextract}, the S node is the lowest
+node headed by \emph{stumbled} that dominates \emph{cat}, and the NP
+node is the highest constituent label headed by \emph{cat}, so the arc
+linking \emph{cat} to \emph{stumbled} is labelled $S/NP$.
+%
+This strategy is very similar to one adopted in the reference
+implementation of labelled-dependency \textsc{SParseval} \cite{roark06:sparseval}, and may be
+considered as a shallow approximation of the rich semantics generated by LFG
+parsers \cite{cahill04lfg}.
+%
+The $A/B$ labels are not as descriptive as the LFG semantics, but they
+have a similar resolution, e.g.\ the $S/NP$ arc label
+usually represents a subject dependent of a sentential verb.
+
 \begin{figure}
@@
     [.\headlabel{VP}{stumbled} 
-      [.\headlabel{VBD}{stumbled} stumbled ]
+      [.\headlabel{VBD}{\it stumbled} stumbled ]
     ]
@@
   \caption{An example constituent tree (heads of each constituent are
-    listed small below the label) and the labelled dependency tree
-    derived from it using the strategy described in
-    section~\ref{sec:paradigm}.}
+    listed below the label) and the labelled dependency tree
+    derived from it.}
   \label{fig:depextract}
 \end{figure}
-For example, in figure~\ref{fig:depextract}, the S node is the lowest
-node headed by \emph{stumbled} that dominates \emph{cat}, and the NP
-node is the highest constituent label headed by \emph{cat}, so the arc
-between \emph{cat} and \emph{stumbled} is labelled \arclabel{S/NP}.
-%
-This strategy is very similar to one adopted in the reference
-implementation of labelled-dependency \textsc{SParseval}, and may be
-considered as an approximation of the rich semantics generated by LFG
-parsers \cite{cahill04lfg} or another parser with a rich semantic
-representation, but using a much shallower representation.
 
-The $A/B$ labels are not as descriptive as the LFG semantics, but they
-have a similar resolution, e.g.\ the $\stackrel{S/NP}{\to}$ arc label
-usually represents a subject dependent of a sentential verb.
+%%MO: I added this though reviewers didn't ask for it, since I thought that
+%% this is actually a tricky concept plus the flattening description should
+%% really go here. You might want to change the example
+For the cases where we have N-best parse hypotheses, we use the associated
+parse probabilities (or confidences) to compute expected
+counts. The sentence will then be represented with more tuples,
+corresponding to alternative analyses. For example, if the N-best
+parses include two different roles for dependent ``Basra'', then two
+different $dl$ tuples are included, each with the weighted count that is the
+sum of the confidences of all parses having the respective role.\endnote{
+    The use of expectations with N-best parses is different from 
+    \textbf{d\_50} and \textbf{d\_50\_pm} in \cite{owczarzak07evaluatingmt}
+    in that the latter uses the best-matching pair of trees rather than an 
+    aggregate over the tree sets and they do not use parse
+    confidences.}
+The parse confidence $\tilde{p}$ is normalized so that the
+N-best confidences sum to one. Because the parser is overconfident, we
+explore a flattened estimate: $\tilde{p}(k) =
+\frac{p(k)^\gamma}{\sum_ip(i)^\gamma},$ where $k,i$ index the parse and
+$\gamma$ is a free parameter.
+
+%%MO: I took this out, since this is not true now that you have more info
+%The uniform distribution ($\gamma = 0$) is equivalent to the \cite{owczarzak07evaluatingmt}
+%  \textbf{d\_50} and \textbf{d\_50\_pm} measures.
 
@@
 In the experiments presented later in this article, we explore the
-space of metrics outlined above along multiple dimensions:
+space of metrics outlined above along three dimensions:
 \begin{description}
-\item[Decompositions.] We consider syntactic decompositions of the
-  following types:
+\item[Decompositions.] We compute precision and recall of:
   \begin{description}
   \item[dlh] $\left\langle \textrm{Dependent}, \textrm{arc Label},
-      \textrm{Head}\right\rangle$ Here, every tuple includes two words
-    and the dependency-label between them.
+      \textrm{Head}\right\rangle$ -- full triple
   \item[dl] $\left\langle \textrm{Dependent}, \textrm{arc Label}
-      \right\rangle$ The tuple is the word plus the label of the arc
-      linking it to its head; this decomposition marks how the word
-      fits into its syntactic context (what it modifies).
+      \right\rangle$ --  marks how the word
+      fits into its syntactic context (what it modifies)
     \item[lh] $\left\langle \textrm{arc Label}, \textrm{Head}
-      \right\rangle$ Each tuple here is the dependency relation plus
-      the word being modified; this decomposition implicitly marks how
-      key the word is to the sentence.
+      \right\rangle$ --  implicitly marks how
+      key the word is to the sentence
     \item[dh] $\left\langle \textrm{Dependent}, \textrm{Head}
-      \right\rangle$ This decomposition drops the arc from the
-      dependency-pair tuple, ignoring the syntactic-role information
-      implied in the arc labels.
-    \item[1g,2g] These decompositions are syntactically-null.  They
-      simply include each 1-gram (or 2-gram) without any syntactic
-      labeling at all.  They are thus simple measures of unigram
+      \right\rangle$ -- drops syntactic-role information.
+    \item[1g,2g] -- simple measures of unigram
       (bigram) precision and recall.
   \end{description}
-\item[Size of $n$-best list.] Although the parser can generate
-  considerably more, we consider $n$-best lists of size 1 and of size
-  50.
-\item[Confidence of $n$-best hypotheses.] An expectation requires a
-  probability distribution over the $n$-best list, and we consider
-  three options: uniform, the parser probabilities, and a flattened
-  version of the parser probabilities such that $\tilde{p}(x) =
-  \frac{p(x)^\gamma}{\sum_ip(i)^\gamma}$ (where $\gamma$ is a free
-  parameter) to account for the fact that the parser tends to be
-  over-confident.  In all cases, the probabilities are normalized to
-  sum to one over the the $n$-best list, where the maximum $n$ in this
-  work is 50.  The uniform distribution ($\gamma = 0$) is intended to
-  be equivalent to the \cite{owczarzak07evaluatingmt}
-  \textbf{d\_50} and \textbf{d\_50\_pm} measures.\endnote{
-    \inlinecite{owczarzak07evaluatingmt} report no use of parse
-    confidence weights, and \textbf{d\_50} and \textbf{d\_50\_pm}
-    return the best-matching pair of trees rather than an aggregate
-    over the tree sets.}
-\item[Score combination.] We consider three possible methods for
-  combining scores across decompositions, as suggested in
-  section~\ref{sec:approach}:
+\item[Parser variations.] When using more than one parse, we explore:
   \begin{description}
-  \item[$F\lbrack\cdot\rbrack$] treats each decomposition as
-    contributing $n$ more items to the bag of tokens extracted from
-    the tree, and then computes a global $F$ score.
-  \item[$\mu_{PR}\lbrack\cdot\rbrack$] constructs a precision and
-    recall score from each decomposition, and then does an
-    evenly-weighted harmonic mean of all the component precision and
-    recall scores.
-  \item Tuned weight combination, like the previous $\mu_{PR}$
-    combination, calculates precision and
-    recall separately for each decomposition but performs a linear
-    weighted combination of the resulting precision and recall scores.
-    For this tuning, we note that TERp \cite{snover09terp} provides a
-    tuneable system over a small number of parameters (they include,
-    among others, separate cost for insertions, deletions,
-    substitutions and synonyms) and an optimization algorithm that
-    assigns weights to each parameter. We use the various
-    decompositions' precision and recall as new parameters to the TERp
-    optimizer.
+  \item[Size of $n$-best list.] 50 (as in \cite{owczarzak07evaluatingmt})
+  \item[Parse confidence.] The distribution flattening
+    parameter is varied from $\gamma=0$ (uniform distribution) to $\gamma=1$
+    (no flattening).
+%%MO: I recall that you tried more than one possible gamma in the middle
+%% range to get the 0.25 value, so I think it is misleading to say you looked
+%% at 3 cases.
   \end{description}
+\item[Score combination.] Global $F$ vs.\ component harmonic mean $\mu_{PR}$.
 \end{description}
-
-\subsection{Reference evaluation measures over parallel translations}
-\label{sec:metricsreference}
-For some of the experiments presented here (sections~\ref{sec:hter1}
-and~\ref{sec:hter2}), we have multiple translations of the same source
-segment, each from a different translation engine and each with its
-own HTER score.  For these translations, reporting correlations of
-metric $m$ with absolute HTER scores can conflate $m$'s power in
-identifying which of two candidate translations is better with $m$'s
-(and HTER's) ability to distinguish which source segments are more
-difficult to translate.
-
-To avoid this conflation, on the HTER corpora we report correlations
-on $\bar{\mu}m$ and $\bar{\mu}\textrm{HTER}$ (rather than measure
-$m$ and HTER), where $\bar{\mu}$ represents mean-removal and $t_i$
-represents the $i$th translation of segment $t$:
-\begin{equation}
-  \label{eq:meansub}
-  \bar{\mu}m(t_i) = m(t_i) - \sum_{j=1}^I\frac{m(t_j)}{I}
-\end{equation}
-Mean-removal lowers the correlation $r$ values reported but ensures
-that the reported correlations are among differences in the
-translations rather than among differences in the underlying segments.
-\endnote{Previous work \cite{kahn08metricsmatr} reported HTER
-  correlations against pairwise differences among translation segments
-  derived from the same source segment, unlike the mean-removal
-  suggested in section~\ref{sec:metricsreference}. Those results are similar to
-  those reported in section~\ref{sec:hter1}. However, we believe that pairwise differences
-  introduce problems with the independence assumptions in the
-  Pearson's $r$ tests, and we use mean-removal throughout this work
-  instead.}
-
-%% MO: Okay, but a little hard to figure out.
-When reporting HTER correlations per-sentence, we use length-weighted
-correlation, since unweighted correlations effectively put too much
-emphasis on short sentences.  We do not weight for length for the
-per-document correlations in section~\ref{sec:hter1}, because we
-assume documents to be equally important (regardless of length).
 
@@
 
-To explore members of the DPM metric family, we explore the
-correlations of various configurations against a corpus of human
-judgments of fluency and adequacy.
+We explore various configurations of the DPM by assessing the results
+against a corpus of human judgments of fluency and adequacy,
+specifically the LDC Multiple Translation Chinese corpus parts
+2~\cite{LDC03MTC2} and 4~\cite{LDC06MTC4}, which are composed of
+translations of written Chinese news stories.  These corpora include
+multiple human judgments of fluency and adequacy for each sentence
+(assigned on a five-point scale), with each judgment using a different
+human judge and a different reference translation.  For a
+rough\endnote{Our segment count differs from
+  \inlinecite{owczarzak07evaluatingmt}, who report 16,807 segments
+  over the same corpus, so the results presented here are not directly
+  comparable with that paper. As a result, the baseline (BLEU and TER)
+  per-segment correlations differ slightly (BLEU$_4$ is higher here,
+  while TER here is lower), but the trends in gains over those
+  baselines are very similar.  } comparison with
+\inlinecite{owczarzak07evaluatingmt}, we treat each judgment as a
+separate segment, which yields 16,815 tuples of $\langle$hypothesis,
+reference, fluency, adequacy$\rangle$.  We compute the per-segment
+correlations\endnote{The use of the same hypothesis translations in
+  multiple comparisons in the Multiple Translation Corpus means that
+  scored segments are not strictly independent, but for methodological
+  comparison with prior work, this strategy is preserved.}
+%
+The baselines for comparison are case-sensitive BLEU (4-grams, with add-one smoothing) and TER.
 
-\subsection{Corpus}
-\label{sec:facorpus}
-For these experiments, we use LDC Multiple Translation Chinese corpus
-parts 2~\cite{LDC03MTC2} and 4~\cite{LDC06MTC4}, which are composed of
-translations of written Chinese news stories.  These corpora
-include multiple human judgments of fluency and adequacy for each
-sentence, with each judgment using a different human judge and a
-different reference translation.  Both fluency and adequacy
-assessments are assigned on a five-point scale. For a rough\endnote{Our segment
-  count reported in section~\ref{sec:facorpus} differs from
-  \inlinecite{owczarzak07evaluatingmt}, who report 16,807 segments
-  over the same corpus. We find baseline (BLEU and TER) per-segment correlations slightly
-  different (BLEU$_4$ is higher here, while TER here is lower) than
-  reported there as well, so the results
-  presented here are not directly comparable with that paper, though
-  we demonstrate similar gains over those baselines in essentially the
-  same corpus.
-} comparison with \inlinecite{owczarzak07evaluatingmt}, we treat
-each judgment as a separate segment.
-%
-This treatment of this corpus yields 16,815 tuples of
-$\langle$hypothesis, reference, fluency, adequacy$\rangle$.  In these
-experiments, we extend this tuple with automatic scores derived from
-$\langle$hypothesis, reference$\rangle$ and examine the per-segment
-correlations\endnote{The independence of the translation segments in
-  the Multiple Translation Corpus described in
-  section~\ref{sec:facorpus} is questionable, since the same
-  hypothesis translations are used in multiple items, but for the sake
-  of methodological comparison with prior work, this strategy is
-  preserved.}  between those automatic scores and the arithmetic mean
-of the fluency and adequacy measures.
+Considering only the 1-best parse, we compare DPM with different
+decompositions to the baseline measures.
+Table~\ref{tab:facorr:subgraphs} shows that all decompositions except
+$[dlh]$ have a better per-segment correlation with the
+fluency/adequacy scores than TER or BLEU$_4$. Dependencies $[dl,lh]$
+and string-local n-grams $[1g,2g]$ give similar results, but the
+combination gives further improvement.  The results also confirm, with
+a PCFG, what \inlinecite{owczarzak07evaluatingmt} found with an LFG
+parser: that partial-dependency matches are better correlated with
+human judgments than full-dependency links. Including progressively
+larger chunks of the dependency graph with $F[1g,dl,dlh]$, inspired by
+the BLEU$_k$ idea of progressively larger $n$-grams, did not give an
+improvement over $[dl,lh]$.  $F$ gives substantially better results than
+$\mu_{PR}$, which is never better than BLEU for these decompositions.
 
-\subsection{Exploring decompositions}
-We begin by comparing some of the simplest DPM measures to baseline
-measures case-sensitive BLEU (4-grams, with add-one smoothing) and
-TER.  For these first experiments, we consider only the 1-best parse
-($n=1$).
 \begin{table}
-  \caption{Comparing $dl,lh$ to $dlh$ and baseline per-segment correlations with
-    human fluency and adequacy judgments.}
+  \caption{Per-segment correlation with
+    human fluency/adequacy judgments of baselines and different decompositions.}
   \label{tab:facorr:subgraphs}
@@
     \hline
+    $F[1g,2g,dl,lh]$      &  0.237 \\
+    $F[1g,2g]$            & 0.227 \\
     $F[dl,lh]$ &   0.226 \\
@@
 \end{table}
-In table~\ref{tab:facorr:subgraphs} we see that using the
-partial-label $dl,lh$ decomposition has a better per-segment correlation with the
-fluency/adequacy scores than TER or BLEU$_4$.  These results confirm,
-with a PCFG, what \inlinecite{owczarzak07evaluatingmt} found with
-an LFG parser: that partial-dependency matches are better correlated
-with human judgments than full-dependency links.
 
-In table~\ref{tab:facorr:combinations}, we compare among several other
-members of the DPM family. 
- \begin{table}
-  \begin{tabular*}{2.5in}{lr}
-    \hline
-    metric  &    \multicolumn{1}{c}{$r$} \\
-    \hline
-    $F[1g,2g,dl,lh]$      &  0.237 \\
-    $\mu_{PR}[1g,2g,dl,lh]$ &  0.217 \\
-    \rlcline{1-1} \rlcline{2-2}
-    $F[1g,2g]$            & 0.227 \\
-    $\mu_{PR}[1g,2g]$            & 0.215 \\
+%In table~\ref{tab:facorr:combinations}, we compare combination methods
+%for different decompositions, finding that $F$ consistently outperforms
+%$\mu_{PR}$ as well as the BLEU basedline (from table~\ref{tab:facorr:subgraphs})$\mu_{PR}$ measures are never better than BLEU.
+
+% \begin{table}
+%  \begin{tabular*}{2.5in}{lr}
+%    \hline
+%    metric  &    \multicolumn{1}{$r$} \\
+%    \hline
+%    $F[1g,2g,dl,lh]$      &  0.237 \\
+%    $\mu_{PR}[1g,2g,dl,lh]$ &  0.217 \\
+%    \rlcline{1-1} \rlcline{2-2}
+%    $F[1g,2g]$            & 0.227 \\
+%    $\mu_{PR}[1g,2g]$            & 0.215 \\
 %    $F[1g,dl,dlh]$        & 0.227 \\
-    \rlcline{1-1} \rlcline{2-2}
-    $F[dl,lh]$            &  0.226 \\
-    $\mu_{PR}[dl,lh]$      &  0.208 \\
-    \hline
-  \end{tabular*}
-  \caption{Correlation with human fluency/adequacy judgments. Comparing combination methods $F[\cdot]$ and
-    $\mu_{PR}[\cdot]$, and comparing the $1g,2g,dl,lh$ decomposition
-    with the $dl,lh$ decomposition and other decompositions.}
-  \label{tab:facorr:combinations}
-\end{table}
-Here, we find that combining the decompositions naively in $F[\cdot]$
-(before computing precision and recall) has a better $r$ than the
-harmonic mean of precision and recall subscores in $\mu_{PR}[\cdot]$.
-In fact, $F[\cdot]$ measures always outperform the BLEU baseline (from
-table~\ref{tab:facorr:subgraphs}), and $\mu_{PR}[\cdot]$ measures are
-never better than BLEU.
-We also see that we can combine the benefits of string-local $n$-grams
-($F[1g,2g]$) with the benefits of dependency information ($F[dl,lh]$)
-for a further improved correlation with human judgment, with the best
-correlation in $F[1g,2g,dl,lh]$. Including progressively larger chunks
-of the dependency graph with $F[1g,dl,dlh]$ (not shown in table, but inspired by
-the BLEU$_k$ idea of progressively larger $n$-grams) does not seem to
-be an improvement over $F[dl,lh]$.
+%    \rlcline{1-1} \rlcline{2-2}
+%    $F[dl,lh]$            &  0.226 \\
+%    $\mu_{PR}[dl,lh]$      &  0.208 \\
+%    \hline
+%  \end{tabular*}
+%  \caption{Per-segment correlation with human fluency/adequacy judgments of different combination methods and decompositions.}
+%  \label{tab:facorr:combinations}
+%\end{table}
 
-% (too many chances for zeroes?)
+Table~\ref{tab:facorr:multiparse} shows the impact of using N-best
+parses for different decompositions.
+For the $n=50$ cases, we set $\gamma=0$ to assign uniform
+probabilities, which was slightly better than $\gamma=1$.
+%%MO: I want to take this out, since it really isn't a comparison
+%% if they are using the best match.  I recalled that 0 was better than 
+%% 1 -- change if not
+%to compare as closely as possible
+%to \inlinecite{owczarzak07evaluatingmt}, which includes a
+%\textbf{d\_var\_50} measure with 50-best parses, with ranks but no
+%weights.  
+While not all of these differences are significant, there is a
+consistent trend of correlation $r$ improving with 50 vs. 1 parse.  Tuning
+experiments find that increasing $\gamma$ to 0.25 can increase the $r$
+reported here for $F[1g,2g,dl,lh]$ (but insignificantly).
 
-\subsection{Exploring parse expectations}
-
-To explore the effect of extracting parse feature-counts from a forest
-instead of a single tree, we present in
-table~\ref{tab:facorr:multiparse} several variants, with $n=1$ and
-with $n=50$.
  \begin{table}
@@
   \end{tabular*}
-  \caption{Correlation with human fluency/adequacy judgments. Comparing $n=1$ to $n=50$ for several variants of the DPM measure.}
+  \caption{Per-segment correlation with human fluency/adequacy judgments
+  comparing $n=1$ vs.\ 50 parses for different decompositions.}
   \label{tab:facorr:multiparse}
 \end{table}
-For the $n=50$ cases, we set $\gamma=0$ to assign uniform
-probabilities to the parse forest, to compare as closely as possible
-to \inlinecite{owczarzak07evaluatingmt}, which includes a
-\textbf{d\_var\_50} measure with 50-best parses, with ranks but no
-weights.  While not all of these differences are significant, the
-trend is universally that the correlation $r$ improves as the number
-of parses ($n$) is increased.  Tuning experiments find that increasing
-$\gamma$ to 0.25 can increase the $r$ reported here for
-$F[1g,2g,dl,lh]$ marginally (but insignificantly).
 
-\subsection{Summary}
-
-In this section, we have presented experiments exploring a number of
+In summary, exploring a number of
 variants of the DPM metric against an average fluency/adequacy
-judgment. The experiments suggest a best-case variant, where we set:
+judgment leads to a best-case of:
 \begin{displaymath}
@@
 \end{displaymath}
-in which we choose a $1g,2g,dl,lh$ sub-graph decomposition based on
-the improvements from better sub-graphs
-(table~\ref{tab:facorr:subgraphs}), multiple parses ($n=50$) based on
-table~\ref{tab:facorr:multiparse}, and $\gamma=0.25$.
-%, hinted at by table~\ref{tab:facorr:parseweight}.  
-We use these EDPM parameter-settings in the experiments exploring
-document-level correlations with HTER (section~\ref{sec:hter1}).
+We use this configuration in experiments assessing correlations with HTER.
 
@@
 \label{sec:hter1}
-In this section, we move from an exploration of the space of possible
-DPM metrics to a single chosen DPM-derived metric (EDPM).  Here, we
-compare EDPM to other metrics in correlation with document-level and
-segment-level HTER performance.
-
-\subsection{HTER corpus}
-
-The GALE 2.5 translation corpus is made up of system translations into
-English from three sites.
+In this section, we compare EDPM to baseline metrics in terms of
+document- and segment-level correlation with HTER scores using the
+GALE 2.5 translation corpus.  The corpus includes system translations
+into English from three sites, all of which use system combination to
+integrate results from several systems, some phrase-based and some
+that use syntax on either the source or target side. No system
+provided system-generated parses.
 %
-The three sites all use system combination to integrate results from
-multiple systems, some of which are phrase-based and some which may
-use syntax on either the source or target side. No system provided
-system-generated parses.
-%
-The corpus being translated comes from Arabic and Chinese in four
+The source data comprises Arabic and Chinese in four
 genres: \texttt{bc} (broadcast conversation), \texttt{bn} (broadcast
@@
 sizes shown in table~\ref{tab:galestats}.
+The corpus includes one English reference translation
+\cite{gale08phase2_5references} for each sentence and a system
+translation for each of the three systems. Additionally,
+each of the system translations has a
+corresponding human-targeted reference aligned at the sentence level,
+and an associated so we have available the HTER score at
+both the sentence and document level.
+
 \begin{table}
@@
 \end{table}
-The corpus includes one English reference translation $r_i$
-\cite{gale08phase2_5references} for each sentence $i$ and a system
-translation $t_{i,z}$ for each of the three systems $z$. Additionally,
-each of the system translations of each segment $i$ has a
-corresponding human-targeted reference aligned at the sentence level,
-so we have available the HTER score of each segment HTER$(t_{i,z})$ at
-both the sentence and document level.
 
-As discussed in section~\ref{sec:paradigm}, we report correlations
-here between $\bar{\mu}m$ and $\bar{\mu}\textrm{HTER}$, rather
-than $m$ and HTER directly, to abstract away from variations in the
-underlying documents.
+HTER and automatic scores all degrade on average for more difficult sentences.
+Since there are multiple system translations in this corpus, it is possible to 
+roughly factor out this source of variabilty by correlating mean normalized
+scores,\endnote{Previous work 
+  \cite{kahn08metricsmatr} reported HTER
+  correlations against pairwise differences among translation segments
+  derived from the same source segment as a way of factoring out 
+  sentence difficulty, but this approach introduces problems with 
+  the independence assumptions in the Pearson's $r$ tests.}
+\begin{equation}
+  \label{eq:meansub}
+  \overline{m}(t_i) = m(t_i) - \sum_{j=1}^I\frac{m(t_j)}{I}
+\end{equation}
+where $m$ can be HTER, TER, BLEU or EDPM and $t_i$
+represents the $i$-th translation of segment $t$.
+Mean-removal lowers the correlation $r$ values reported but ensures
+that the reported correlations are among differences in the
+translations rather than among differences in the underlying segments.
 
 In table~\ref{tab:hterperdoc}, we show per-document Pearson's $r$
-between $\bar{\mu}$EDPM and $\bar{\mu}$HTER, as well as two other
-baselines: $\bar{\mu}$TER and $\bar{\mu}$BLEU$_4$.
+between $\overline{\textrm{EDPM}}$ and $\overline{\textrm{HTER}}$, as well as the $\overline{\textrm{TER}}$ and $\overline{\textrm{BLEU}}$$_4$
+baselines. EDPM has the highest correlation in each of the
+subcorpora created by dividing by genre or by source language, as well
+as the corpus as a whole.   In structured data
+(\texttt{bn} and \texttt{nw}), these differences are not always
+significant, but in the unstructured domains (\texttt{wb} and
+\texttt{bc}), EDPM is always significantly better than at least one of
+the comparison baselines.
+
 \begin{subtable}
@@
 \end{subtable}
-We see that the EDPM has the highest correlation in each of the
-subcorpora created by dividing by genre or by source language, as well
-as the corpus as a whole.  Not every difference is above 95\%
-confidence, but the trend is the same across all of the
-subcorpora.
 
-%% MO: I still think this is a bit weird given that what you are
-%% adding is structure. If time, you might want to look at some
-%% example sentences
 Table~\ref{tab:hterpersent} presents per-sentence correlations,
-weighted by sentence length.  Though many of the $r$ values are
-smaller magnitude in this correlation, EDPM again has the largest
-correlation in each category.  TER, however, does quite well with
-length-weighted sentence correlation, with $r$ values within 95\%
+weighted by sentence length in order to get a per-word measure which
+reduces variance across sentences.  (Even with length weighting, the
+$r$ values are smaller magnitude due to the higher variability at the
+sentence level.)  EDPM again has the largest correlation in each
+category, but TER has $r$ values within 95\%
 confidence of EDPM scores on nearly every breakdown.
-%
-In all cases in tables~\ref{tab:hterperdoc} and~\ref{tab:hterpersent},
-EDPM is the most strongly correlated with HTER. In structured data
-(\texttt{bn} and \texttt{nw}), these differences are not always
-significant, but in the unstructured domains (\texttt{wb} and
-\texttt{bc}), EDPM is always significantly better than at least one of
-the comparison baselines.
 
-\section{Weight-tuning to combine syntax and other knowledge sources}
+\section{Combining syntax, edit and semantic knowledge sources}
 \label{sec:hter2}
 
-In the previous section, we observed that TER and EDPM each perform
-better than BLEU when considering mean-removed length-weighted
-correlations with HTER.  We observe that TER's basic operations
-(insertions, deletions, shifts and substitutions) are fundamentally
-different from those of EDPM, which is an $F$-style overlap measure
-and is seeking to make explicit comparison based on an inferred
-syntactic structure.  These two different approaches (word-level and
-syntactic features) are complementary, and we seek in this section to
-explore mechanisms for combining them.
+While the results above show that EDPM is as good or better than other
+measures, the correlation is still low, which is not surprising when
+seeing that the EDPM score is much less than 1 for the good
+translation in the example in section~\ref{sec:approach}. For that
+reason, we investigated combining the alternative wording features of
+TERp with the syntactic features of EDPM.
 
-As discussed in section~\ref{sec:intro}, the TER Plus (TERp) research
-tools \cite{snover09terp} provide an optimizer for weighting multiple
-simple subscores, including selecting weights for the TER basic
-operations. TERp's feature list extends TER to additionally include
-synonymy, stemming, and automatically-derived paraphrases.
+The TER Plus (TERp) tools \cite{snover09terp} provide an optimizer for
+weighting multiple simple subscores.  The TERp optimizer performs a
+hill-climbing search, with randomized restarts, to maximize the
+correlation of a linear combination of the subscores with a set of
+human judgments.  Within the TERp framework the subscores are the
+counts of the various edit types normalized for the length of the
+reference, where the counts are determined after aligning the MT
+output to the reference using default edit costs.
 
-%% MO: important to describe basic model before listing features.
-The TERp optimizer performs a hill-climbing search, with randomized
-restarts, to maximize the correlation of a linear combination of the
-features with a set of human judgments.  Within the TERp framework the
-features are the counts of the various edit types normalized for the
-length of the reference so that the dot product of the features with a
-vector of edit costs will give the TERp score for a given sentence.  In
-these experiments, we follow~\inlinecite{snover09terp} by first
-aligning the MT outputs with the reference, using TERp with a set of
-default edit costs, to find the counts of the various edit types.
-These alignments and edit costs are then held constant to optimize the
-edit costs. In the case of comparing TERp without the paraphrasing
-feature, a separate alignment was performed without that feature
-enabled.  Additional features can be added to the TERp optimization
-framework and optimized in the same manner as the edit costs.
-
-In the experiments presented in this section, we extend this feature
-list by including the syntactic overlap features, and we use the TERp
-optimizer to tune the relative weight of each feature.
-
-\subsection{Features}
-
-We train a model using features from one or more of the following
-sets:
+The experiments presented here leverage the TERp optimizer but extend
+the set of subscores by including the syntactic and n-gram overlap
+features modified to reflect false and missed detection rates (rather than precision and recall as used in the DPM family). The subscores explored include:
 \begin{description}
-\item[E]: the fully syntactic features from the DPM family.
-  Specifically, this feature set includes error-counts
-  ($\frac{|\textrm{hyp}| - |\textrm{correct}|}{|\textrm{ref}|}$ and
-  $\frac{|\textrm{ref}| - |\textrm{correct}|}{|\textrm{ref}|}$) 
-  for the $dl$, $lh$, $dlh$, and $dh$ decompositions of the syntactic
-  structure, using $n=50$ and $\gamma=0.25$.  The \textbf{E} set
-  includes 8 features.
-\item[N]: the non-syntactic features from the DPM family;
-  specifically, error-counts for the $1g$ and $2g$ decompositions.
-  The \textbf{N} set includes 4 features.
-\item[T]: the features from basic TERp, excluding paraphrases, which include a separate cost
+\item[E]: the 8 fully syntactic subscores from the DPM family, including
+  false/miss error rates for the $dl$, $lh$, $dlh$, and $dh$ decompositions,
+  using $n=50$ and $\gamma=0.25$. 
+\item[N]: the 4 n-gram subscores from the DPM family;
+  specifically, error rates for the $1g$ and $2g$ decompositions.
+\item[T]: the 11 subscores from TERp, which include a separate cost
   for each of matches, insertions, deletions, substitutions, shifts, and
-  synonym matches and stem matches. The \textbf{T} set includes 7 features.
-\item[P]: features from the automatically-derived TERp paraphrase
-  table. These are four features which are a function of the edit
-  distance between the two phrases in each paraphrase and the
-  probability of the paraphrase.
-\item[B]: two ``length penalties'': brevity and prolixity. One is assigned the count of
-  hypothesis words beyond the count of reference words, and one
-  assigned the count of reference words beyond the count of hypothesis
-  words.
-  % MO: out of curiosity, why do you need 2 features?
-  We assign two features here to let the system separately penalize
-  ``too short'' translations and ``too long''.
+  synonym and stem matches, edit distances between paraphrases, and the. 
 \end{description}
 
-\subsection{Corpus}
+For these experiments, we again use the GALE 2.5 data, but with 2-fold
+cross-validation in order to have independent tuning and test data.
+Documents are partitioned randomly, such that each subset has the same
+document distribution across source-language and genre. The objective
+is length-weighted per-sentence correlation with HTER, using
+mean-removed scores as before. In figure~\ref{fig:tuneresults}, we
+plot the Pearson's $r$ (with 95\% confidence interval) for the results
+on the two test sets combined, after linearly normalizing the
+predicted scores to account for magnitude differences in the learned
+weight vectors.  The baseline scores, which involve no tuning, are not
+normalized.
 
-For these tuning experiments we use the same GALE 2.5 documents and
-sentences from the previous section.  Since this approach requires
-tuning, we assign documents randomly into two groups, such that each
-group has the same document distribution across source-language and
-genre.
-%
-These two corpus halves are used in two-fold cross-validation, testing
-on one and training on the other.  We tune on (and report correlation
-with) length-weighted per-sentence correlation with HTER, using
-mean-removed scores as before.
-
-% tuning target: weighted correlation with mean-removed segment HTER
-\subsection{Effects of features on tuning}
-% [we use weighted correlation to avoid emphasis on short segments
-% (which hurts document-level and system-level utility) ]
-In figure~\ref{fig:tuneresults}, we plot the Pearson's $r$ (with 95\%
-confidence interval) for the results on the two test sets (each
-trained on one and tested on the other).
 \begin{figure}
   \begin{center}
-    \includegraphics[scale=0.45]{tuning.eps}
+    \includegraphics[scale=0.45]{tuning}
   \end{center}
-  \caption{Pearson's $r$ values for various feature tunings, with 95\%
-    confidence intervals. BLEU$_4$ (add-one smoothed), TER, and EDPM
-    scores are provided on the left for comparison.}
+  \caption{Pearson's $r$ values for the two halves of the training
+    set, with 95\% confidence intervals. }
   \label{fig:tuneresults}
 \end{figure}
-%
-For each set of features, we plot the $r$ (and confidence) on both
-halves of the corpus (for each half, tuned on the other half). In the
-leftmost group of the figure, we include baselines: EDPM and BLEU and
-TER (which are not tuned).  Next, the tuned E feature, which (by
-itself) does not perform as well as TER or EDPM.
-%
-The remaining groups of three show that for each set of features, as E
-(and EB) are included with others, correlation improves (though not
-significantly).  These all move in the same direction. By the same
-measurement, P features help beyond T -- as reported elsewhere -- but
-N features do not seem to find a better correlation than TP features.
-% MO suggestions:
 
-Overall, we find that this approach for combining the underlying idea
-of EDPM with the underlying ideas of TERp is promising, in that the E
-features offer small (though non-significant) improvements to the
-features already within TERp. However, this combination approach still
-has more in common with the $\mu_{PR}$ combination method (explored in
-section~\ref{sec:faexpts}) than with EDPM, which suggests a reason
-that the EN features --- though tuned --- do not outperform EDPM.
-%% do n-fold with >2 N -- better robustness to variation?
+The figure shows that TER and EDPM are significantly more correlated with HTER than BLEU, consistent with the overall results of the previous section.
+The N+E combination outperforms E alone (i.e.\ it is helpful to use both n-gram and dependency overlap), but the combination gives lower performance than EDPM because of the particular combination technique (essentially the $\mu_{PR}$ vs.\ F difference).
+Both findings are consistent with the fluency/adequacy experiments.
+The TERp features, which account for synonym/paraphrains differences, have much higher correlation with HTER than the syntactic E+N subscores.
+However, a significant additional improvement is obtained by adding syntactic features to TERp (T+E).
+Adding the n-gram features to TERp (T+N) gives almost as much improvement, probably because most dependencies are local.
+There is no further gain from using all three aubscore types.
 
@@
 \label{sec:conclusion}
-In this research, we explore dependency pair match (DPM) a family of
+In summary, we explore dependency pair match (DPM) a family of
 syntactically-oriented measures.  Through a corpus of human fluency
@@
 underlying document- and segment-level difficulty).
-We experiment with including syntactic features in TERp-style tuning,
-with linear weight combination, and find that including these features
-improves correlation with mean-subtracted HTER, although these
-improvements do not reach significance.
+We experiment with including syntactic and synonym/paraphrase 
+features in a TERp-style linear combination, and find that the combination
+improves correlation with mean-subtracted HTER over either method alone.
 
-Several areas of future work are open to us. One obstacle to using
-this method widely is the computational cost of parsing compared to
-word-based measures
-such as BLEU or TER.  Using these syntactic techniques may be possible
-as a late-pass evaluation, to identify how candidate translation
-systems are performing overall.  Alternatively, the various
-tree-decompositions of the DPM family could be used as system
-diagnostics, by looking at relative quality of these component scores
-compared to those of an alternative candidate system.
-%
-Another possible approach is to store packed forests
-\cite{huang08packedforests} rather than generating an $n$-best list
-only to sum across it again in calculating the expectation.
+One difference with respect to the work of \inlinecite{owczarzak07evaluatingmt} is the use of a PCFG vs.\ an LFG parser. The PCFG has the advantage that it
+is publicly available and easily adaptable to new domains.
+However, the performance varies depending on the amount of labeled data for the domain, which raises the question of how sensitive EDPM and related measures are to parser quality.
 
-Another research area relates to the quality of the parser.  In
-this work, we demonstrated that the
-\inlinecite{charniak-johnson:2005:ACL} PCFG parser could serve in
-place of the LFG parser used in
-\inlinecite{owczarzak07evaluatingmt}. PCFG parsers, however, are
-fairly easily adaptable to another domain. One research
-direction explores the intersection of genre (and genre-mismatch),
-parser quality, and prediction of human performance.
-% Another
-% consideration is the speed of the parser: parse time currently takes
-% much, much longer than BLEU, TER, or other word-based measures. 
-We are interested in the trade-offs --- if any --- among parse
-quality, genre adaptation, and predictive power as a translation metric.
+A limitation of this method for MT evaluation is the computational
+cost of parsing compared to word-based measures such as BLEU or TER.
+Using these syntactic techniques may be possible as a late-pass
+evaluation, to identify how candidate translation systems are
+performing overall.  Alternatively, the various tree-decompositions of
+the DPM family could be used as system diagnostics, by looking at
+relative quality of these component scores compared to those of an
+alternative candidate system.
+
+%%MO: put back if room,
+%Another possible approach is to store packed forests
+%\cite{huang08packedforests} rather than generating an $n$-best list
+%only to sum across it again in calculating the expectation.
 

### 2009-08-11 reviewers/review-response.tex
Mari's changes to the review-response

--- 2009-08-05 reviewers/review-response.tex
@@
 
+Note from the authors: all references to sections and tables in our responses are with respect to numbering in the revised version of the paper.
+
 \section{Comments for the Author: Reviewer \#1}
@@
 \begin{response}
-  Thank you for this feedback. Our responses are below.
+  Thank you for this very useful feedback. Our responses are below.
 \end{response}
@@
 \begin{response}
-  We added discussion in section 3 and refer back to the example in
-  section two, which illustrates this point. [TO DO -- emend the discussion].
+  We added discussion in the second paragraph of section 3.1 and refer back to the new example in
+  section two, which illustrates this point. 
 \end{response}
@@
 \begin{response}
-  Section 4 has now included several additional sentences to clarify
-  these points.
+  Section 4 has now included information on granularity (5-point scale),
+genre (Chinese news stories), and a note that these are per-segment correlations.
 \end{response}
@@
 \begin{response}
- [TO DO -- add a footnote]
+ The correlations are similar to those reported in Owczarzak et al., and they
+are low in part because of the fact that they are computed at the segment level.
+We added a parenthetical in section 5, paragraph 4, indicating that sentence-level correlations are lower than document-level correlations because of the greater variability.
+Of course, there are also differences simply associated with the fact that these are different data sets.
 \end{response}
@@
 \begin{response}
-  The sentence at the end of the section is revised to indicate that
-  ``in the unstructured domains (\texttt{wb} and \texttt{bc}), EDPM is
+  The sentence at the end of the second-to-last paragraph of section 5
+  is revised to indicate that ``In structured data (\texttt{bn} and
+  \texttt{nw}), these differences are not always significant, but in
+  the unstructured domains (\texttt{wb} and \texttt{bc}), EDPM is
   always significantly better than at least one of the comparison
@@
     eliminate section 6, we have chosen to compress it, but we have
-    adopted the rest of the suggestions above.
+    adopted the rest of the suggestions above.  Details about the 
+    changes to section 6 are in the response to the last comment.  We think that these 
+    changes make the contribution
+    of section 6 more clear.
   \end{response}
@@
   \begin{response}
-    We include an example in section 2 to address this point.  [TO DO
-    -- using motivating examples?]
+    We include an example in section 2 to address this point. 
   \end{response}
@@
   \begin{response}
-    Done.
+    Done in the section 2 example.
   \end{response}
@@
   \begin{response}
-    This section has also been compressed.
+    This paragraph has also been compressed.
   \end{response}
@@
   \begin{response}
-    [TO DO -- expand to rebuttal on expanding similarities and
-    differences -- note that sec. 4 is essentially this].  The
-    reference has been updated.
+    Both references are cited in the introduction.  In the comparative
+    discussions, the earlier reference is cited since we were not aware of
+    the MT journal paper when doing the work and experiments were designed
+    with the first paper in mind.  We feel that our paper did have 
+    discussion of the similarities and differences, but with the compression
+    of other material these are better highlighted in the revisions. In addition, the last
+    paragraph of section 1 now states that section 4 is in part aimed at making
+    the comparison.
   \end{response}
@@
   \begin{response}
-    [TO DO: motivating examples?]
+  We have added a sentence at the end of the paragraph pointing to the example
+to motivate the ``overweighting''. (now 1st paragraph on page 4)
   \end{response}
@@
   \begin{response}
-    [TO DO] reduce the discussion of $\mu_{PR}$ further? [TO DO:
-    revisit after compression]
+  We have compressed the discussion somewhat in section 2, substantially in section 3 (to a single line), and combined the original Tables I and II without the $\mu$ results, and then mention $\mu$ findings in a sentence of the text.
+Since
+this finding is significant and provides insight into differences between other methods in terms of score combination, we did not want to eliminate it entirely.
   \end{response}
@@
   \begin{response}
-    [TO DO]
+    As mentioned above, this was reduced to one line.
   \end{response}
@@
   \begin{response}
-    [TO DO]
+    This method is used for evaluating the evaluation metric, not for
+    evaluation itself, so comparability of scores across sentences is
+    not really an issue for us.  That said, we agree that mean 
+    normalization is not possible in all
+    scenarios and in fact did not use it in section four.  Hence, we
+    moved the mean normalization description to section 5, since it
+    is only used when we have multiple translations.  
   \end{response}
@@
   \begin{response}
-    Thank you! this oversight is now repaired and $t$ is explicit.
+    Fixed, but to save space the equation is now inline and not numbered.
   \end{response}
@@
   \begin{response}
-    Section 4 is intended to match the Owczarzak \textit{et al.} data
-    selection, but sections 5 and 6 do not; in sections 5 and 6 the
-    data available \emph{does} allow multiple translations of the same
-    segment and thus we can use mean-subtraction. Mean subtraction is
-    only used in sections 5 and 6, and the discussion of
-    mean-subtraction has been moved to the beginning of section 5 for
-    clarity.
+    Section 4 is intended to match the Owczarzak \textit{et al.} paradigm
+    as closely as possible, but sections 5 and 6 do not since the focus 
+    is on HTER. As mentioned above, mean subtraction is {\bf not} used 
+    in section 4 and discussion of that was moved for clarity.  We think
+    that this concern is really a lack of clarity in the original submission
+    which has hopefully now been fixed.
   \end{response}
@@
   \begin{response}
-    [TO DO]
+   Yes.  This is no longer in the tables because of our response to
+   your earlier comment abut F vs. $\mu$, but we include this observation
+   in the text (end of second paragraph, section 4)
   \end{response}
@@
   \begin{response}
-    [TO DO]
+  We have tried to make it more clear that only the sentence-level results (not document level) use length-normalized scores in computing correlations.
+We explain that this is ``in order to get a per-word measure which reduces variance across sentences'' and we clarify that this is not giving the longer sentences more weight but rather the same weight by using the term ``normalized'' rather than ``weighted'' and saying that it is per-word.
   \end{response}
@@
 \begin{response}
-  [TO DO]: We have made a number of changes to section 6.  We have
+  As described earlier, we have made a number of changes to section 6.  We have
   revisited the two-fold validation and provided a new statistical
-  analysis that simplifies both the presentation (figure 4) and the
-  discussion (elsewhere in section 6).
+  analysis on a smaller set of results that simplifies both the presentation (figure 4) and the
+  discussion (elsewhere in section 6). In particular, the changes
+  include:
+  \begin{enumerate}
+  \item dropping the discussion of $B$ brevity penalties, since these are not in the  referenced TERp and are not relevant to the point of the combining
+    syntax and paraphrasing features
+  \item collapsing the $T$ (edit and synonymy) and $P$ (paraphrasing)
+    categories together (now presented together as $T$), again since the
+    separation does not contribute to the point of the paper
+  \item folding together and normalizing the values produced for the
+    cross-validation-tuned combination measures to provide a single value and
+    confidence interval rather than presenting both splits, which leads to a simpler 
+    figure with more significant differences.
+  \item adding a motivational introductory paragraph
+  \end{enumerate}
 \end{response}

### 2009-08-11 mtjournal.tex
minor caption revision.
Mari's further compression work

@@
   matches with standard local n-gram matches using a statistical
-  parser, and taking advantage of N-best parse probabilities in deriving
-  scores for the hypothesized
-  translation.  The new scoring metric, Expected Dependency
-  Pair Match (EDPM), is shown to be superior to BLEU and TER in terms
-  of correlation to human judgments and as a per-document and
-  per-sentence predictor of HTER. Further, we combine the syntactic features of EDPM with the
+  parser, and taking advantage of N-best parse probabilities.  The new scoring metric, Expected Dependency
+  Pair Match (EDPM), is shown to outperform BLEU and TER in terms
+  of correlation to human judgments and as a predictor of HTER. Further, we combine the syntactic features of EDPM with the
   alternative wording features of TERp, showing a benefit to accounting for syntactic structure on top of
@@
 translations may be quite different in surface form. 
-Currently, the most popular evaluation measures include a measure
-based on $n$-gram precision known as BLEU \cite{papineni02bleu} and
-the edit-distance measure Translation Edit Rate (TER)
-\cite{snover06ter}.  These measures can only account for variability when 
+Currently, the most popular approaches are BLEU \cite{papineni02bleu},
+based on $n$-gram precision, and Translation Edit Rate (TER), an
+edit distance \cite{snover06ter}.
+%evaluation measures include a measure
+%based on $n$-gram precision known as BLEU \cite{papineni02bleu} and
+%the edit-distance measure Translation Edit Rate (TER)
+%\cite{snover06ter}.  
+These measures can only account for variability when 
 given multiple translations, and studies show that they may
-not accurately track translation quality both empirically
-\cite{charniak03syntaxlmmt} and theoretically
-\cite{callisonburch06bleuproblems}.
+not accurately track translation quality 
+%both empirically
+\cite{charniak03syntaxlmmt,callisonburch06bleuproblems} 
+%and theoretically
+%\cite{callisonburch06bleuproblems}.
 
@@
 $n$-gram precision in various configurations of constituency and
-dependency trees \cite{liu05syntaxformteval};\endnote{The dependency-based SParseval measure
-\cite{roark06:sparseval}, designed as a parse-quality metric for
-speech, is a similar approach, in that it is an F-measure over a
-decomposition of reference and hypothesis trees.} and the \textbf{d} and
-\textbf{d\_var} measures proposed by \inlinecite{owczarzak07evaluatingmt} that compare relational tuples derived from a
+dependency trees \cite{liu05syntaxformteval};
+%%MO: Since sparseval is cited elsewhere, I don't think we need the endnote
+%\endnote{The dependency-based SParseval measure
+%\cite{roark06:sparseval}, designed as a parse-quality metric for
+%speech, is a similar approach, in that it is an F-measure over a
+%decomposition of reference and hypothesis trees.} 
+and the \textbf{d} and
+\textbf{d\_var} measures proposed by \inlinecite{owczarzak07evaluatingmt}
+%%MO: This is a hack to get the second year without the name
+(2007b) 
+that compare relational tuples derived from a
 lexical functional grammar (LFG)
-over reference and hypothesis translations.
+over reference and hypothesis translations.\endnote{
+  \inlinecite{owczarzak07evaluatingmt} extend their previous line of
+  research \cite{owczarzak07labelleddepseval}  by variably-weighting
+%%MO: you actually do variable weighting using parse probabilities, so
+%% would be good to mention what they use.  If you can say this in a few
+%% words, then you won't add a line
+  dependencies and by including synonym matching, two directions not
+  pursued here. Hence, the earlier paper is cited in comparisons.
+}
 %
@@
 
-We describe our approach to including syntactic information in MT
+We describe our approach to including syntax in MT
 evaluation by outlining a family of metrics in
-section~\ref{sec:approach} and provide implementation details in
+section~\ref{sec:approach} and implementation details in
 section~\ref{sec:paradigm}.  Section~\ref{sec:faexpts} examines the
-correlation of several members of this family with human judgments of
-fluency and adequacy, using the paradigm from
-\cite{owczarzak07evaluatingmt} to provide comparisons to that work and
+correlation of members of this family with human judgments of
+fluency and adequacy, using the 
+\cite{owczarzak07evaluatingmt} paradigm to provide comparisons and
 select a best case configuration, Expected Dependency Pair Match
@@
 extensions of the dependency-pair F measures found in
-\inlinecite{owczarzak07labelleddepseval}\endnote{
-  \inlinecite{owczarzak07evaluatingmt} extend their previous line of
-  research \cite{owczarzak07labelleddepseval} by variably-weighting
-%%MO: you actually do variable weighting using parse probabilities, so
-%% would be good to mention what they use.  If you can say this in a few
-%% words, then you won't add a line
-  dependencies and by including synonym matching, two directions not
-  pursued here.
-}.  Rather than comparing
+\inlinecite{owczarzak07labelleddepseval}.  Rather than comparing
 string sequences, as BLEU does with its $n$-gram precision, this
@@
 %  $\frac{1}{4}$: thus $F[dlh] = \frac{2}{7}$
-  \caption{Example hypothesis and reference dependency trees and the
-    $dlh$ decomposition of each.}
+  \caption{Example dependency trees and their $dlh$ decompositions.}
   \label{fig:decompexample}
@@
   \caption{The $dl$ and $lh$ decompositions of the hypothesis tree in
-    figure~\ref{fig:decompexample}. The items extracted here
-    are individually less restrictive in their ability to match
-    against reference tuples. 
+    figure~\ref{fig:decompexample}.
 %     $\precision{dl} = \frac{2}{3}$,
@@
 words in the tree, are included multiple times in the decomposition
-(once for each inbound link).  This ``overcounting'' leverages
+(once for each inbound link).  This ``overweighting'' leverages
 syntactic indications of which words are more
@@
 
-A probabilistic parser provides confidences associated with parses
-in an $n$-best list, which we can use to compute
-expected counts for each decomposition in both the reference and
+A statistical parser provides confidences associated with parses
+in an $n$-best list, which we use to compute
+expected counts for each decomposition in both reference and
 hypothesized translations. The expected counts lead to partial matches
-which are used to compute precision and recall as weighted counts.
-Using the expectations can help cope with both error
+(or weighted counts) used in computing precision and recall.
+This approach addresses both error
 in the best parse and ambiguity in the translations (reference and
@@
 
-\section{Experimental implementation}
-\label{sec:paradigm}
-% In this section, we describe the details of the implementation and
-% experimental variables explored in the experiments presented here.
-
-\subsection{Parsing and dependency extraction}
+\section{Parsing and dependency extraction}
 
@@
 
-\subsection{Dimensions of metric family}
-\label{sec:metricdimensions}
-In the experiments presented later in this article, we explore the
-space of metrics outlined above along three dimensions:
+\section{Correlation with human judgments of fluency \& adequacy}
+\label{sec:faexpts}
+
+We explore various configurations of the DPM by assessing the results
+against a corpus of human judgments of fluency and adequacy,
+specifically the LDC Multiple Translation Chinese corpus parts
+2~\cite{LDC03MTC2} and 4~\cite{LDC06MTC4}, which are composed of
+translations of written Chinese news stories.  These corpora include
+multiple human judgments of fluency and adequacy for each sentence
+(assigned on a five-point scale), with each judgment using a different
+human judge and a different reference translation.  For a
+rough\endnote{Our segment count differs slightly from
+  \inlinecite{owczarzak07evaluatingmt}
+  for the same corpus: 16,807 vs.\ 16,815. 
+  As a result, the baseline 
+  per-segment correlations differ slightly (BLEU$_4$ is higher here,
+  while TER here is lower), but the trends in gains over those
+  baselines are very similar.  } comparison with
+\inlinecite{owczarzak07evaluatingmt}, we treat each judgment as a
+separate segment, which yields 16,815 tuples of $\langle$hypothesis,
+reference, fluency, adequacy$\rangle$.  We compute the per-segment
+correlations\endnote{The use of the same hypothesis translations in
+  multiple comparisons in the Multiple Translation Corpus means that
+  scored segments are not strictly independent, but for methodological
+  comparison with prior work, this strategy is preserved.}
+%
+The baselines for comparison are case-sensitive BLEU (4-grams, with add-one smoothing) and TER.
+
+The specific dimensions of DPM explored include:
 \begin{description}
@@
 \end{description}
-
-\section{Correlation with human judgments of fluency \& adequacy}
-\label{sec:faexpts}
-
-We explore various configurations of the DPM by assessing the results
-against a corpus of human judgments of fluency and adequacy,
-specifically the LDC Multiple Translation Chinese corpus parts
-2~\cite{LDC03MTC2} and 4~\cite{LDC06MTC4}, which are composed of
-translations of written Chinese news stories.  These corpora include
-multiple human judgments of fluency and adequacy for each sentence
-(assigned on a five-point scale), with each judgment using a different
-human judge and a different reference translation.  For a
-rough\endnote{Our segment count differs from
-  \inlinecite{owczarzak07evaluatingmt}, who report 16,807 segments
-  over the same corpus, so the results presented here are not directly
-  comparable with that paper. As a result, the baseline (BLEU and TER)
-  per-segment correlations differ slightly (BLEU$_4$ is higher here,
-  while TER here is lower), but the trends in gains over those
-  baselines are very similar.  } comparison with
-\inlinecite{owczarzak07evaluatingmt}, we treat each judgment as a
-separate segment, which yields 16,815 tuples of $\langle$hypothesis,
-reference, fluency, adequacy$\rangle$.  We compute the per-segment
-correlations\endnote{The use of the same hypothesis translations in
-  multiple comparisons in the Multiple Translation Corpus means that
-  scored segments are not strictly independent, but for methodological
-  comparison with prior work, this strategy is preserved.}
-%
-The baselines for comparison are case-sensitive BLEU (4-grams, with add-one smoothing) and TER.
 
@@
 Since there are multiple system translations in this corpus, it is possible to 
-roughly factor out this source of variabilty by correlating mean normalized
+roughly factor out this source of variability by correlating mean normalized
 scores,\endnote{Previous work 
   \cite{kahn08metricsmatr} reported HTER
-  correlations against pairwise differences among translation segments
-  derived from the same source segment as a way of factoring out 
-  sentence difficulty, but this approach introduces problems with 
-  the independence assumptions in the Pearson's $r$ tests.}
-\begin{equation}
-  \label{eq:meansub}
-  \overline{m}(t_i) = m(t_i) - \sum_{j=1}^I\frac{m(t_j)}{I}
-\end{equation}
+  correlations against pairwise differences among translations
+  derived from the same source to factor out 
+  sentence difficulty, but this violates
+  independence assumptions used in the Pearson's $r$ tests.}
+$\overline{m}(t_i) = m(t_i) - \frac{1}{I}\sum_{j=1}^Im(t_j)$
+%\begin{equation}
+%  \label{eq:meansub}
+%  \overline{m}(t_i) = m(t_i) - \sum_{j=1}^I\frac{m(t_j)}{I}
+%\end{equation}
 where $m$ can be HTER, TER, BLEU or EDPM and $t_i$
 represents the $i$-th translation of segment $t$.
-Mean-removal lowers the correlation $r$ values reported but ensures
+Mean-removal ensures
 that the reported correlations are among differences in the
@@
 
-Table~\ref{tab:hterpersent} presents per-sentence correlations,
-weighted by sentence length in order to get a per-word measure which
+Table~\ref{tab:hterpersent} presents per-sentence correlations based on
+scores normalized by sentence length in order to get a per-word measure which
 reduces variance across sentences.  (Even with length weighting, the
@@
 
-While the results above show that EDPM is as good or better than other
-measures, the correlation is still low, which is not surprising when
-seeing that the EDPM score is much less than 1 for the good
-translation in the example in section~\ref{sec:approach}. For that
+While our results show that EDPM is as good or better than other
+measures, the correlation is still low, which is consistent with
+the example in section~\ref{sec:approach}, where the EDPM score is much less than 1 for the good
+translation. For that
 reason, we investigated combining the alternative wording features of
-TERp with the syntactic features of EDPM.
+TERp with the EDPM syntactic features.
 
-The TER Plus (TERp) tools \cite{snover09terp} provide an optimizer for
+The TERp tools \cite{snover09terp} provide an optimizer for
 weighting multiple simple subscores.  The TERp optimizer performs a
@@
 
-The experiments presented here leverage the TERp optimizer but extend
+The experiments here leverage the TERp optimizer but extend
 the set of subscores by including the syntactic and n-gram overlap
-features modified to reflect false and missed detection rates (rather than precision and recall as used in the DPM family). The subscores explored include:
+features (modified to reflect false and missed detection rates for the TERp
+format rather than precision and recall). The subscores explored include:
 \begin{description}
-\item[E]: the 8 fully syntactic subscores from the DPM family, including
-  false/miss error rates for the $dl$, $lh$, $dlh$, and $dh$ decompositions,
-  using $n=50$ and $\gamma=0.25$. 
+\item[E]: the 8 fully syntactic subscores from the EDPM family, including
+  false/miss error rates for the $dl$, $lh$, $dlh$, and $dh$ decompositions. 
 \item[N]: the 4 n-gram subscores from the DPM family;
   specifically, error rates for the $1g$ and $2g$ decompositions.
-\item[T]: the 11 subscores from TERp, which include a separate cost
-  for each of matches, insertions, deletions, substitutions, shifts, and
+\item[T]: the 11 subscores from TERp, which include matches, insertions, deletions, substitutions, shifts, and
   synonym and stem matches, edit distances between paraphrases, and the. 
@@
 document distribution across source-language and genre. The objective
-is length-weighted per-sentence correlation with HTER, using
+is length-normalized per-sentence correlation with HTER, using
 mean-removed scores as before. In figure~\ref{fig:tuneresults}, we
@@
   \end{center}
-  \caption{Pearson's $r$ values for the two halves of the training
-    set, with 95\% confidence intervals. }
+  \caption{Pearson's $r$ for various feature tunings, with 95\%
+    confidence intervals.}
   \label{fig:tuneresults}
@@
 The figure shows that TER and EDPM are significantly more correlated with HTER than BLEU, consistent with the overall results of the previous section.
-The N+E combination outperforms E alone (i.e.\ it is helpful to use both n-gram and dependency overlap), but the combination gives lower performance than EDPM because of the particular combination technique (essentially the $\mu_{PR}$ vs.\ F difference).
+The N+E combination outperforms E alone (i.e.\ it is helpful to use both n-gram and dependency overlap) but gives lower performance than EDPM because of the particular combination technique.
 Both findings are consistent with the fluency/adequacy experiments.
-The TERp features, which account for synonym/paraphrains differences, have much higher correlation with HTER than the syntactic E+N subscores.
+The TERp features, which account for synonym/paraphrase differences, have much higher correlation with HTER than the syntactic E+N subscores.
 However, a significant additional improvement is obtained by adding syntactic features to TERp (T+E).
 Adding the n-gram features to TERp (T+N) gives almost as much improvement, probably because most dependencies are local.
-There is no further gain from using all three aubscore types.
+There is no further gain from using all three subscore types.
 
@@
 \label{sec:conclusion}
-In summary, we explore dependency pair match (DPM) a family of
-syntactically-oriented measures.  Through a corpus of human fluency
+In summary, we explore a family of dependency pair match measures.  Through a corpus of human fluency
 and adequacy judgments,
@@
 correlation with human judgments and as a per-document and
-per-sentence predictor of HTER (using mean-subtraction to account for
-underlying document- and segment-level difficulty).
-We experiment with including syntactic and synonym/paraphrase 
+per-sentence predictor of mean-normalized HTER.
+We also experiment with including syntactic and synonym/paraphrase 
 features in a TERp-style linear combination, and find that the combination
-improves correlation with mean-subtracted HTER over either method alone.
+improves correlation with HTER over either method alone.
 
@@
 
-A limitation of this method for MT evaluation is the computational
+A limitation of this method for MT system tuning is the computational
 cost of parsing compared to word-based measures such as BLEU or TER.
-Using these syntactic techniques may be possible as a late-pass
-evaluation, to identify how candidate translation systems are
-performing overall.  Alternatively, the various tree-decompositions of
-the DPM family could be used as system diagnostics, by looking at
+Two alternative low-cost use scenarios include late-pass
+evaluation, for choosing between different system architectures,
+or system diagnostics, looking at
 relative quality of these component scores compared to those of an
-alternative candidate system.
+alternative configuration.
 

### 2009-08-11 reviewers/review-response.tex
additional comments on why the detailed list of decompositions is moved to s. 4

@@
     which has hopefully now been fixed.
+
+    In addition, we move the detailed list of decompositions explored
+    from section 3 to section 4, since section 4 explores
+    decompositions --- later sections use fixed (s.~5) and
+    greatly-reduced (s.~6) sets of decompositions, so the detailed
+    list fits more succinctly and appropriately in section 4 alone.
   \end{response}

### 2009-08-11 reviewers/review-response.tex
one more change to the review response, suggested by Mari

@@
   \begin{response}
-  We have compressed the discussion somewhat in section 2, substantially in section 3 (to a single line), and combined the original Tables I and II without the $\mu$ results, and then mention $\mu$ findings in a sentence of the text.
-Since
+  We have compressed the discussion somewhat in section 2,
+  substantially in section 3 (to a single line, now in section 4), and combined the original Tables I and II without the $\mu$ results, and then mention $\mu$ findings in a sentence of the text. Since
 this finding is significant and provides insight into differences between other methods in terms of score combination, we did not want to eliminate it entirely.

### 2009-08-11 mtjournal.tex
missing paradigm section label
grammar fixups at last minute

@@
 %there is little data for tuning.
-Other metrics model syntactically-local (rather than string-local)
+Other metrics modeling syntactically-local (rather than string-local)
 word-sequences include: tree-local
@@
 \section{Parsing and dependency extraction}
+\label{sec:paradigm}
 
@@
 separate segment, which yields 16,815 tuples of $\langle$hypothesis,
-reference, fluency, adequacy$\rangle$.  We compute the per-segment
-correlations\endnote{The use of the same hypothesis translations in
+reference, fluency, adequacy$\rangle$.  We compute per-segment
+correlations.\endnote{The use of the same hypothesis translations in
   multiple comparisons in the Multiple Translation Corpus means that
@@
 corresponding human-targeted reference aligned at the sentence level,
-and an associated so we have available the HTER score at
+so we have available the HTER score at
 both the sentence and document level.

### 2009-08-11 mtjournal.tex
repaired description of T features

@@
   specifically, error rates for the $1g$ and $2g$ decompositions.
-\item[T]: the 11 subscores from TERp, which include matches, insertions, deletions, substitutions, shifts, and
-  synonym and stem matches, edit distances between paraphrases, and the. 
+\item[T]: the 11 subscores from TERp, which include matches, insertions, deletions, substitutions, shifts,
+  synonym and stem matches, and four paraphrase edit scores.
 \end{description}

### 2009-08-11 mtjournal.tex
inline citation and notation fix

@@
 PCFG parser with deterministic head-finding, while 
-\cite{owczarzak07evaluatingmt} extract the semantic dependency
+\inlinecite{owczarzak07evaluatingmt} extract the semantic dependency
 relations from an LFG parser \cite{cahill04lfg}.
@@
 judgments of fluency and adequacy.  In other work, e.g.\
-\cite{snover06ter}, measures are compared to
+\inlinecite{snover06ter}, measures are compared to
 human-targeted TER (HTER), a distance to a human-revised reference
@@
 fluency and adequacy, using the 
-\cite{owczarzak07evaluatingmt} paradigm to provide comparisons and
+\inlinecite{owczarzak07evaluatingmt} paradigm to provide comparisons and
 select a best case configuration, Expected Dependency Pair Match
@@
 In table~\ref{tab:hterperdoc}, we show per-document Pearson's $r$
-between $\overline{\textrm{EDPM}}$ and $\overline{\textrm{HTER}}$, as well as the $\overline{\textrm{TER}}$ and $\overline{\textrm{BLEU}}$$_4$
+between $\overline{\textrm{EDPM}}$ and $\overline{\textrm{HTER}}$, as well as the $\overline{\textrm{TER}}$ and $\overline{\textrm{BLEU}_4}$
 baselines. EDPM has the highest correlation in each of the
